### 2020-04-12 TOG/note.tex
simplify

\section{Blog}

\begin{description}
\item[December 18, 2016]
{
\qisun{
I am starting to double this as a tutorial on how to use paper drafts to manage project progress.
This is still work in progress and not yet done, but I hope to finish a draft within the next few months.
}%qisun
}%item

\end{description}

### 2020-12-31 SIG21/0-abstract.tex
download latex template from official website and set SIG21/main.tex as entry point.

\begin{abstract}
Writing papers and managing projects are essential for conducting research.
A research paper describes the final outcomes, but can also help manage intermediate progress.
This document provides a template for writing papers and describes how to manage projects via drafts.

Start the abstract when you have some rough ideas, before they are forgotten amid your busy lives.
It does not need to be fancy, as long as you can understand your own writing later.
The eventual abstract should be the minimalist version of the entire paper, with key information outlined in the introduction.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
% End generated code
%

### 2020-12-31 SIG21/1-intro.tex
download latex template from official website and set SIG21/main.tex as entry point.

\section{Introduction}

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note

Writing research documents, such as conference/journal/white papers, patent applications, grant proposals, and course reports, is a core activity for many poor souls including professors, researchers, engineers, and students.
Since we must do it one way or another, sooner or later, we might as well do it as happily and effectively as possible.
Here are my personal practices and experiences, which I have repeatedly shared with my collaborators and students throughout all these years\note{~to the point that I finally decided to write up to save the troubles}.

\note{
What prior works have done, and why they are not adequate.
(Note: this is just high level big ideas. Details should go to a previous work section.)
}%note

Some people, including very successful ones, write papers only at the end of a project, like days if not hours prior to a submission deadline.
This almost always leads to total chaos and breakdown, unless you have other means to keep track and organize all the relevant pieces of information.

\note{
What our method has to offer, sales pitch for concrete benefits, not technical details.
Imagine we are doing a TV advertisement here.
}%note

This document provides a template for writing research papers, and describes how to manage project progress via iterating drafts.
Doing this correctly can help you achieve productive research and happy life, e.g. working anywhere anytime without synchronous meetings with your collaborators.

\note{
Our main idea, giving people a take home message and (if possible) see how clever we are.
}%note

I learned from my PhD adviser to start writing from the moment I have the faintest idea, and gradually update the drafts to reflect progress.
During my first job with NVIDIA, I learned why and how revision control is essential for sharing and editing source files, especially among large engineering teams.
The iterative nature of writing and coding matches well with the capability of revision control.
Thus I prefer Latex + git/svn (e.g. github, bitbucket, and the svn server under my hosting service), even though I can still manage with MS Word + cloud drive (e.g. Google drive, Box, and Dropbox). 
A good computer scientist writes papers like programs, and manages Latex files like source codes.
These repos are external memories and communication mediums for the collective brains of my teams.

\note{
Our algorithms and methods to show technical contributions and that our solutions are not trivial.
}%note

\input{TOG/example_fig}

Translating thoughts into words, diagrams, equations, pseudo-codes, or codes is a good way to clarify and consolidate.
If you are writing a graphics/HCI paper, design the figures so that they alone can provide a high level picture of the main points.
As exemplified in \Cref{fig:example}, a figure can contain images, drawings, or combinations.

\note{
Results, applications, and extra benefits.
}%note

You can use the associated files as a starting template for your research papers.
Start with the \filename{Makefile}, and notice the two build targets: final for official submission and public disclosure, and draft for internal sharing among your collaborators (including yourself).

### 2020-12-31 SIG21/4-method.tex
download latex template from official website and set SIG21/main.tex as entry point.

\section{Method}
\label{sec:method}

Written words are more concrete than ideas in your head.
Math equations and pseudo-codes are even more concrete than words, and can help, even though not essential, for exposition.
Programs that produce the actual results are the most concrete, but the amount and level of information is not suitable for a research paper.
I like to write down math and pseudo-code first as a way to design the algorithms before actual implementation.

\subsection{Math}
\label{sec:math}
\begin{align}
\energy = \mass \lightspeed^2
\label{eqn:emc2}
\end{align}

Always use defined instead of naked math symbols for clarify and consistency.
For example, in \Cref{eqn:emc2} I define all symbols inside \filename{TOG\\symbols.tex} so later if I need to change $\energy$ from $E$ to $e$ I just need to change one line in \filename{TOG\\symbols.tex} instead of chasing $E$ everywhere.
This can save you a lot of time and sanity if you have many math symbols.

\subsection{Code}
\label{sec:code}

\input{TOG/dart_throw_alg}

Pseudo-code can help summarize and explain the algorithms, but only if presented with sufficient clarity and simplicity. 
\Cref{alg:dart_throw} is an example.
All algorithms should be understandable from the main texts without looking at the pseudo-codes, which are usually more suitable for summarization than explanation.

### 2020-12-31 TOG/title_author.tex
download latex template from official website and set SIG21/main.tex as entry point.

\title{Using Paper Drafts to Manage Research Projects}

\author{Anonymous Authors}

### 2021-01-05 TOG/0-abstract.tex
v0.01: title, abstract, submission ID

\begin{abstract}
Streaming-based graphics has revolutionized the entertainment industry. It transfers the computational pressure from individual personal computers to clustered servers. This unifies development/consumption experience and reduces local storage.
Virtual and augmented realities are the future medium that would significantly benefit from the cloud computing via reliving their currently bulky wires, demands for high-end personal computers, and consequently, low resolution/field-of-view. 
However, the current off-the-shelf streaming solutions apply only to console-based platforms due to their lower requirement of latency and quality, thus transmission performance.

Tailored for the future cloud-based virtual reality, we present the first gaze-contingent 3D neural representation and view synthesis method. Our method considers both visual acuity and neural representation for 6DoF immersive interaction. Both objective analysis and subjective studies evidenced our effectiveness of significantly reducing local storage (\warning{xxx MB} for any high quality 3D scene), transmission latency (\warning{xxx ms}), and rendering time (\warning{xxx ms}).
\end{abstract}

### 2021-01-05 TOG/1-introduction.tex
v0.01: title, abstract, submission ID

--- 2020-12-31 SIG21/1-intro.tex
@@
 
-Writing research documents, such as conference/journal/white papers, patent applications, grant proposals, and course reports, is a core activity for many poor souls including professors, researchers, engineers, and students.
-Since we must do it one way or another, sooner or later, we might as well do it as happily and effectively as possible.
-Here are my personal practices and experiences, which I have repeatedly shared with my collaborators and students throughout all these years\note{~to the point that I finally decided to write up to save the troubles}.
+Started from remote desktops and online videos, cloud-based computing has show unprecedented potential to reshape computer graphics pipeline by completely relieving end-user's computational workload thus device bulkiness.
+In recent years, we have observed growing interactive streaming services such as Google Stadia or NVIDIA Shield.
+However, these off-the-shelf services have only been deployed in less latency sensitive platforms such as TV-based game consoles, due to the inevitable latency (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}.
+Immersive medias, such as virtual and augmented realities, demand remarkably lower latency and higher fidelity/speed for sickness-free and natural experience.
+Due to the consequently massive spatial-temporal data volume, there is no solution for streaming frames from server to individual end users today.
 
@@
 }%note
-
-Some people, including very successful ones, write papers only at the end of a project, like days if not hours prior to a submission deadline.
-This almost always leads to total chaos and breakdown, unless you have other means to keep track and organize all the relevant pieces of information.
 
@@
 
-\input{TOG/example_fig}
+\input{example_fig}
 

### 2021-01-12 TOG/0-abstract.tex
Updates from Overleaf

@@
 \begin{abstract}
-Streaming-based graphics has revolutionized the entertainment industry.
+%Streaming-based graphics has revolutionized the entertainment industry.
 It transfers the computational pressure from individual personal computers to clustered servers.
 This unifies development/consumption experience and reduces local storage.
-Virtual and augmented realities are the future medium that would significantly benefit from the cloud computing via reliving their currently bulky wires, demands for high-end personal computers, and consequently, low resolution/field-of-view.
-However, the current off-the-shelf streaming solutions apply only to console-based platforms due to their lower requirement of latency and quality, thus transmission performance.
+%Virtual and augmented realities are the future medium that would significantly benefit from the cloud computing via reliving their currently bulky wires, demands for high-end personal computers, and consequently, low resolution/field-of-view.
+%However, the current off-the-shelf streaming solutions apply only to console-based platforms due to their lower requirement of latency and quality, thus transmission performance.
 
-Tailored for the future cloud-based virtual reality, we present the first gaze-contingent 3D neural representation and view synthesis method.
-Our method considers both visual acuity and neural representation for 6DoF immersive interaction.
-Both objective analysis and subjective studies evidenced our effectiveness of significantly reducing local storage (\warning{xxx MB} for any high quality 3D scene), transmission latency (\warning{xxx ms}), and rendering time (\warning{xxx ms}).
+Traditional high-quality 3D graphics requires large volume of detailed data to be rendering.
+This compromises efficiency and local storage.
+In recent years, cloud-based streaming and neural representations have significantly revolutionized the pipeline by redistributing and compressing the computation/storage.
+However, they typically suffer from high latency and low quality for practically large virtual scenes with 6 degree-of-freedom camera motions.
+These problems are particularly exacerbated in immersive medias such as virtual and augmented reality due to the demand of high frequency and resolution.
+
+Tailored for the future cloud-based, portable, and low-storage VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis approach.
+Our method mutually bridges both visual acuity and neural representation for 6DoF immersive interaction.
+Moreoever, it optimizes between temporal computational latency and spatial predictive precision.
+Both objective analysis and subjective studies evidenced our effectiveness of significantly reducing local storage (\warning{xxx MB, xx\% size reduction} for large 3D scenes) while preserves original perceptual quality without introducing additional latency.
 \end{abstract}

### 2021-01-12 TOG/1-introduction.tex
Updates from Overleaf

--- 2020-12-31 SIG21/1-intro.tex
@@
 
-Writing research documents, such as conference/journal/white papers, patent applications, grant proposals, and course reports, is a core activity for many poor souls including professors, researchers, engineers, and students.
-Since we must do it one way or another, sooner or later, we might as well do it as happily and effectively as possible.
-Here are my personal practices and experiences, which I have repeatedly shared with my collaborators and students throughout all these years\note{~to the point that I finally decided to write up to save the troubles}.
+With the advent of remote desktops and streaming services, cloud-based computing has shown unprecedented potential reshaping traditional computer graphics pipeline by completely relieving end-user's computational workload, local storage volume, thus advance device portability.
+Recent years has seen growing interactive streaming services such as Google Stadia or NVIDIA Shield.
+However, these off-the-shelf services have only been deployed in less latency-sensitive platforms such as TV-based game consoles, due to the inevitable latency (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}.
+Immersive media such as virtual and augmented reality (VR/AR), however, demand remarkably lower latency and higher fidelity/speed for sickness-free and natural experience.
+Due to the consequently massive spatial-temporal high-quality data volume, there is no available solution today that is capable of streaming such dense content to the individual end user.
 
@@
 }%note
+\qisun{(Jan 5, 2021) We should include two main fields: neural representation and foveated rendering}
 
-Some people, including very successful ones, write papers only at the end of a project, like days if not hours prior to a submission deadline.
-This almost always leads to total chaos and breakdown, unless you have other means to keep track and organize all the relevant pieces of information.
+Recent years have seen a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets.
+Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf} and many more.
+However, existing solutions suffer from either high time consumption or low image fidelity.
+
+To accelerate real-time rendering tailored for VR, gaze-contingent foveated rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Patney:2016:TFR}.
+This is achieved by harnessing high FoV head-mounted displays and high precision eye-tracking technologies.
+Unfortunately, foveated rendering approaches still require full local 3D asset storage, resulting in a significant loading time and memory shortage.
 
@@
 
-This document provides a template for writing research papers, and describes how to manage project progress via iterating drafts.
-Doing this correctly can help you achieve productive research and happy life, e.g. working anywhere anytime without synchronous meetings with your collaborators.
+In this paper, we present the first gaze-contingent real-time 6DoF neural synthesis of scenes, aiming at the next generation cloud-based wearable immersive platforms.
+Specifically, our method significantly reduces 3D data storage (\warning{xxx ms}), and rendering time (\warning{xxx ms}) compared with the existing methods.
+It enables low-latency interactive VR/AR 6DoF viewable 3D virtual content without compromising perceived quality.
+Tailoring for future light weight, power and memory efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for any high quality 3D scenes.
 
@@
 
-I learned from my PhD adviser to start writing from the moment I have the faintest idea, and gradually update the drafts to reflect progress.
-During my first job with NVIDIA, I learned why and how revision control is essential for sharing and editing source files, especially among large engineering teams.
-The iterative nature of writing and coding matches well with the capability of revision control.
-Thus I prefer Latex + git/svn (e.g. github, bitbucket, and the svn server under my hosting service), even though I can still manage with MS Word + cloud drive (e.g. Google drive, Box, and Dropbox). 
-A good computer scientist writes papers like programs, and manages Latex files like source codes.
-These repos are external memories and communication mediums for the collective brains of my teams.
+In this work, we take an interdisciplinary approach achieving 6-DoF deep neural representation of complex scenes and gaze-contingent real-time rendering.
+This is realized by our novel data representation and a coarse-to-fine prediction of perceptually high quality imagery while balancing the latency.
 
@@
 
-\input{TOG/example_fig}
-
-Translating thoughts into words, diagrams, equations, pseudo-codes, or codes is a good way to clarify and consolidate.
-If you are writing a graphics/HCI paper, design the figures so that they alone can provide a high level picture of the main points.
-As exemplified in \Cref{fig:example}, a figure can contain images, drawings, or combinations.
+Specifically, we represent an arbitrary virtual 3D scene as a compact neural representation in a viewer centered inside-out fashion. 
+We account for the challenges such as low quality/resolution/performance common to existing 3D deep neural representation approaches, we further leverage the human visual perception to synthesize gaze-contingent and vision-matched immersive frames, without compromising the speed and image fidelity.
+The method is further optimized with in the spatio-temporal domain to balance the latency and precision.
 
@@
 
-You can use the associated files as a starting template for your research papers.
-Start with the \filename{Makefile}, and notice the two build targets: final for official submission and public disclosure, and draft for internal sharing among your collaborators (including yourself).
+We validate our approach by conducting both numerical analysis and user evaluations on commercially available AR/VR display devices.
+They both reveal the effectiveness of our method on delivering a real-time VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, without compromising on local storage volume.
+Our main contributions include the following:
+\begin{itemize}
+    \item A low-latency and significantly low-storage 6DoF immersive application that preserves perceptual quality
+    \item A gaze-contingent neural representation and synthesis method tailored for first-person viewing
+    \item A spatio-temporally joint optimization for balancing latency and perceptual quality
+\end{itemize}

### 2021-01-12 TOG/2-prior.tex
Updates from Overleaf

\section{Related Work}
\label{sec:prior}

\subsection{Deep View Synthesis}

\subsection{Gaze-Contingent Rendering}
\cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}

### 2021-01-12 TOG/4-method.tex
Updates from Overleaf

\section{Method}
\label{sec:method}
As visualized in \Cref{fig:pipeline}, our method consists of three major components: foveal-periphery image synthesis (\Cref{sec:method:net}), and real-time blending with optimized spatial-temporal precision (\Cref{sec:method:blending}).

%\subsection{Pipeline}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\linewidth]{TOG/figs/sys_pipeline.png}
    \Caption{System pipeline.}
    {%
    }
    \label{fig:pipeline}
\end{figure*}

%To meet the requirement of real-time, use fewer channels/layers, much fewer samples, predict a coarse fovea image, then refine using local view guidance

\subsection{GazeNet}
\label{sec:method:net}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\linewidth]{TOG/figs/spher_net_flow.png}
    \caption{GazeNet Network.}
    \label{fig:my_label}
\end{figure*}

%Follow NeRF's idea, represent scene using a MLP but in spherical space because we are inside-out (viewer is located at center area and look around)

\qisun{(Jan 9, 2021) Avoid usage of subsubsection.}
\paragraph{Multi spherical representation}
As reviewed in \autoref{sec:prior}, prior work contributed a lot to represent a static scene via object-focused manner. Inspired by that, we investigate on how to represent a static scene for VR environment to render real-timely without loading the entire scene. However, VR environment is user-centered. That is, unlike the object-focused ``outside-in'' way such as \cite{mildenhall2020nerf}, the views are highly varied in an ``inside-out'' fashion. 
Consequently, the overlaps among views are diversified, resulting in training challenges for large virtual scenes.
To address this problem, we depict a VR scene as concentric spherical volumes with the center as the original point, as visualized in \Cref{fig:teaser:scene}.

%\begin{figure}[htb]
%    \centering
%    \includegraphics[width=0.96\linewidth]{TOG/figs/spher_coord.pdf}
%%    \caption{Visualization of our multi-spherical coordinate system.}
    \label{fig:method:coordinate}
%\end{figure}
% different position (10cm working, 50cm not working),
% different rotation

\dnc{(Jan 11) Ray to spherical coord} For a sphere with radius $r$, the intersection of ray and sphere $\pt_r$ can be calculated by solving:

\begin{equation}
    \|\pt_r\|^2=\|\rayo + k\rayd\|^2=r^2
\end{equation}

Convert $\pt_r$ to spherical coordinate $(r, \theta, \phi)$:

\begin{equation}
    \begin{cases}
        \theta = \mathrm{atan2}(x_{\pt_r}/z_{\pt_r})\\
        \phi = \mathrm{acos}(y_{\pt_r} / r)
    \end{cases}
\end{equation}

\dnc{(Jan 11) A simple derivation for the requirement of depth range and samples} Take 2D as an example. Consider two view centers $\rayo_1$, $\rayo_2$ on X axis and a point $\pt$ in 3D space. The radius of nearest sampling sphere of $\pt$ is $r_l$. The intersections of the sphere and rays from $\rayo_1$, $\rayo_2$ to $\pt$ are $\pt'_{1,r_l}$ and $\pt'_{2,r_l}$, respectively. $\Delta\pt'_{r_l}=\|\pt'_{2,r_l}-\pt'_{1,r_l}\|$ is small enough, then we have:

\begin{equation}
    \Delta\pt'_{r_l} \approx \Delta\rayo \frac{|z_\pt - r_l|}{z_\pt}
\end{equation}

\begin{equation}
    hello
\end{equation}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/depth_disparity.pdf}
    \caption{Depth disparity}
    \label{fig:method:eccentricity}
\end{figure}

With this coordinate system, the input of individual element in the 3D spaces is defined as $p=(r, \theta, \phi)$ and the output is color information $c=(r,g,b)$ and volume density $d$. We train the scene representation with an multilayer perception (MLP) network first. 
For the training uniformity and run-time efficiency, we do not distinguish different eccentricities during the training stage. Instead, the spatially variant visual acuity is represented as varying field of views in the rendering, i.e., data creation.
Specifically, we depict the whole visual field as three semantic layers the fovea ({0-10 deg}), mid-periphery {6-60 deg}, and far periphery {y-110 deg}. In each frame, The uniformized networks returns three images with a identical resolution yet, as seen from the separation range in deg, gradually larger areas along eccentricity.

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/layers_together_zh.pdf}% try not overlapping the text on the image since we want the reviewers to see each pixel clearly
    \caption{Visualization of the output at three eccentricity ranges.}
    \label{fig:method:eccentricity}
\end{figure}

%estimate the fovea, mid-peripheral, peripheral field of view for a novel view next, fine-tune the fovea result with local view streamed from the cloud, and lastly warp/blend the three results for VR display.

\paragraph{Input encoding}
Since the input of our representation network is only three dimensional, we map the inputs to a higher dimensional space via high-frequency functions. \qisun{(Jan 9, 2021) What is bad if we have low-dim input? Harder to converge, higher error etc.?} \zh{(Jan 9) similar to NeRF, three components are not sufficient for achieving state-of-the-art quality, as demonstrated in Section 6.4). We introduce a positional encoding of the input coordinates that assists the MLP in
representing high-frequency functions.}
Thus we encoded the input through function:
\begin{align}
\iota (p) = (\sin(2^0 p), \cos(2^0 p), ..., \sin(2^{L-1}p, \cos(2^{L-1}p))
\label{eqn:inputencode}
\end{align}
The function $\iota (\cdot)$ is applied to each of the 3D location representation ($r$, $\theta$, and $\phi$), we choose $L$ to be 10 in practice. Hence, the input dimension becomes 3 x 20 = 60 after encoding. The MLP network is composed with $N_{layer}$ fully-connected layers, each uses ReLu as activation function and $N_{channel}$ channels per layer and output the color and volume density.

\paragraph{Synthesis}
% the description of the volume density is the same as chapter 4 in NeRF paper.
Similar to \cite{mildenhall2020nerf}, the trained representation is then deployed to synethesis final images via ray marching. For accelerating the performance, we bounded the near/far range to $20$ and $50$ meters respectively. The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
\begin{figure}
    \centering
    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
    \caption{Visualization of ray marching bounding.}
    \label{fig:method:bounding}
\end{figure}
%For each pixel, we perform a ray from the camera and accumulate all the intersection's color weighted by the density accordingly between near bound and far bound. In practice, we experimented with different far bound values: 20 (meters) and 50 (meters), the difference between output is minimal. (\note{add figure for comparison}). We also experimented with the sampling number ($N_{sample}$) for accumulation between near bound and far bound to balance time cost and effects. Hence, we have an output image for the camera perspective.
However, as a gaze-contingent method, the compromised quality would be unnoticeable in periphery. Thus, we further refine the visual quality in foveal vision, as detailed in \Cref{sec:method:refine}.
Since we aim to render a novel view in real time and with perceptually high quality, we estimate the foveated image (10 degree), middle peripheral image (60 degree), and full peripheral image (110 degree) from the same novel view for further synthesis. Regarding foveated rendering, we estimate an image with size of 64 x 64, $N_{layer}=12$, $N_{channel}=64$, and $N_{sample}=16$. Regarding middle peripheral and full peripheral rendering, we estimate the two images with size of 256 x 256, $N_{layer}=8$, $N_{channel}=64$, and $N_{sample}=4$. We next upsample middle peripheral to ??? x ??? and full peripheral to 1440 x 1440 for blending.
\nothing{
\subsection{Refinement for Fovea Layer}
\label{sec:method:refine}
Due to the depth bounding and prediction errors, it is particularly challenging to preserve pixel-wise quality without compromising real-time performance. However, the human foveal vision is remarkably sensitive to subtle details across all frequency ranges. To 
To maximize the quality of foveated image, we utilize the limited but consistent network bandwidth by retrieving a local peripheral image from the cloud. The retrieval was realized by cloud-based rendering using users' upstreamed head and gaze parameters.
Due to the inevitable dual-way transmission, it is impossible to instantly obtain the image without latency. Displaying outdated frames have been studied as a major cause of simulator sickness.
}%nothing
\note{describe when to retrieve new full peripheral image.} We leverage local guidance views to fine tune the output foveated image. Local guidance view is decided according to current camera view. We chose the nearest 4 views (or 8 views) as reference views and estimate the color as well as volume density for all reference views. Every time we receive a new peripheral image from the cloud, we calculate the delta between the ground truth and our result for reference views for preparation. Then we warp the delta into target camera view, calculate the average and apply to the fovea result for each frame.

\subsection{Real-time Rendering}
% \subsection{Fovea-Peripheral Blending}
\label{sec:method:blending}
Lastly, we apply alpha blend on the fine-tuned fovea view and middle peripheral view, as well as on the middle peripheral view and full peripheral view via function smo/:

\subsection{Latency-quality joint optimization}
\label{sec:method:optimization}
Increasing rendering sampling naturally improves the image output quality. However, it increases the latency, causing quality drop stretched along time, and more critically, simulator sickness.
Inspired by \cite{Li:2020:TSP}, we perform a spatial-temporal joint optimization to determine the sampling that adapts to individual computational resources. This is achieved via a Pareto efficiency optimization.

$E_{q}\triangleq $

$E_{l}$ is defined as the loss introduced by latency. 
\begin{align}
    \int E_{l} \mathbf{d} t
\end{align}

\begin{figure*}
    \centering
    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \caption{Latency-quality joint-optimization.}
    \label{fig:optimization}
\end{figure*}

%\subsection{Cloud-Based System}
%\label{sec:method:system}
%Finally, we deploy our gaze-contingent synthesizer to a real-time cloud-based streaming system.

### 2021-01-12 TOG/5-result.tex
Updates from Overleaf

\section{Evaluation}
\label{sec:result}
% use one net to represent a complicated scene
% we still have other option for real-time, like streaming
\note{
Ablation studies for the network training/losses
Target scene | uniformly downsampled image | local foveated rendered image | network predicted images ---> at least 4 different scenes.
streaming time for 4-5 different scenes: full resolution vs our network prediction for “same data size”
image quality of rendered scene vs our network prediction for “same amount of streaming time”
User study results
Table on performance comparison, streaming time, PSNR, SSIM, etc
}%note
\subsection{User Study}
\label{sec:study:user}
\zh{(Jan 10) all the Qs are answered offline, will update the section soon.}

% In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
In this section, we report a perceptual experiment to investigate how users perceive our solution compared to the ground truth and state-of-the-art (we choose to apply NeRF on our dataset).

% \zh{usually we report the study design here too, like within-subject user study and analyze method, like ANOVA. But I am not sure what we should use here.}
% \zh{(Jan 12), if we include NeRF, I'd say we keep one of the down-sample conditions}

\paragraph{Stimuli}
We designed four conditions for the perceptual experiment (see \autoref{fig:results:comparison}). The first column is the target scene, which we treat as the ground truth. The size of the image per eye is 1440 x 1600. The second column is a uniformly down-sample image at size \warning{??? x ???}. Assuming our network bandwidth is \warning{???}, to stream the images from the cloud in real time, the largest size of the image per eye is up to \warning{??? = ??? / 50 fps / 2 eyes}. So we down sample the entire image with ratio \warning{???}.
% The third column enables foveated rendering. The foveated area is preserved as the original resolution while the rest of the image is down sampled. We apply the same alpha blending method that we used in our solution (see \autoref{sec:method:blending}). Similarly, we calculate the size of the foveated image based on the bandwidth. Hence the down-sample ratio of peripheral area is \warning{???}. 
The fourth column is generated by NeRF implementation. We assign $N_{sample}=16$. The last column is our result. We predict the image from the camera transformation, fine-tune the foveated layer with local guided view, and render the image by alpha blending the three layers. The duration of the stimuli was \warning{??} seconds for each condition. 
% \zh{unsure if we want to render real-time results according to users gaze movement, or show static images with predefined gaze positions for certain duration?}

\paragraph{Setup}
?? subjects (\warning{? females and ? males, $M=?$, $SD=?$}) participated in the study. They were not involved in the project previously. Each participant was asked to wear a HTC Vive Pro Eye headset during the experiment. The headset is connected to a Windows machine with \warning{Intel i9-7900X CPU @ 3.30 GHz}, 64GB RAM and one NVIDIA GTX 3080 graphics card. During the experiment, the participants wearing the headset explored our stimuli. We used the \textit{two-alternative forced choice} (2AFC) method for both of our studies.

\paragraph{Task: Our method v.s. NeRF v.s. Ground Truth}
We aim at collecting user's feedback in VR display. Each participant were provided with \warning{30?} trials where each trial consisted a foveated area indicator and a pair of two stimuli. We randomize the order of the pairs for counterbalancing. To control the images to be shown for each participant, we choose to render the results for all conditions on predefined gaze positions rather than passing user's real time gaze input to the system. We indicate the location of the foveated area to users in advance to all the conditions first. For each condition, participants were asked to report \zhtd{whether they think the rendered result is better}.

% \zh{Since we will show at least four conditions to each participant, are we gonna ask them to pick up ONE as the favorite or ask them to judge each condition as ACCEPTED or NOT ACCEPTED? Or we should apply 3 trials that each contains ours and one of other alternatives? For example, Condition1 v.s. ConditionOurs, Condition2 v.s. ConditionOurs, and Condition3 v.s. ConditionOurs}
To minimize the effect of accumulating discomfort during the experiment, we enforced at least a 60-second break between the two conditions in each trial, but the participants were instructed to take as much time as needed to recover.
% \zh{I think rendering static images rather than taking real gaze input is more controllable and good for comparison, because we cannot control users gaze input, thus the images shown to participants are not the same. If we wanna follow a within subject study design, we probably need to pre-generated the gaze path. HOWEVER, if it does not matter if the participants need to look at the same images, which means the study is not ``paired'', we can still take real gaze input. But I think it increases the pressure for implementation, and we are not designing a gaze tracking system actually. How do you think about it?}

\qisun{Here is what I think. The conditions, OURS, UNIFORM (i.e. downsampled), NERF could be described in the stimuli part. Then in the task paragraph, we simply compare any given pair of these 3 conditions.}
% \zh{no more down samples?}
\zh{no GT included?}

\paragraph{Results}
\begin{figure}[htb]
    \centering
    \includegraphics[width=0.96\linewidth]{example-image-a}
    \caption{The users' preference votes from our evaluation experiment.}
    \label{fig:results:2afc}
\end{figure}
The user preference votes are visualized in \autoref{fig:results:2afc}.
% apply binomial test
% or apply ANOVA, there is a value of votes for each condition and we want to know if there exists significant effects on different condition.

\paragraph{Discussion}

\subsection{Ablation Study}
% the performance of an AI system by removing certain components, to understand the contribution of the component to the overall system

\subsection{Visual Quality}
As a complementary to the subjective measurement \Cref{sec:study:user}, we evaluate the perceived quality in an objective fashion. Specifically, we compare our method with both full resolution rendering and \cite{mildenhall2020nerf} with the same time consumption.

\begin{figure*}[htb]
    \centering
    \begin{minipage}{0.32\linewidth} %full res
        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_gt1.png}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_gt1.png}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
    \end{minipage}
    \begin{minipage}{0.32\linewidth} %ours
        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_fovea_result1.png}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_fovea_result1.png}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
    \end{minipage}
    \begin{minipage}{0.32\linewidth} % NERF
        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}

        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
    \end{minipage}

    \caption{Comparing our synthesis method with full resolution rendering and alternative solutions.}
    \label{fig:results:comparison}
\end{figure*}

\subsection{Spatial-Temporal Optimality}
%\zh{about streaming? network?}

### 2021-01-12 TOG/teaser_fig.tex
Updates from Overleaf

%\begin{figure}[htb]
\begin{teaserfigure}
  \centering
  \subfloat[scene representation]{
    \label{fig:teaser:scene}
    \includegraphics[width=0.45\linewidth]{TOG/figs/scene.pdf}
  }%subfloat
  \subfloat[latency]{
    \label{fig:teaser:latency}
    \includegraphics[width=0.25\linewidth]{example-image-a}
    % nerf: 108s/per infer
  }%subfloat
  \subfloat[quality]{
    \label{fig:teaser:quality}
    \includegraphics[width=0.25\linewidth]{example-image-a}
  }%subfloat
 \Caption{Illustration of our system.}
 {%
\subref{fig:teaser:scene} shows visualizes our active-viewing tailored scene representation.
Comparing with \protect\cite{mildenhall2020nerf}, \subref{fig:teaser:latency} shows our benefits of reduced latency.
As a zoom-in, \subref{fig:teaser:quality} compares our foveal quality preservation.
\qisun{Cross out half-half in  \subref{fig:teaser:hmd}. TODO for myself: I should make the background scene semi-transparent. Will do so soon.}
 }
 \label{fig:teaser}
%\end{figure}
\end{teaserfigure}

### 2021-01-12 TOG/title_author.tex
v0.01: title, abstract, submission ID
Updates from Overleaf

@@
-\title{Using Paper Drafts to Manage Research Projects}
+\title{Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display}
+\title{Immersive 6DoF Streaming with Gaze-Contingent Neural Synthesis} 
+\title{Gaze-Contingent Neural Synthesis for Low Latency VR in-the-Cloud} 
+\title{GazeNet: Gaze-Contingent Neural Synthesis for Low Storage VR} 
+% Representing Scenes as Neural Radiance Fields for View Synthesis
+
+% perceptually-aware neural synthesis for low latency 6dof VR?
+% perceptual neural rendering for ultra fast 6dof VR
 

### 2021-01-19 TOG/1-introduction.tex
snapshot

@@
 (Note: this is just high level big ideas. Details should go to a previous work section.)
-}%note
+}%notez
 \qisun{(Jan 5, 2021) We should include two main fields: neural representation and foveated rendering}
@@
     \item A low-latency and significantly low-storage 6DoF immersive application that preserves perceptual quality
-    \item A gaze-contingent neural representation and synthesis method tailored for first-person viewing
+    \item A gaze-contingent scene parameterization and neural representation/synthesis method tailored for first-person viewing
     \item A spatio-temporally joint optimization for balancing latency and perceptual quality

### 2021-01-19 TOG/2-prior.tex
snapshot

@@
 
-\subsection{Deep View Synthesis}
+\subsection{Image-based View Synthesis}
+Image-based rendering (IBR) has been proposed in computer graphics to complement the traditional 3D mesh + ray propagation pipelines \cite{levoy1996light,gortler1996lumigraph}.
+Benefiting from the 2D representation, it delivers pixel-wise fast rendering quality without compromising performance.
+However, one major limitation of IBR is the demand of large volume pre-captured data.
+This requires significant large volume of data and limited view ranges.
+To address this problem, recent years have witnessed extensive research on synthesizing novel views instead of storing.
+Examples include synthesizing light fields \cite{LearningViewSynthesis}, panoramas \cite{Lin:DeepPanorama}, and volumetric videos \cite{Broxton:immersiveLF}.
+However, despite being more compact, the synthesis is usually in a locally interpolated fashion than being globally consistent.
 
-\subsection{Gaze-Contingent Rendering}
-\cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}
+\subsection{Implicit Scene Representation for Neural Rendering}
+To fully represent a 3D object and environment without storing images and meshes, neural representations have drawn attentions in both computer graphics and computer vision communities.
+With a implicit deep neural network and 3D world tailored format, it generates an objects appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf}.
+However, current implicit representations focuses on ``out-side-in'' view with single objects.
+In immersive medias, the cameras are typically user-centered, resulting in lack of coverage thus quality drop.
+Instead of full scene representation, our method is VR tailored via representing the scene for user-centered views.
+
+\subsection{Gaze-Contingent Virtual Reality Rendering}
+Both rendering and neural synthesis suffers from computational load, thus computational and transmissional latency.
+With a full mesh representation, foveated rendering has been proposed to proposed to accelerate local rendering performance \cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}.
+This is typically achieved through high-field-of view displays (such as VR HMDs) and eye-trackers.
+In comparison, synthesizing views if far from being real-time or interactive.
+Our method introduces the vision-matched deep synthesis with both high quality and low latency, suitable for cloud-based and local representations.

### 2021-01-19 TOG/4-method.tex
snapshot

@@
 \label{sec:method}
-As visualized in \Cref{fig:pipeline}, our method consists of three major components: foveal-periphery image synthesis (\Cref{sec:method:net}), and real-time blending with optimized spatial-temporal precision (\Cref{sec:method:blending}).
+In this section, we discuss our approach to achieve gaze-contingent neural rendering for 6DoF immersive virtual reality.
+As illustrated in \Cref{fig:pipeline}, our method predominantly comprises of two steps: 1) visual eccentricity aware frame synthesis, discussed in \Cref{sec:method:net}, and real-time rendering with optimized spatial-temporal precision, discussed in \Cref{sec:method:blending}.
 
 %\subsection{Pipeline}
-\begin{figure*}[htb]
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=\linewidth]{TOG/figs/sys_pipeline.png}
-    \Caption{System pipeline.}
+    %\subfloat[pipeline]{\includegraphics[width=\linewidth]{TOG/figs/pipeline.pdf}
+    %\label{fig:pipeline}
+    %}
+    \subfloat[representation and annotation]{\includegraphics[width=0.8\linewidth]{TOG/figs/variables.pdf}}
+
+    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/fovea.png}}\vspace{1em}
+    \subfloat[mid-pheripery]{\includegraphics[width=0.3\linewidth]{TOG/figs/mid.png}}\vspace{1em}
+    \subfloat[far-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/periph.png}}
+
+    \subfloat[blend]{\includegraphics[width=0.9\linewidth]{TOG/figs/blended_with_mask.png}}
+
+    %\subfloat[GazeNet]{\includegraphics[width=\linewidth]{TOG/figs/spher_net_flow.png}
+    %\label{fig:net}
+    %}
+    \Caption{System and network visualization.}
     {%
+    \subref{fig:pipeline} shows breakdown of our system.
+    \subref{fig:net} visualizes our neural network design.
+    \warning{Label fovea, periphery FoVs in the figure}
     }
-    \label{fig:pipeline}
-\end{figure*}
+    \label{fig:system}
+\end{figure}
 
-%To meet the requirement of real-time, use fewer channels/layers, much fewer samples, predict a coarse fovea image, then refine using local view guidance
+% \subsection{First-Person Tailored Representation}
+\subsection{Egocentric View Representation}
+\label{sec:method:representation}
+\paragraph{Multi spherical representation}
+We tailor our method specifically for egocentric (first-person) view in immersive VR environments.
+Unlike the object-focused ``outside-in'' view synthesis~\cite{sitzmann2019deepvoxels,mildenhall2020nerf}, rendering of immersive ``inside-out'' scenes is challenging owing to the rapid variation of scene content.
+As a consequence of extreme diversity among overlapping views, neural representation of large virtual scenes typically result in significant artifacts as demonstrated in \cref{fig:teaser:scene}.
+We overcome this problem by depicting the rapidly varying 360-degree virtual scenes with concentric spherical volumes.
+This representation not only enables robust rendering of complex immersive environments at real-time rates, but also allows for 6DoF interaction for exploring virtual spaces, as will be discussed later.
 
-\subsection{GazeNet}
-\label{sec:method:net}
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\linewidth]{TOG/figs/spher_net_flow.png}
-    \caption{GazeNet Network.}
-    \label{fig:my_label}
-\end{figure*}
+Inspired by \cite{Broxton:immersiveLF,Lin:DeepPanorama}, we address this problem via depicting a virtual scene with concentric spherical volumes.
+\zh{shall we emphasize here that the root of concentric spherical volumes is the origin of the virtual scene/world?} \qisun{I intentionally didn't emphasize this part since I was wondering whether that gives people the impression of weak support of larger range translation.
+What do you think?}
 
-%Follow NeRF's idea, represent scene using a MLP but in spherical space because we are inside-out (viewer is located at center area and look around)
-
-\qisun{(Jan 9, 2021) Avoid usage of subsubsection.}
-\paragraph{Multi spherical representation}
-As reviewed in \autoref{sec:prior}, prior work contributed a lot to represent a static scene via object-focused manner.
-Inspired by that, we investigate on how to represent a static scene for VR environment to render real-timely without loading the entire scene.
-However, VR environment is user-centered.
-That is, unlike the object-focused ``outside-in'' way such as \cite{mildenhall2020nerf}, the views are highly varied in an ``inside-out'' fashion.
-Consequently, the overlaps among views are diversified, resulting in training challenges for large virtual scenes.
-To address this problem, we depict a VR scene as concentric spherical volumes with the center as the original point, as visualized in \Cref{fig:teaser:scene}.
-
+Specifically, as shown in \Cref{fig:teaser:scene},  
+%Consequently, the overlaps among views are diversified, resulting in training challenges for large virtual scenes.
+%To address this problem, we depict a VR scene as concentric spherical volumes with the center as the original point, as %visualized in \Cref{fig:teaser:scene}. 
 %\begin{figure}[htb]
@@
 %%    \caption{Visualization of our multi-spherical coordinate system.}
-    \label{fig:method:coordinate}
+%    \label{fig:method:coordinate}
 %\end{figure}
@@
 % different rotation
+For a sphere (located at origin point) with radius $\sphereRadius$, its intersection (if exists) with a directional ray $\{\rayo,\rayd\}$ is
+\begin{equation}
+\pt(\sphereRadius,\rayo,\rayd) = \rayo + \left({\left((\rayo\cdot\rayd)^2-\norm{\rayo}^2+\sphereRadius^2\right)}^{\frac{1}{2}}-\rayo\cdot\rayd\right)\rayd,
+    %\pt(\sphereRadius,\rayo,\rayd)=\rayo + \sphereRadius\rayd,
+\end{equation}
+where $\rayo$ and $\rayd$  are the ray's origin point and normalized direction respectively.
+%Convert $\pt_r$ to spherical coordinate $(r, \theta, \phi)$:
+%\begin{equation}
+%    \begin{cases}
+%        \theta = \mathrm{atan2}(x_{\pt_r}/z_{\pt_r})\\
+%        \phi = \mathrm{acos}(y_{\pt_r} / r).
+%    \end{cases}
+%\end{equation}
+%\qisun{(Jan 14) Todo for myself: I haven't done revision on this part yet. Will do so after the description becomes more understandable.}
 
-\dnc{(Jan 11) Ray to spherical coord} For a sphere with radius $r$, the intersection of ray and sphere $\pt_r$ can be calculated by solving:
-
+% ZH: coordiante system? Parameterization system, parameterized representation etc.
+In our first-person tailored neural representation, the scene is parameterized through the number ($\sphereNum$) and radii ($\mathbf{\sphereRadius}=\{\sphereRadius_i\},i\in [1,\sphereNum]$) of concentric spheres.
+Under this system, all the intersections are
 \begin{equation}
-    \|\pt_r\|^2=\|\rayo + k\rayd\|^2=r^2
+\left\{\pt(\sphereRadius_i,\rayo,\rayd) \right\}\ |\ i\in[1,2,\dots,\sphereNum].
 \end{equation}
 
-Convert $\pt_r$ to spherical coordinate $(r, \theta, \phi)$:
-
-\begin{equation}
-    \begin{cases}
-        \theta = \mathrm{atan2}(x_{\pt_r}/z_{\pt_r})\\
-        \phi = \mathrm{acos}(y_{\pt_r} / r)
-    \end{cases}
+Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt=\SpatialPt(\sphereRadius, \theta, \phi)$ represented in spherical coordinates, the ray connect them is with the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$.
+This ray may intersect with more than one spheres.
+Among them, the closest intersection point to $\SpatialPt$ is
+\begin{equation}\label{eq:closestPoint}
+\begin{aligned}
+    \hat{\pt}&(\rayo,\SpatialPt, \sphereNum, \mathbf{\sphereRadius}) \triangleq \\
+    &\pt(\sphereRadius_k,\rayo,\rayd(\SpatialPt,\rayo))\ |\ k=\arg\min_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right),
+\end{aligned}
 \end{equation}
 
-\dnc{(Jan 11) A simple derivation for the requirement of depth range and samples} Take 2D as an example.
-Consider two view centers $\rayo_1$, $\rayo_2$ on X axis and a point $\pt$ in 3D space.
-The radius of nearest sampling sphere of $\pt$ is $r_l$.
-The intersections of the sphere and rays from $\rayo_1$, $\rayo_2$ to $\pt$ are $\pt'_{1,r_l}$ and $\pt'_{2,r_l}$, respectively.
-$\Delta\pt'_{r_l}=\|\pt'_{2,r_l}-\pt'_{1,r_l}\|$ is small enough, then we have:
+% \begin{align}
+% \pt(\sphereRadius_k,\rayo,\rayd(\SpatialPt,\rayo))\ |\ 
+% k=\min (\arg\min_j(\norm{\norm{\SpatialPt}-\sphereRadius_j}), \sphereNum).
+% \label{eq:intersection}
+% \end{align}
 
+The formula \Cref{eq:closestPoint} and \Cref{fig:intersection} reveal our concentric representation's spatial approximation error due to its sparsity.
+Specifically we compute the offset distance across all $\SpatialPt$ in a 3D static scene
 \begin{equation}
-    \Delta\pt'_{r_l} \approx \Delta\rayo \frac{|z_\pt - r_l|}{z_\pt}
+\begin{aligned}
+\sparseError(\sphereNum, \mathbf{\sphereRadius})  &= \iint \norm{\SpatialPt-\hat{\pt}(\rayo,\SpatialPt, \sphereNum, \mathbf{\sphereRadius})} \mathbf{d}\SpatialPt \mathbf{d}\rayo,\\
+&\forall \{\rayo, \SpatialPt\} \text{ pair without occlusion in between},
+\label{eq:sparseError}
+\end{aligned}
 \end{equation}
+given it's parameters ($\sphereNum$,$\mathbf{\sphereRadius}$). Higher losses cause less precise spatial approximation, i.e., prediction error from a neural network.
+Specifically, we observe
+\begin{itemize}
+    \item Given a fixed min/max range of $\mathbf{\sphereRadius}$, $\sphereNum$ is negatively correlated to the spatial approximation error
+    \item With a fixed $\sphereNum$, the correlation between distribution of $\sphereRadius_j$ and content (i.e., distribution of $\SpatialPt$s) is positively correlated to the spatial approximation error
+\end{itemize}
+We determine the coordinate variables with a content-aware optimization, as further detailed in \Cref{sec:method:optimization}.
 
-\begin{equation}
-    hello
-\end{equation}
+\begin{figure}
+    \centering
+    \subfloat[Visualization of \protect\Cref{eq:sparseError}]{\includegraphics[width = 0.47\linewidth]{example-image-a}\label{fig:intersection:visualization}}
+    \subfloat[Plot]{\includegraphics[width = 0.47\linewidth]{example-image-a}}
+    \caption{The relationship among $\sphereRadius$, N}
+    \label{fig:intersection}
+\end{figure}
+%Under this definition, we obtain the nearest sphere ID of a given 3D virtual point $\pt=\pt(r, \theta, \phi)$ as:
+%\begin{equation}
+%    \sphereID(\pt,\sphereNum, \mathbf{\sphereRadius}) = \min (\arg\min_k(\norm{\norm{\pt}-\sphereRadius_k}), \sphereNum),
+%\label{eq:sphereID}
+%\end{equation}
+%where $r=\norm{\pt}$  and  $\theta, \phi$ are the radius and angles of the spherical coordinate
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.96\linewidth]{TOG/figs/depth_disparity.pdf}
-    \caption{Depth disparity}
-    \label{fig:method:eccentricity}
-\end{figure}
+%Consider two rays $\{\rayo_i,\rayd_i\},i=1,2$ on X axis and a point $\pt$ in 3D space. The intersections of $\pt$'s nearest sphere and the two rays are:
+%\begin{equation}
+%%\pt^\prime (\pt,\rayo_i,\rayd_i) = \rayo_i + \sphereRadius_{\sphereID(\pt,\sphereNum)} \rayd_i.
+%\label{eq:raySphereIntersection}
+%\end{equation}
+%from $\rayo_1$, $\rayo_2$ to $\pt$ are $\pt'_{1,r_l}$ and $\pt'_{2,r_l}$, respectively. %$\Delta\pt'_{r_l}=\|\pt'_{2,r_l}-\pt'_{1,r_l}\|$ is small enough, then we have:
+%Then, we obtain
+%\begin{equation}
+%\pt^\prime (\pt,\rayo_1,\rayd_1) - \pt^\prime (\pt,\rayo_2,\rayd_2) \approx 
+%    \Delta\pt'_{r_l} \approx \Delta\rayo \frac{|z_\pt - r_l|}{z_\pt}
+%\label{eq:deltaApprox}
+%\end{equation}
+%\qisun{(Jan 14, 2021) Fill in the eq. above.}
 
-With this coordinate system, the input of individual element in the 3D spaces is defined as $p=(r, \theta, \phi)$ and the output is color information $c=(r,g,b)$ and volume density $d$.
-We train the scene representation with an multilayer perception (MLP) network first.
-For the training uniformity and run-time efficiency, we do not distinguish different eccentricities during the training stage.
-Instead, the spatially variant visual acuity is represented as varying field of views in the rendering, i.e., data creation.
-Specifically, we depict the whole visual field as three semantic layers the fovea ({0-10 deg}), mid-periphery {6-60 deg}, and far periphery {y-110 deg}.
-In each frame, The uniformized networks returns three images with a identical resolution yet, as seen from the separation range in deg, gradually larger areas along eccentricity.
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.96\linewidth]{TOG/figs/depth_disparity.pdf}
+%     \caption{Depth disparity}
+%     \label{fig:method:eccentricity}
+% \end{figure}
+%\qisun{(Jan 14, 2021) The figure above is not even referred in the current draft. Feel free to add them back whenever there is any real content or at least related to the textual content.}
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.96\linewidth]{TOG/figs/layers_together_zh.pdf}% try not overlapping the text on the image since we want the reviewers to see each pixel clearly
-    \caption{Visualization of the output at three eccentricity ranges.}
-    \label{fig:method:eccentricity}
-\end{figure}
+%\begin{figure}[htb]
+%    \centering
+%    \includegraphics[width=0.96\linewidth]{TOG/figs/layers_together_zh.pdf}% try not overlapping the text on the image since we want the %reviewers to see each pixel clearly
+%    \caption{Visualization of the output at three eccentricity ranges.}
+%    \label{fig:method:layer}
+%\end{figure}
 
 %estimate the fovea, mid-peripheral, peripheral field of view for a novel view next, fine-tune the fovea result with local view streamed from the cloud, and lastly warp/blend the three results for VR display.
+\subsection{Neural Synthesis}
+Our goal is to create perceptually identical quality frames in each frame.
+Thus, we train networks that synthesize the foveal ($\imageFoveal(\rayo,\rayd)$, $0-10$deg), mid-peripheral ($\imageMid(\rayo,\rayd)$, $6-60$deg), and far-peripheral ($\imageFar(\rayo,\rayd)$, $60-110$deg) images in real-time.
+Here, camera position $\rayo$ and gaze direction $\rayd$ are obtained from the eye-tracked HMD.
 
 \paragraph{Input encoding}
-Since the input of our representation network is only three dimensional, we map the inputs to a higher dimensional space via high-frequency functions.
+%Since the input of our representation network is only three dimensional, we map the inputs to a higher dimensional space via high-frequency functions.
 \qisun{(Jan 9, 2021) What is bad if we have low-dim input?
@@
 We introduce a positional encoding of the input coordinates that assists the MLP in
-representing high-frequency functions.}
-Thus we encoded the input through function:
+%representing high-frequency functions.}
+
+Similar to \cite{mildenhall2020nerf}, we extend the input dimensions for higher training degree-of-freedom. This is achieved via coding the input as:
 \begin{align}
@@
 \end{align}
-The function $\iota (\cdot)$ is applied to each of the 3D location representation ($r$, $\theta$, and $\phi$), we choose $L$ to be 10 in practice.
-Hence, the input dimension becomes 3 x 20 = 60 after encoding.
-The MLP network is composed with $N_{layer}$ fully-connected layers, each uses ReLu as activation function and $N_{channel}$ channels per layer and output the color and volume density.
+The function $\iota (\cdot)$ is applied to dimension of $\SpatialPt=(r_\SpatialPt,\theta_\SpatialPt, \phi_\SpatialPt)$.
+We train the scene representation with an multilayer perception (MLP) network. 
+In our experiment we choose $L=10$.\nothing{ Hence, the input dimension becomes 3 x 20 = 60 after encoding.} The MLP network is composed with $\mlpLayerNum$ fully-connected layers, each uses ReLu as activation function and $\mlpChannelNum$ channels per layer.
+The MLP outputs $c(\SpatialPt)\triangleq(r,g,b,d)$ for each 3D virtual voxel with our concentric parameterization in \Cref{sec:method:representation}.
+Here $(r,g,b)$ and $d$ represent the color and density respectively.
+
+\paragraph{Training}
+With concentric parameterization, we predicts  $c(\SpatialPt)$ for all $\SpatialPt$s along any given light ray $\{\rayo,\rayd\}$.
+\nothing{This follows a similar definition of \cite{mildenhall2020nerf}.} Note that besides the rays themselves, the prediction is also a function of  $\sphereNum$, $\sphereRadius_k$, and the network parameters ($\mlpLayerNum$/$\mlpChannelNum$) as evidenced in \Cref{eq:raySphereIntersection}.
+
+Note that foveal/peripheral retinal images requires high/low pixel-wise precision but low/high FoV (number of pixels).
+We trained two independent neural networks that are tailored for the perceptual nature.
+%For the training uniformity and run-time efficiency, we do not distinguish different eccentricities during the training stage.
+Instead, the spatially variant visual acuity is represented as varying field of views in the rendering, i.e., data creation.
+%Specifically, we depict the whole visual field as three retinal layers the fovea ({0-10 deg}), mid-periphery {6-60 deg}, and far periphery {60-110 deg}.
+\note{
+we estimate an image with size of 64 x 64, $N_{layer}=12$, $N_{channel}=64$, and $N_{sample}=16$.
+Regarding middle peripheral and full peripheral rendering, we estimate the two images with size of 256 x 256, $N_{layer}=8$, $N_{channel}=64$, and $N_{sample}=4$.
+We next upsample middle peripheral to ???
+x ???
+and full peripheral to 1440 x 1440 for blending.
+}
+Specifically, to incorporate the display hardware properties and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $64\times64$/$256\times256$/$1440\times1440$ respectively.
+Thus the $\imageFoveal$ has the highest spatial resolution of \warning{xx} pixel per degree (PDD), higher than those of $\imageMid$ \warning{yy PDD} and $\imageFar$ \warning{zz}.
+In each frame, three uniformized networks independently return three retinal images, as seen from the separation range in deg, gradually larger areas along eccentricity.
+An example is shown in \Cref{fig:system}.
 
@@
 % the description of the volume density is the same as chapter 4 in NeRF paper.
-Similar to \cite{mildenhall2020nerf}, the trained representation is then deployed to synethesis final images via ray marching.
-For accelerating the performance, we bounded the near/far range to $20$ and $50$ meters respectively.
-The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
+After training, the nerual network is then deployed to synthesize final images via ray marching.
+For accelerating the performance and quality (as compared in \Cref{fig:method:bounding}), we bound the min/max of $\mathbf{\sphereRadius}$.
+The bounding range, together with $\sphereNum$, $\mlpLayerNum$ and $\mlpChannelNum$, are determined from a spatial-temporal optimization, as detailed in \Cref{sec:method:optimization} %The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
 \begin{figure}
     \centering
-    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
-    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
+    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}\hspace{1em}
+    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}
     \caption{Visualization of ray marching bounding.}
@@
 Hence, we have an output image for the camera perspective.
-However, as a gaze-contingent method, the compromised quality would be unnoticeable in periphery. Thus, we further refine the visual quality in foveal vision, as detailed in \Cref{sec:method:refine}.
-Since we aim to render a novel view in real time and with perceptually high quality, we estimate the foveated image (10 degree), middle peripheral image (60 degree), and full peripheral image (110 degree) from the same novel view for further synthesis.
-Regarding foveated rendering, we estimate an image with size of 64 x 64, $N_{layer}=12$, $N_{channel}=64$, and $N_{sample}=16$.
-Regarding middle peripheral and full peripheral rendering, we estimate the two images with size of 256 x 256, $N_{layer}=8$, $N_{channel}=64$, and $N_{sample}=4$.
-We next upsample middle peripheral to ???
-x ???
-and full peripheral to 1440 x 1440 for blending.
+ %Specifically, the network for $\imageFoveal$ generates $64\times64$ resolution (\todo{xxx} deg) with 
+%However, as a gaze-contingent method, the compromised quality would be unnoticeable in periphery. Thus, we further refine the visual quality in foveal vision, as detailed in \Cref{sec:method:refine}.
+
+\subsection{Latency-Quality Joint Optimization}
+\label{sec:method:optimization}
+\paragraph{3D to image space}
+\Cref{eq:sparseError} depicts a generic approximation error for a given camera position and a 3D virtual space.
+However, typical 3D representations are not uniformly distributed points.
+Moreover, the network only predicts pixel colors/intensities but not depth.
+Thus, we further extend \Cref{eq:sparseError} to image space.
+For each camera view, its projection matrix can be defined by the ray that forms with the camera's position and viewing direction $\projectionMatrix(\rayo,\rayd)$.
+By extending from \Cref{eq:sparseError} to the captured image space data, we obtain
+\begin{align}
+\imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo, \rayd)  = \int \norm{\projectionMatrix(\rayo,\rayd)\cdot\SpatialPt-\projectionMatrix(\rayo,\rayd)\cdot
+\pt(\sphereRadius_k,\rayo,\rayd)}\mathbf{d}\SpatialPt 
+\label{eq:imageError}
+\end{align}
+
+\paragraph{Integrating with latency-introduced error}
+Increasing rendering sampling naturally improves the image output quality. However, it increases the latency, causing quality drop stretched along time, and more critically, simulator sickness.
+Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform a spatial-temporal joint optimization to determine the spherical sampling that adapts to individual computational resources.
+This is achieved via a latency-precision modeling and the corresponding optimization.
+
+%Specifically, we model the prediction image $P$ with gaze position ($G$) at time $t$ as $P_G(t)$ and the ground truth retinal image (from a simulated local rendering) as $\hat{P}_G(t)$.
+In practice, increasing the number of spheres $\sphereNum$ or number of MLP layers may affect the computational and transmissional latency but will improve the quality of individual frames (as in \Cref{eq:imageError}).
+Thus, we define the perceptual error as:
+\begin{equation}
+\begin{aligned}
+&\finalError(\sphereNum, \mathbf{\sphereRadius}) = \\
+&= \sum_t\int \norm{\projectionMatrix(\rayo_t,\rayd_t)\cdot\SpatialPt-\projectionMatrix(\rayo_{t-\latency},\rayd_{t-\latency})\cdot\pt(\sphereRadius_k,\rayo_{t-\latency},\rayd_{t-\latency})}\mathbf{d}\SpatialPt,
+%&\int  \norm{ \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_t, \rayd_t) - \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_{t-l(\sphereNum, \mathbf{\sphereRadius})}, \rayd_{t-l(\sphereNum, \mathbf{\sphereRadius})})) } \mathbf{d} t
+\end{aligned}
+\end{equation}
+where $\latency\triangleq\latency(\sphereNum, \mathbf{\sphereRadius})$ is the latency caused by alternating \warning{the representation coordinate} $\sphereNum, \mathbf{\sphereRadius}$ and network layers.
+
+As indicated in \cite{albert2017latency}, the latency shall be below the perceptual range \warning{$L=xx$ ms}.
+So, we determine the optimal $\{\sphereNum, \mathbf{\sphereRadius}\}$ to balance latency and precision as
+\begin{equation}
+\begin{aligned}
+    &\arg\min_{\sphereNum, \mathbf{\sphereRadius}} E(\sphereNum, \mathbf{\sphereRadius}), \\
+    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L
+\end{aligned}
+\end{equation}
+
+%$E_{q}\triangleq $
+%$E_{l}$ is defined as the loss introduced by latency. 
+%\begin{align}
+%    \int E_{l} \mathbf{d} t
+%\end{align}
+
+\begin{figure*}
+    \centering
+    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
+    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
+    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
+    \caption{Latency-quality joint-optimization.}
+    % talking about sample
+    % (a) we need several figures with gaze movement
+    \label{fig:optimization}
+\end{figure*}
+
+%\subsection{Cloud-Based System}
+%\label{sec:method:system}
+%Finally, we deploy our gaze-contingent synthesizer to a real-time cloud-based streaming system.
+
+\subsection{Real-time Rendering}
+% \subsection{Fovea-Peripheral Blending}
+\label{sec:method:blending}
+%Lastly, we apply alpha blend on the fine-tuned fovea view and middle peripheral view, as well as on the middle peripheral view and full peripheral view via function:
+
+In real-time, the \netName predicts the two eye's projected retinal images (3 for each, 6 in total).
+We then perform an image-based rendering to generate final frames for the VR head-mounted-displays.
+
+To enhance consistency of edge area between layers, our fragment shader retrieves the three frames and applied a pixel wise bilinear interpolation at each $10\%$ connecting areas between two adjunct layers.
+To further preserve peripheral quality due to its low PDD, we enhanced the contrast following the mechanism of \cite{Patney:2016:TFR}, the details are visualized in \Cref{fig:method:blending}.
+
+\begin{figure}
+    \centering
+    \subfloat[blending]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
+
+    \subfloat[Contrast enhancement]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
+    \caption{Our multi-layer real-time rendering system.}
+    \label{fig:method:blending}
+\end{figure}
+
 \nothing{
@@
 
-\subsection{Real-time Rendering}
-% \subsection{Fovea-Peripheral Blending}
-\label{sec:method:blending}
-Lastly, we apply alpha blend on the fine-tuned fovea view and middle peripheral view, as well as on the middle peripheral view and full peripheral view via function smo/:
-
-\subsection{Latency-quality joint optimization}
-\label{sec:method:optimization}
-Increasing rendering sampling naturally improves the image output quality. However, it increases the latency, causing quality drop stretched along time, and more critically, simulator sickness.
-Inspired by \cite{Li:2020:TSP}, we perform a spatial-temporal joint optimization to determine the sampling that adapts to individual computational resources.
-This is achieved via a Pareto efficiency optimization.
-
-$E_{q}\triangleq $
-
-$E_{l}$ is defined as the loss introduced by latency. 
-\begin{align}
-    \int E_{l} \mathbf{d} t
-\end{align}
-
-\begin{figure*}
-    \centering
-    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \caption{Latency-quality joint-optimization.}
-    \label{fig:optimization}
-\end{figure*}
-
-%\subsection{Cloud-Based System}
-%\label{sec:method:system}
-%Finally, we deploy our gaze-contingent synthesizer to a real-time cloud-based streaming system.
+\input{TOG/implementation}

### 2021-01-19 TOG/5-result.tex
snapshot

@@
 \label{sec:study:user}
-\zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
+% \zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
 
@@
 ??
-subjects (\warning{?
+users (\warning{?
 females and ?
@@
 Each participant was asked to wear a HTC Vive Pro Eye headset during the experiment.
-The headset is connected to a Windows machine with \warning{Intel i9-7900X CPU @ 3.30 GHz}, 64GB RAM and one NVIDIA GTX 3080 graphics card.
 During the experiment, the participants wearing the headset explored our stimuli.
@@
 
-\paragraph{Task: Our method v.s. NeRF v.s. Ground Truth}
-We aim at collecting user's feedback in VR display.
+\paragraph{Task}
+The task was a two-alternative-forced-choice (2AFC). 
 Each participant were provided with \warning{30?} trials where each trial consisted a foveated area indicator and a pair of two stimuli.
@@
 We indicate the location of the foveated area to users in advance to all the conditions first.
-For each condition, participants were asked to report \zhtd{whether they think the rendered result is better}.
+For each condition, participants were instructed to indicate which of the two conditions appeared to be with higher quality.
 
@@
 
-\qisun{Here is what I think.
+%\qisun{Here is what I think.
 The conditions, OURS, UNIFORM (i.e.
@@
 % \zh{no more down samples?}
-\zh{no GT included?}
+%\zh{no GT included?}
 
-\paragraph{Results}
+\paragraph{Results and discussion}
+\Cref{fig:results:2afc} plots the study results.
+
 \begin{figure}[htb]
@@
 \end{figure}
-The user preference votes are visualized in \autoref{fig:results:2afc}.
-% apply binomial test
+
+% apply binomial test https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.binom_test.html
 % or apply ANOVA, there is a value of votes for each condition and we want to know if there exists significant effects on different condition.
 
-\paragraph{Discussion}
+The results about indicated the similar perceptual quality between {\bf OURS} and {\bf GT}. However, the participants significantly observed the quality drop from {\bf NERF}.
 
-\subsection{Ablation Study}
-% the performance of an AI system by removing certain components, to understand the contribution of the component to the overall system
+To further validate the perceptual quality, we conducted measurements with objective metrics in the following section.
 
@@
 
+\begin{figure}
+    \centering
+    \subfloat[ANOVA on FWQI]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
+
+    \subfloat[FWQI gaze heatmap]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
+    \caption{Objective measurements on visual quality.}
+    \label{fig:result:visual}
+\end{figure}
+
 \begin{figure*}[htb]
@@
     \begin{minipage}{0.32\linewidth} %full res
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_gt1.png}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_gt4.png}}
 
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_gt1.png}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_gt3.png}}
 
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/bed_gt0.png}}
     \end{minipage}
     \begin{minipage}{0.32\linewidth} %ours
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_fovea_result1.png}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_our4_inset.pdf}}
 
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_fovea_result1.png}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_our3_inset.pdf}}
 
@@
     \begin{minipage}{0.32\linewidth} % NERF
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_nerf4_inset.pdf}}
 
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_nerf3_inset.pdf}}
+
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/bed_nerf0.png}}      
+    \end{minipage}
+
+    \caption{Comparing our synthesis method with full resolution (2nd column) rendering and NeRF (3rd column).}
+    {\zh{add inset for 2nd and 3rd col}}
+    \label{fig:results:comparison}
+\end{figure*}
+
+\begin{figure*}[htb]
+    \centering
+    \begin{minipage}{0.32\linewidth} %full res
+
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/study_gt0.png}}
+    \end{minipage}
+    \begin{minipage}{0.32\linewidth} %ours
 
@@
     \end{minipage}
+    \begin{minipage}{0.32\linewidth} % NERF
 
-    \caption{Comparing our synthesis method with full resolution rendering and alternative solutions.}
-    \label{fig:results:comparison}
+        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/study_nerf0.png}}        
+    \end{minipage}
+
+    \caption{Comparing our synthesis method with full resolution (2nd column) rendering and NeRF (3rd column) (Cont.).}
+    {\zh{add inset for 2nd and 3rd col}}
+    \label{fig:results:comparison2}
 \end{figure*}
 
+\subsection{Ablation Study}
+% the performance of an AI system by removing certain components, to understand the contribution of the component to the overall system
+
 \subsection{Spatial-Temporal Optimality}
-%\zh{about streaming? network?}
+
+% apply objective metric to Fig 7, aka different sample numbers

### 2021-01-19 TOG/teaser_fig.tex
snapshot

@@
     \label{fig:teaser:scene}
-    \includegraphics[width=0.45\linewidth]{TOG/figs/scene.pdf}
+    % \includegraphics[height=4.5cm]{TOG/figs/scene.pdf}
+    \includegraphics[height=4.5cm]{TOG/figs/teaser_1a.pdf} %try other colors a little bit
   }%subfloat
@@
     \label{fig:teaser:latency}
-    \includegraphics[width=0.25\linewidth]{example-image-a}
-    % nerf: 108s/per infer
+    \includegraphics[height=4.5cm]{TOG/figs/teaser_1b_small.pdf}
+    % \includegraphics[height=4.5cm]{TOG/figs/teaser.pdf}
+    % nerf: 15+s/per infer
   }%subfloat
-  \subfloat[quality]{
+  \subfloat[foveal quality]{
     \label{fig:teaser:quality}
-    \includegraphics[width=0.25\linewidth]{example-image-a}
+    \includegraphics[height=4.5cm]{TOG/figs/teaser_1c.pdf}
   }%subfloat
@@
 As a zoom-in, \subref{fig:teaser:quality} compares our foveal quality preservation.
-\qisun{Cross out half-half in  \subref{fig:teaser:hmd}. TODO for myself: I should make the background scene semi-transparent. Will do so soon.}
+\qisun{Cross out half-half in  \subref{fig:teaser:latency}. TODO for myself: I should make the background scene semi-transparent. Will do so soon.}
  }

### 2021-01-21 TOG/2-prior.tex
add two figures

@@
 To address this problem, recent years have witnessed extensive research on synthesizing novel views instead of storing.
-Examples include synthesizing light fields \cite{LearningViewSynthesis}, panoramas \cite{Lin:DeepPanorama}, and volumetric videos \cite{Broxton:immersiveLF}.
+Examples include synthesizing light fields \cite{LearningViewSynthesis}, panoramas \cite{Lin:DeepPanorama,Benjamin:2020:RTV}, and volumetric videos \cite{Broxton:immersiveLF}.
 However, despite being more compact, the synthesis is usually in a locally interpolated fashion than being globally consistent.

### 2021-01-21 TOG/4-method.tex
add two figures

@@
 In this section, we discuss our approach to achieve gaze-contingent neural rendering for 6DoF immersive virtual reality.
-As illustrated in \Cref{fig:pipeline}, our method predominantly comprises of two steps: 1) visual eccentricity aware frame synthesis, discussed in \Cref{sec:method:net}, and real-time rendering with optimized spatial-temporal precision, discussed in \Cref{sec:method:blending}.
+As illustrated in \Cref{fig:system}, our method predominantly comprises of two steps: 1) visual eccentricity aware frame synthesis, discussed in \Cref{sec:method:net}, and real-time rendering with optimized spatial-temporal precision, discussed in \Cref{sec:method:blending}.
 
@@
 
-    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/fovea.png}}\vspace{1em}
-    \subfloat[mid-pheripery]{\includegraphics[width=0.3\linewidth]{TOG/figs/mid.png}}\vspace{1em}
-    \subfloat[far-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/periph.png}}
+    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_fovea.png}}\vspace{1em}
+    \subfloat[mid-pheripery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_mid.png}}\vspace{1em}
+    \subfloat[far-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_periph.png}}
 
-    \subfloat[blend]{\includegraphics[width=0.9\linewidth]{TOG/figs/blended_with_mask.png}}
+    \subfloat[blend]{\includegraphics[width=0.9\linewidth]{TOG/figs/layer_blend/lobby_blended_with_mask.png}}
 
@@
 We overcome this problem by depicting the rapidly varying 360-degree virtual scenes with concentric spherical volumes. 
-This representation not only enables robust rendering of complex immersive environments at real-time rates, but also allows for 6DoF interaction for exploring virtual spaces, as will be discussed later.
+Inspired by \cite{Lin:DeepPanorama,Benjamin:2020:RTV}, the representation not only enables robust rendering of complex immersive environments at real-time rates, but also allows for 6DoF interaction for exploring virtual spaces, as will be discussed later.
 
@@
     \centering
-    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
+    \subfloat[quality-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/quality_only.png}}\hspace{1em}
+    \subfloat[latency-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/latency_only.png}}\hspace{1em}
+    \subfloat[ours]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/our.png}}
     \caption{Latency-quality joint-optimization.}
@@
     \centering
-    \subfloat[blending]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
+    \subfloat[blending]{\includegraphics[width = 0.8\linewidth]{TOG/figs/system/blended.png}}
 
-    \subfloat[Contrast enhancement]{\includegraphics[width = 0.8\linewidth]{example-image-a}}
-    \caption{Our multi-layer real-time rendering system.}
+    \subfloat[Contrast enhancement]{\includegraphics[width = 0.8\linewidth]{TOG/figs/system/blended_enhanced.png}}
+    \caption{Our multi-layer real-time rendering system.\zh{Or three-image pipeline}}
     \label{fig:method:blending}

### 2021-01-21 TOG/5-result.tex
add two figures

--- 2021-01-12 TOG/5-result.tex
@@
 Ablation studies for the network training/losses
-Target scene | uniformly downsampled image | local foveated rendered image | network predicted images ---> at least 4 different scenes.
+Target scene | Foveated target scene | NeRF | network predicted images ---> at least 4 different scenes.
 streaming time for 4-5 different scenes: full resolution vs our network prediction for “same data size”
@@
 }%note
+With a variety of scenes as shown in \Cref{fig:results:comparison2}, we evaluate our method with subjective studies (\Cref{sec:study:user}), objective analysis (\Cref{sec:study:quality}), and intra-system efficiency (\Cref{sec:study:intra}).
+\input{TOG/fig_tile_results}
+
 \subsection{User Study}
 \label{sec:study:user}
-\zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
+% \zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
 
 % In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
-In this section, we report a perceptual experiment to investigate how users perceive our solution compared to the ground truth and state-of-the-art (we choose to apply NeRF on our dataset).
+We first conducted a perceptual experiment to investigate how users perceive our solution ({\bf OURS}) compared to existing solution (\cite{mildenhall2020nerf}, {\bf NeRF}) and original mesh with foveated rendering (\cite{perry2002gaze}, {\bf F-GT}).
 
@@
 \paragraph{Stimuli}
-We designed four conditions for the perceptual experiment (see \autoref{fig:results:comparison}).
-The first column is the target scene, which we treat as the ground truth.
-The size of the image per eye is 1440 x 1600.
-The second column is a uniformly down-sample image at size \warning{???
+
+The stimuli were the images rendered via {\bf F-GT} or synthesized via  {\bf OURS}/{\bf NeRF}, as shown in \Cref{fig:results:comparison}. The resolution of the image per eye was $1440 \times 1600$. 
+For fair comparison, each group of stimuli were generated with the same, randomly defined gaze p-osition. The position was indicated as a green cross on the stimuli images.
+The resolution of the images per eye was $1440 \times 1600$. 
+We used the {\it gas} and {\it minecraft} scenes.
+Each condition from individual scene consists of $1$ view and $3$ different gaze positions.
+
+Specifically, we controlled the foveal kernel size in {\bf F-GT} to match our display size and eye-display distances.
+For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $N_{sample}=16$, $N_{layers}=4$, $N_{channel}=128$.
+The last column of \Cref{fig:results:comparison} shows the result.
+%We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.
+
+%\zh{We probably don't need to attach foveated GT here}
+% The second column is a uniformly down-sample image at size \warning{???
 x ???}.
@@
 So we down sample the entire image with ratio \warning{???}.
-% The third column enables foveated rendering.
-The foveated area is preserved as the original resolution while the rest of the image is down sampled.
-We apply the same alpha blending method that we used in our solution (see \autoref{sec:method:blending}).
+%The second column enables foveated rendering. We applied \warning{foveated algorithms} described in \cite{perry2002gaze} and \cite{jiang2015salicon}. 
+% and The foveated area is preserved as the original resolution while the rest of the image is down sampled.
+same alpha blending method that we used in our solution (see \autoref{sec:method:blending}).
 Similarly, we calculate the size of the foveated image based on the bandwidth.
 Hence the down-sample ratio of peripheral area is \warning{???}.
-The fourth column is generated by NeRF implementation.
-We assign $N_{sample}=16$.
+%The third column is generated by NeRF implementation.
+We retrained a new model on our dataset with parameters $N_{sample}=16$, $N_{layers}=4$, $N_{channel}=128$.
 The last column is our result.
-We predict the image from the camera transformation, fine-tune the foveated layer with local guided view, and render the image by alpha blending the three layers.
-The duration of the stimuli was \warning{??} seconds for each condition.
-% \zh{unsure if we want to render real-time results according to users gaze movement, or show static images with predefined gaze positions for certain duration?}
+We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.
 
 \paragraph{Setup}
-??
-subjects (\warning{?
-females and ?
-males, $M=?$, $SD=?$}) participated in the study.
-They were not involved in the project previously.
-Each participant was asked to wear a HTC Vive Pro Eye headset during the experiment.
-The headset is connected to a Windows machine with \warning{Intel i9-7900X CPU @ 3.30 GHz}, 64GB RAM and one NVIDIA GTX 3080 graphics card.
-During the experiment, the participants wearing the headset explored our stimuli.
-We used the \textit{two-alternative forced choice} (2AFC) method for both of our studies.
+Each participant was asked to wear a eye-tracked HTC Vive Pro Eye headset during the experiment. During the experiment, the participants wearing the headset examining  the stimuli. 
+Fourteen users participated in the study and 13 of them managed to complete (4 females and 9 males, $M=23.08$, $SD=1.75$).
+One participant (age=23, female) did not complete the experiment due to gaze drifting (please refer to the task description below).
+None of the subjects were aware of the research, the experimental hypothesis, or the number of rendering methods. All participants had normal or corrected-to-normal vision.
 
-\paragraph{Task: Our method v.s. NeRF v.s. Ground Truth}
-We aim at collecting user's feedback in VR display.
-Each participant were provided with \warning{30?} trials where each trial consisted a foveated area indicator and a pair of two stimuli.
+\paragraph{Task}
+The task was a \textit{two-alternative-forced-choice} (2AFC).
+Each trial consists of a pair of stimuli from the three conditions ({\bf OURS} / {\bf NeRF} / {\bf F-GT}).
+Each stimulus appeared for $300$ms on the the display.
+A forced $1.5$sec break was introduced between conditions.
+During the study, the participants were instructed to fix their gazes on the green cross.
+To prevent fixation from shifting away, we tracked the users gaze during the whole experiment.
+Whenever the gaze is $1$deg away from the target, a trial was dropped immediately with black green informing the participant.
+After each trial, the participants were instructed to select which one of the two stimuli appeared with higher visual quality using keyboard.
+Before each experiment, a warm-up session with 1 trial was provided to have the participants familiarize the study procedure.
+The order of conditions cmong trials were randomized with counterbalancing.
+%To start with, each participant was informed that they were required to choose an alternative with higher quality after watching each pair of stimuli.
+We calibrated the device and performed a warm-up trial to help participants understand our system.
+Each participant were provided with 1 warm-up trial and 6 official trials (2 scenes x 1 view x 3 gaze positions) where each trial consisted all the combinations of the pairs of every two stimuli.
 We randomize the order of the pairs for counterbalancing.
-To control the images to be shown for each participant, we choose to render the results for all conditions on predefined gaze positions rather than passing user's real time gaze input to the system.
-We indicate the location of the foveated area to users in advance to all the conditions first.
-For each condition, participants were asked to report \zhtd{whether they think the rendered result is better}.
-
-% \zh{Since we will show at least four conditions to each participant, are we gonna ask them to pick up ONE as the favorite or ask them to judge each condition as ACCEPTED or NOT ACCEPTED?
-Or we should apply 3 trials that each contains ours and one of other alternatives?
-For example, Condition1 v.s.
-ConditionOurs, Condition2 v.s.
-ConditionOurs, and Condition3 v.s.
-ConditionOurs}
-To minimize the effect of accumulating discomfort during the experiment, we enforced at least a 60-second break between the two conditions in each trial, but the participants were instructed to take as much time as needed to recover.
-% \zh{I think rendering static images rather than taking real gaze input is more controllable and good for comparison, because we cannot control users gaze input, thus the images shown to participants are not the same.
-If we wanna follow a within subject study design, we probably need to pre-generated the gaze path.
-HOWEVER, if it does not matter if the participants need to look at the same images, which means the study is not ``paired'', we can still take real gaze input.
-But I think it increases the pressure for implementation, and we are not designing a gaze tracking system actually.
-How do you think about it?}
-
-\qisun{Here is what I think.
-The conditions, OURS, UNIFORM (i.e.
-downsampled), NERF could be described in the stimuli part.
-Then in the task paragraph, we simply compare any given pair of these 3 conditions.}
-% \zh{no more down samples?}
-\zh{no GT included?}
+We showed a yellow cross on the screen.
+After participants move their gaze positions to the area close to the cross, the yellow cross turns green.
+The first stimulus will be rendered as well as the cross after participants fixate the cross for 1 second.
+Each stimulus is displayed for 300 milliseconds.
+The second one will be rendered after 1.5-second dark screen.
+If the participants does not fixate at the cross when stimuli are displayed, the current pair will be discarded and displayed to the participants later in the same trial.
+After each pair, participants reported their opinions on which image has higher quality.
+To minimize the effect of accumulating discomfort during the experiment, we enforced at least a 60-second break between each trial.
+Meanwhile, the participants were instructed to take as much time as needed to recover.
 
 \paragraph{Results}
+\Cref{fig:results:2afc} plots the study results. 
+Among all three conditions, we observed near to random guess among trails that compares  {\bf F-GT} and {\bf OURS} (50.0\% voted for {\bf OURS}).
+However, participants significantly prefers  {\bf F-GT} over {\bf NeRF} (98.7\% voted for NeRF, binomial test showed $p<0.005^{***}$).
+The preference applies to {\bf OURS} vs.
+{\bf F-GT} as well (97.4\% voted for NeRF, binomial test showed $p<0.005^{***}$).
+
+% our/gt 78/156 0.5318897993488445
+% our/nerf 152/156 2.6681140960754964e-40
+% nerf/gt 2/156 1.3407579916082842e-43
+
 \begin{figure}[htb]
     \centering
-    \includegraphics[width=0.96\linewidth]{example-image-a}
-    \caption{The users' preference votes from our evaluation experiment.}
+    \includegraphics[width=0.96\linewidth]{TOG/figs/vote.pdf}
+    \caption{The users' preference votes from our evaluation experiment. X axis shows all the pair combination and Y axis shows the ratio of voting for the second one.}
     \label{fig:results:2afc}
 \end{figure}
-The user preference votes are visualized in \autoref{fig:results:2afc}.
-% apply binomial test
+
+% apply binomial test https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.binom_test.html
 % or apply ANOVA, there is a value of votes for each condition and we want to know if there exists significant effects on different condition.
+\paragraph{Discussion}
+The near-to-random-guess indicated the statistically similar perceptual quality between {\bf OURS} and foveated ground truth ({\bf F-GT}). 
+Given the perceptual identicality between foveated and full-resolution rendering, the subjective study indicates that {\bf OURS} can achieve similar quality as a rendering with a locally fully stored high-quality mesh.
+Meanwhile, both of the two conditions showed significant subjective quality preference than {\bf NeRF}.
+That is, under an immersive, high FoV, and first-person view, the current volume-based scene representation may not fully reproduce the perceptually identical retinal images.
+To further validate the perceptual quality by breakdown across the whole visual field, we conducted measurements with objective metrics in the following section.
+
+\subsection{Visual Quality}
+\label{sec:study:quality}
+Complementary to the subjective measurement (\Cref{sec:study:user}), we evaluate the perceived quality in an objective fashion.
+Specifically, under four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT} and {\bf NeRF} at individual eccentricity ranges across the whole visual field (up to $110$ deg, the capability of the VR HMD).
+%\zh{gonna comment this: To validate the non-uniform perceptual quality on the retina, we apply a foveated filtering to {\bf GT} \cite{perry2002gaze,jiang2015salicon}.}
+For each eccentricity range, we compare the deep perceptual similarity (LPIPS) \cite{zhang2018unreasonable} across all scenes.
+LPIPS uses deep neural networks to evaluate perceptual similarity between the image provided and a reference image.
+Smaller values indicate higher perceptual similarity.
+For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $420$ data per scene.
+We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity value and together.
+Paired t-tests with Holm correction were used for all pairwise comparisons between stimuli.
+All tests for significance were made at the $\alpha=0.05$ level.
+% The error bars in the graphs show the 95\% confidence intervals of the means.
+
+\paragraph{Results} 
+\Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg per sample). 
+From foveal to near periphery (<=$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.84$).
+That is, \textbf{OUR} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), while significantly higher LPIPS than \textbf{F-GT} ($p<.001^{***}$).
+For example, the main effects of stimuli ($F(2,38)=42.023, p=.001^{***}$) was significant on eccentricity $=25$  deg  in scene \textit{gas} (\Cref{fig:lpips:gas}).
+\textbf{OUR} was significantly lower than \textbf{NeRF} ($t(19)=-18.557, p<.001^{***}$) and \textbf{F-GT} ($t(19)=-10.342, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
+
+From mid- to far- periphery (>$40$ deg ), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.19$).
+\textbf{OUR} shows significantly higher LPIPS than \textbf{NeRF} due to the foveated nature of our design.
+Whereas, comparing with \textbf{F-GT}, we observed  significant lower scores ($p<.001^{***}$).
+For instance, the main effects of stimuli was significant on eccentricity $=60$ deg in scene \textit{gas} ($F(2,38)=493.51, p<.001^{***}$, \Cref{fig:lpips:gas}).
+\textbf{OUR} was significantly lower than \textbf{F-GT} ($t(19)=-12.28, p<.001^{***}$), and higher than \textbf{NeRF} ($t(19)=18.90, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
+
+The general trend applies to all 4 scenes being validated.
 
 \paragraph{Discussion}
+The observation revealed our method's over-performance than alternative view synthesis solutions in the foveal and near peripheral vision.
+This is evidenced by the significantly increased perceptual similarity to \textbf{GT} while comparing between \textbf{OUR} and \textbf{NeRF}.\zh{``significantly increased LPIPS'' requires analysis}
 
-\subsection{Ablation Study}
+Meanwhile, \textbf{OUR}'s increased perceptual similarity comparing with \textbf{F-GT} beyond $40$ deg indicated their similar visual quality.
+This shows that \textbf{OUR} doesn't compromise the peripheral vision's quality with its significantly enhanced foveal image.
+
+The discoveries also agrees with the trend from \Cref{sec:study:user}.
+That is, in addition to the faster performance, our method showed remarkably perceptual quality gain than \textbf{NeRF} for first-person 6DoF immersive viewing of 3D scenes: our synthesized views are perceptually identical to the traditionally rendered images from complete mesh.
+It yet achieves significant data storage saving (\warning{xxx\%-xxx\%}).
+
+% ZH: why \Cref{fig:lpips:gas} shows 'section 5.2'
+\begin{figure*}
+    \centering
+    \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenegas.pdf}}\label{fig:lpips:gas}
+    \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}
+
+    \subfloat[gallery]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenegallery.pdf}}
+    \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
+    \caption{LPIPS analysis of all scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity.}
+    {\zh{Add intersection lines later.}}
+    \label{fig:lpips}
+\end{figure*}
+
+\subsection{Intra-System Workings}
+\label{sec:study:intra}
+\paragraph{Spatial-Temporal Optimality}
+
+% apply objective metric to |cref{fig:optimization}, aka different sample numbers
+
+\paragraph{Ablation Study}
 % the performance of an AI system by removing certain components, to understand the contribution of the component to the overall system
 
-\subsection{Visual Quality}
-As a complementary to the subjective measurement \Cref{sec:study:user}, we evaluate the perceived quality in an objective fashion.
-Specifically, we compare our method with both full resolution rendering and \cite{mildenhall2020nerf} with the same time consumption.
-
-\begin{figure*}[htb]
-    \centering
-    \begin{minipage}{0.32\linewidth} %full res
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_gt1.png}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_gt1.png}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
-    \end{minipage}
-    \begin{minipage}{0.32\linewidth} %ours
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/gas_fovea_result1.png}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{TOG/figs/mc_fovea_result1.png}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
-    \end{minipage}
-    \begin{minipage}{0.32\linewidth} % NERF
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
-
-        \subfloat[]{\includegraphics[width=0.96\linewidth]{example-image-a}}
-    \end{minipage}
-
-    \caption{Comparing our synthesis method with full resolution rendering and alternative solutions.}
-    \label{fig:results:comparison}
-\end{figure*}
-
-\subsection{Spatial-Temporal Optimality}
-%\zh{about streaming? network?}
+% nmsl v.s. msl

### 2021-01-21 TOG/fig_lpips.tex
add two figures

\begin{figure*}
    \centering
    \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips-fov-gas-group.pdf}}\label{lpips:gas}
    \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}

    \subfloat[studyroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenestudyroom.pdf}}
    \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
    \caption{LPIPS analysis of all scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity. $^*$: $p<.05$, $^{**}$: $p<.01$.}
    {We found a significant difference in \warning{ecc range} between \warning{stimulus} and \warning{stimulus};}
    \label{fig:lpips}
\end{figure*}

### 2021-01-28 TOG/0-abstract.tex
last day

@@
 
-Traditional high-quality 3D graphics requires large volume of detailed data to be rendering.
-This compromises efficiency and local storage.
-In recent years, cloud-based streaming and neural representations have significantly revolutionized the pipeline by redistributing and compressing the computation/storage.
-However, they typically suffer from high latency and low quality for practically large virtual scenes with 6 degree-of-freedom camera motions.
-These problems are particularly exacerbated in immersive medias such as virtual and augmented reality due to the demand of high frequency and resolution.
+Traditional high-quality 3D graphics requires large volumes of fine-detailed scene data for rendering.
+This demand compromises computational efficiency and local storage resources.
+Specifically, it becomes more concerning for future wearable and portable virtual and augmented reality (VR/AR) displays.
+Recent approaches to combat this problem include remote rendering/streaming and neural representations of 3D assets.
+These approaches have redefined the traditional local storage-rendering pipeline by distributed computing or compression of large data.
+However, these methods typically suffer from high latency or low quality for practical visualization of large immersive virtual scenes, notably with extra high resolution and refresh rate requirements for VR applications such as gaming and design.
 
-Tailored for the future cloud-based, portable, and low-storage VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis approach.
-Our method mutually bridges both visual acuity and neural representation for 6DoF immersive interaction.
-Moreoever, it optimizes between temporal computational latency and spatial predictive precision.
-Both objective analysis and subjective studies evidenced our effectiveness of significantly reducing local storage (\warning{xxx MB, xx\% size reduction} for large 3D scenes) while preserves original perceptual quality without introducing additional latency.
+Tailored for the future portable, low-storage, and energy-efficient VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis method.
+We incorporate the human psychophysics of visual- and stereo-acuity into egocentric neural representation of 3D scenery.
+Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high quality immersive interaction.
+Both objective analysis and subjective study demonstrate the effectiveness of our approach in significantly reducing local storage volume and synthesis latency (up to 99\% reduction in both data size and computational time), while simultaneously presenting high-fidelity rendering, with perceptual quality identical to that of fully locally stored and rendered high-quality imagery.
 \end{abstract}

### 2021-01-28 TOG/1-introduction.tex
last day

--- 2021-01-05 TOG/1-introduction.tex
@@
 }%note
+Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
+Thus, streaming services and remote cloud-based rendering (such as with Google Stadia and Nvidia Shield) have rapidly emerged in recent years.
+These storage-friendly systems have shown an unprecedented potential to reshape the traditional computer graphics pipeline by completely relieving the end user's computational workload and storage pressure, thus advancing device portability.
+Future spatial and immersive computing platforms such as VR and AR are potentially major beneficiaries due to the extensive demands of portability and energy-efficiency  \footnote{\url{https://www.vxchnge.com/blog/virtual-reality-data-storage-concern}}.
 
-Started from remote desktops and online videos, cloud-based computing has show unprecedented potential to reshape computer graphics pipeline by completely relieving end-user's computational workload thus device bulkiness.
-In recent years, we have observed growing interactive streaming services such as Google Stadia or NVIDIA Shield.
-However, these off-the-shelf services have only been deployed in less latency sensitive platforms such as TV-based game consoles, due to the inevitable latency (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}.
-Immersive medias, such as virtual and augmented realities, demand remarkably lower latency and higher fidelity/speed for sickness-free and natural experience.
-Due to the consequently massive spatial-temporal data volume, there is no solution for streaming frames from server to individual end users today.
+However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles, due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}).
+In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for sickness-free and natural experience.
+As a consequence, there is a demand for massive high-quality spatial-temporal data.
+Unfortunately, no solution exists today to deliver high-quality and dynamic/interactive immersive content at low latency and with low local storage.
 
@@
 (Note: this is just high level big ideas. Details should go to a previous work section.)
-}%note
+}%notez
+%\qisun{(Jan 5, 2021) We should include two main fields: neural representation and foveated rendering}
 
+Recent years have witnessed a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets.
+Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf} and many more.
+However, existing solutions suffer from either high time consumption or low image fidelity with large scenes.
+
+Acceleration without compromising perceptual quality is also an ultimate goal of real-time rendering.
+Gaze-contingent rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}.
+This is achieved by harnessing high field-of-view(FoV) head-mounted displays and high precision eye-tracking technologies.
+However, existing foveated rendering approaches still require locally stored full 3D assets, resulting in long transmission time and storage volume, whose availability is unfortunately limited in wearable devices.
+
+\qisun{(Jan 22, 2021) Do we want to discuss 6DoF video synthesis here?}
 \note{
@@
 
-This document provides a template for writing research papers, and describes how to manage project progress via iterating drafts.
-Doing this correctly can help you achieve productive research and happy life, e.g. working anywhere anytime without synchronous meetings with your collaborators.
+To overcome these open challenges mentioned above, we present the first gaze-contingent and interactive neural synthesis for high-quality 3D scenes.
+Aiming at the next-generation wearable, low-latency and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions.
+The predicted frames are perceptually identical to traditional real-time rendering of high-quality scenes.
+With wide FoV and high resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB to 0.5MB) thus enables instant viewing (without long loading/transmission waits) compared to traditional rendering pipelines; preserves visual quality, and improves responsiveness (from 9s to as low as 20 ms) compared with alternative view neural synthesis methods.
+Tailoring to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.
+%\praneeth{---- somoene please check the text in the brackets and make sure it is ok :)}
 
@@
 
-I learned from my PhD adviser to start writing from the moment I have the faintest idea, and gradually update the drafts to reflect progress.
-During my first job with NVIDIA, I learned why and how revision control is essential for sharing and editing source files, especially among large engineering teams.
-The iterative nature of writing and coding matches well with the capability of revision control.
-Thus I prefer Latex + git/svn (e.g. github, bitbucket, and the svn server under my hosting service), even though I can still manage with MS Word + cloud drive (e.g. Google drive, Box, and Dropbox). 
-A good computer scientist writes papers like programs, and manages Latex files like source codes.
-These repos are external memories and communication mediums for the collective brains of my teams.
+To achieve immersive viewing with low storage, high visual quality as well as low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis.
+%To this end, we build both the retina-matched neural models and VR-tailored representation coordinates, towards high visual quality as well as low systematic latency.
+\praneeth{-- this line is a little confusing to me}.
+Specifically, we represent 3D scenes with concentric spherical coordinates.
+Our representation also allows for depicting the foveated color sensitivity and stereopsis, during both the training and inference phases.
+Furthermore, we derive an analytical spatial-temporal perception model to optimize our neural scene representation towards imperceptible loss in image quality and latency.\qisun{better?
+This paragraph was a bit messy before.
+Tried to clean up.}
 
@@
 
-\input{example_fig}
+% In 3D scenes represented by concentric spherical coordinates, we model each camera ray's perceivablity on the retina.
+It particularly depicts the foveated color sensitivity and stereopsis during both training and inference.
+% By deriving an analytical spatial-temporal model, we further optimize our representation and networks towards imperceptible quality/precision loss and speed/latency.
 
-Translating thoughts into words, diagrams, equations, pseudo-codes, or codes is a good way to clarify and consolidate.
-If you are writing a graphics/HCI paper, design the figures so that they alone can provide a high level picture of the main points.
-As exemplified in \Cref{fig:example}, a figure can contain images, drawings, or combinations.
+%Specifically, we represent an arbitrary virtual 3D scene as a compact neural representation in a viewer centered inside-out fashion. 
+%We account for the challenges such as low quality/resolution/performance common to existing 3D deep neural representation approaches, we further leverage the human visual perception to synthesize gaze-contingent and vision-matched immersive frames, without compromising the speed and image fidelity.
+%The method is further optimized with in the spatio-temporal domain to balance the latency and precision.
 
@@
 
-You can use the associated files as a starting template for your research papers.
-Start with the \filename{Makefile}, and notice the two build targets: final for official submission and public disclosure, and draft for internal sharing among your collaborators (including yourself).
+We validate our approach by conducting numerical analysis and user evaluations on commercially available AR/VR display devices.
+The series of experiments reveal our method's effectiveness in delivering an interactive VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, with significantly lowered data storage.
+We make the following major contributions:
+\begin{itemize}
+    \item A low-latency, interactive, and significantly low-storage immersive application. It also offers full perceptual quality with instant high resolution and high FoV first-person VR viewing.
+    \item A scene parameterization and vision-matched neural representation/synthesis method considering both visual- and stereo- acuity.
+    \item A spatiotemporally analytical model for jointly optimizing systematic latency and perceptual quality.
+\end{itemize}

### 2021-01-28 TOG/2-prior.tex
last day

--- 2021-01-12 TOG/2-prior.tex
@@
 
-\subsection{Deep View Synthesis}
+\subsection{Image-based View Synthesis}
+Image-based rendering (IBR) has been proposed in computer graphics to complement the traditional 3D assets and ray propagation pipelines \cite{levoy1996light,gortler1996lumigraph,shum2000review}.
+Benefiting from the uniform 2D representation, it delivers pixel-wise high quality without compromising performance.
+However, one major limitation of IBR is the demand for large and dense data acquisition, resulting in limited and sparse view ranges.
+To address this problem, recent years have witnessed extensive research on synthesizing novel image-based views instead of fully capturing or storing them.
+Examples include synthesizing light fields \cite{LearningViewSynthesis,Li2020LF,mildenhall2019llff} and multi-layer volumetric videos \cite{Broxton:immersiveLF}.
+However, despite being more compact, the synthesis is usually in a locally interpolated fashion than globally applicable with highly dynamic VR camera motions.
+
+\subsection{Implicit Scene Representation for Neural Rendering}
+To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attentions in both computer graphics and computer vision communities. 
+With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf}.
+Prior arts also investigated on implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects.
+Inspired by \cite{park2019deepsdf}, signed distance function has been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}.
+The input information was varied from 2D images ~\cite{lin2020sdf, yariv2020multiview}, 3D context ~\cite{saito2019pifu, oechsle2019texture}, time-varying vector field~\cite{niemeyer2019occupancy}, and local features~\cite{tretschk2020patchnets, liu2020neural}.
+Current implicit representations primarily focus on locally ``out-side-in'' viewing of individual objects, with low field-of-view, speed and resolution.
+For large scenes, the lack of coverage may cause quality drop.
+However, in virtual reality, the cameras are typically first-person and highly dynamic, requiring low latency, high resolution, and 6DoF coverage. 
+%Our method, in comparison, is VR tailored via representing the scene with user-centered views.
 
 \subsection{Gaze-Contingent Rendering}
-\cite{Guenter:2012:F3G,Patney:2016:TFR,Sun:2017:PGF,Kaplanyan:2019:DNR}
+Both rendering and neural synthesis suffer from the heavy computational load, thus systematic latency and lags.
+With a full mesh representation, foveated rendering has been proposed to accelerate local rendering performance, including mesh- \cite{Guenter:2012:F3G,Patney:2016:TFR} or image-based \cite{Sun:2017:PGF,Kaplanyan:2019:DNR} methods.
+The accelerations are typically achieved through high field-of-view displays (such as VR head-mounted displays) and eye-tracking technologies.
+However, these all require full access to the original 3D assets or per-frame images.
+Consequently, the systems usually require large local storage, long transmission waits, or constant frame-based transmissions with latency.
+%On the other hand, most view synthesizing methods are designed for offline viewing or processing, other than real-time or interactive consumption, especially with high resolution and FoV.  
+Our method bridges human vision to deep synthesis.
+It accelerates neural synthesis beyond local rendering, enabling both computational and storage efficiency.
+It is suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
+%\zh{seems not sufficient?}
+
+\subsection{Panorama-Based 6DoF Immersive Viewing}
+Panorama images and videos greatly fit VR applications thanks to their 360 FoV coverage.
+However, the main challenge is their single projection center in each frame, limiting free-form camera translation, thus 6DoF immersive viewing experience.
+Recently, extensive research has been proposed to address this problem.
+Serrano et al.
+\shortcite{Serrano_TVCG_VR-6dof} presented a depth-based dis-occlusion method that dynamically reprojects to 6DoF cameras, enabling natural viewing with motion parallax.
+Pozo et al.
+\shortcite{Pozo:2019:I6V} jointly optimizes capture and viewing processes.
+Machine learning approaches have advanced the robustness of various geometric and lighting conditions \cite{Attal:2020:ECCV,Lin:DeepPanorama,Benjamin:2020:RTV}. 
+As a spherical image, panoramas are designed for capturing physical worlds with flexible reprojections to displays.
+The 6DoF viewing typically suffer from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
+With an orthogonal mission, our model represents and reproduces arbitrary 3D virtual worlds for low storage VR.
+It lowers memory usage by replacing traditional mesh- or volume-based data than predicting image-space occlusions.

### 2021-01-28 TOG/4-method.tex
last day

--- 2021-01-12 TOG/4-method.tex
@@
 \label{sec:method}
-As visualized in \Cref{fig:pipeline}, our method consists of three major components: foveal-periphery image synthesis (\Cref{sec:method:net}), and real-time blending with optimized spatial-temporal precision (\Cref{sec:method:blending}).
+In this section, we present our perception-guided neural scene synthesis for virtual reality.
+Based on a concentric spherical representation (\Cref{sec:method:representation}), our system predominantly comprises of two main steps: visual- and stereo- acuity adaptive view synthesis (\Cref{sec:synthesis}), and image-based real-time rendering to composite stereo display frames (\Cref{sec:method:blending}).
+We then craft an analytical spatial-temporal model to guide the optimization of our intra-system variables for desired precision-performance balance (\Cref{sec:method:optimization}).
+\input{TOG/fig_notations}
+\subsection{Egocentric Neural Representation}
+\label{sec:method:representation}
+The recent single-object-oriented ``outside-in'' view synthesis methods \cite{sitzmann2019deepvoxels,mildenhall2020nerf} typically represent the training targets using uniform voxelization.
+However, immersive VR environments introduce open challenges for such parameterization due to the rapid variation of the egocentric (first-person) viewing (e.g., \Cref{fig:teaser:scene}).
+As a consequence, the neural representation on large virtual environments typically suffer from ghosting artifacts, low resolution, or slow speed (e.g., the right triangle in \Cref{fig:teaser:quality}).
+%We tailor our method specifically for egocentric (first-person) views in immersive VR environments.
+%rendering of immersive viewer-centered ``inside-out'' scenes is challenging owing to the rapid variation of scene content. 
+\paragraph{Egocentric coordinates}
+To tackle this problem we are inspired by the recent panorama dis-occlusion methods \cite{Lin:DeepPanorama,Benjamin:2020:RTV,Broxton:immersiveLF} to depict the rapidly varying first-person views with concentric spheres.
+This representation enables robust rendering at real-time rates and allows for 6DoF interaction to navigate complex immersive environments.
+As visualized in \Cref{fig:notations}, our representation is parameterized with the number of concentric spheres ($\sphereNum$) and their respective radii ($\mathbf{\sphereRadius}=\{\sphereRadius_i\},i\in [1,\sphereNum]$).
+Our network aims at predicting the function $\mlpFunc(\hat{\SpatialPt})\triangleq(r,g,b,d)$ of any position $\mathbf{\hat{\SpatialPt}}({\sphereRadius_{i,i\in[1,\sphereNum]},\theta,\phi})$ on these $\sphereNum$ spherical surfaces.
+%\dnc{$\mathbf{q}$ may be more proper}\qisun{Just to avoid confusion with \Cref{fig:notations}.
+What do you think?}\dnc{Then perhaps we can use $\hat{q}$ as shown in \Cref{fig:notations} I think the alphabet 'q' means it's in spherical system, 'p' never exists anywhere else}
+Here, $(r,g,b)$ and $d$ are the predicted color and density, respectively. 
+We then employ a ray marching through this intermediate function to produce images of specified views.
 
-%\subsection{Pipeline}
+\paragraph{Training and synthesis}
+%\dnc{Perhaps we can merge this paragraph with above section? And move the last paragraph of above section to here as a intro. Then this section is more focusing on GAZE-AWARE}
+For each scene, we approximate the function $\mlpFunc$ by training a multilayer perceptron (MLP) network, similar to \cite{mildenhall2020nerf}.
+The network is composed of $\mlpLayerNum$ fully-connected layers of $\mlpChannelNum$ channels, each using ReLu as the activation function.
+We will further determine $\mlpLayerNum$, $\mlpChannelNum$ and $\sphereNum$ via a spatial-temporal optimization in \Cref{sec:method:optimization}.
+Once trained, the network achieves 4D prediction on the spherical voxels, on which a ray marching rendering is performed to synthesize the final frames for display.% based on users' frame-wise views. 
+Similar to \cite{mildenhall2020nerf}, we apply a trigonometric input encoding and compute the L2-distance between the volume-rendered and mesh-rendered images as the training loss function.
+We discuss the specifics of training data creation and implementations in \Cref{sec:impl}.
+
+\subsection{Gaze-Aware Acceleration}
+\label{sec:synthesis}
+\qisun{@nianchen: does this section title sounds reasonable to you?}
+%The scene representation above is created to parameterize a given 3D space for neural synthesis. 
+%the foveated rendering manner of VR display through training two orthogonal neural networks that are tailored for the perceptual acuity variances.
+Our concentric spherical coordinate system, as described in \Cref{sec:method:representation}, addresses the view variance problem in VR.
+However, it sill suffers from significant rendering latency (seconds per frame).
+This is one of the core challenges causing neural representation to be unsuitable yet for VR.
+To combat the common performance problem of view synthesis, we predict multiple elemental images based on the viewer's real-time head and gaze motions\nothing{ instead of synthesizing a globally uniform image}.
+The elemental images are tailored for the visual acuity variances.
+We detail our network design and elemental image synthesis in the following section.
+We enable real-time responsiveness via introducing gaze-adaptive synthesis mechanisms.
+\input{fig_system}
+\paragraph{Adaptive visual acuity} 
+%the user's current view $[\rayo,\camDir]$ and gaze position $\gazeDir$ that are tracked in real-time by a VR headset.
+Existing neural scene representation networks predict uniform resolution images for offline third-person viewing.
+Due to the trade-off between quality and latency during inference, they are typically executed with low quality and/or high latency when it comes large field-of-view and high resolution imagery.
+\praneeth{this line is confusing}
+Our method, without compromising performance, generates {\it perceptually} high quality frames matching the retinal acuity, i.e. how the retina perceives dynamic and stereo frames.
+% Note that foveal/peripheral retinal images requires high/low pixel-wise precision but low/high FoV (number of pixels).
+Thus, we train two orthogonal neural networks that are tailored for the perceptual acuity variances.
+Because of foveation of imagery by the human visual system \cite{Guenter:2012:F3G}, we train two orthogonal networks (i.e., $\mlpFunc$s) to synthesize foveal ($\imageFoveal(\rayo,\camDir,\gazeDir)$, $0-20$deg), and peripheral elemental images (i.e., two $\mlpFunc$s to deploy for ray marching) for each eye.
+The periphery includes mid- ($\imageMid(\rayo,\camDir,\gazeDir)$, $0-45$deg) and high- eccentricity visual fields ($\imageFar(\rayo,\camDir$), $0-110$deg), see \Cref{fig:system}.
+Note that $\imageFar$ is independent from the gaze direction $\gazeDir$. 
+%The images are defined as functions that maps individual camera rays ($\rayo,\rayd$) to an RGB color.
+\dnc{Could move to section 4?
+I think it's implementation related}\qisun{I think the PDD numbers give readers a rough impression of the varied quality across each of the image, right?} Specifically, to incorporate the display hardware properties and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $128^2$/$256^2$/$230\times256$ resolutions respectively.
+That is, the $\imageFoveal$ has the highest spatial resolution of $6.4$ pixels per degree (PDD), higher than those of $\imageMid$ ($5.7$ PDD) and $\imageFar$ ($2.33$ PDD).
+%In each frame, the two networks independently return three elemental images, as seen from the separation range in deg, gradually larger areas along eccentricity.
+An example is shown in \Cref{fig:system}.
+
+\paragraph{Adaptive stereo-acuity}
+{
+To be rendered on head-mounted displays, stereo elemental images $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$ ($l$ and $r$ represent left and right eyes respectively) are required.
+This doubles the computational cost that is critical for latency- and frame-rate- sensitive VR experience
+\nothing{The separation of the three elemental images considers the varied visual acuity.
+However, in a head-mounted stereo VR displays, the rendered image are for two eyes, resulting in double computation for $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$.
+Here $l$ and $r$ represent the projection images for left ($\rayo^l$) and right ($\rayo^r$) eyes respectively.
+Whereas, because $\imageMid^{\{l,r\}}$ and $\imageFar^{\{l,r\}}$ demand high spatial resolution due to their large eccentricity coverage, }
+}
+(please refer to \Cref{sec:study:intra} for detailed comparisons).
+
+To tackle this problem towards real-time experience, we accommodate the inference process with the adaptive stereo-acuity in the periphery.
+Specifically, besides the visual acuity, psychophysical studies have revealed human's significantly declined stereopsis while receding from the gaze point \cite{mochizuki2012magnitude}.
+Inspired this characteristic, instead of inferring $6$ elemental images, we perform the computation with $\imageFoveal^{\{l,r\}}$, $\imageMid^m$, and $\imageFar^m$, where $m$ indicates the view at the midpoint of the left and right eyes.
+The individual images are further blended for display, as detailed in \Cref{sec:method:blending}.
+\Cref{fig:mono} visualizes the stereo changes from the adaptation.
+\input{TOG/fig_mono_comparison}
+
+\subsection{Real-Time Frame Composition}\dnc{May be moved to section 4 (merge with the paragraph 'Integration').
+I think it's not closely related to our methodology and we have no key contributions about that.
+It's more like an engineering issue}\qisun{This whole section?
+I was concerned that would break the method completeness.
+Intentionally kept it short according to contribution level here.}
+\label{sec:method:blending}
+In real-time, we perform a image-based rendering that blends the elemental images. The output frames are rendered on the stereo VR displays.
+To accommodate the uniformized $\imageMid^m$, we shift it with $-/+ \frac{\rayo^r-\rayo^l}{2}$ for left/right eye respectively. 
+\dnc{Not accurate. shifting is related to mean of fovea depthmap inferred in actual implementation to minimize the misalignment between fovea and mid}
+{
+With the obtained elemental images as input, a fragment shader is executed to generate final images for each eye.
+Two adjunct layers are blended using a smoothstep function across $40\%$ of inner layer.
+\nothing{After obtaining the individual images, our fragment shader retrieves the three frames and applied a pixel wise bilinear interpolation at each $10\%$ connecting areas between two adjunct layers.}
+}
+This enhances visual consistency on the edges between layers \cite{Guenter:2012:F3G}.
+Moreover, to further preserve peripheral elemental images' visual fidelity due to its low PDD, we enhance the contrast following the mechanism of \cite{Patney:2016:TFR}, the details are visualized in \Cref{fig:method:blending}.
+%\dnc{Note: the contrast enhancement is performed on elemental images, before blending.
+Different enhancement parameters are applied for fovea, mid and periph layers}
+
 \begin{figure*}[htb]
     \centering
-    \includegraphics[width=\linewidth]{TOG/figs/sys_pipeline.png}
-    \Caption{System pipeline.}
+    \subfloat[network overlay]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/overlap2.pdf}\label{fig:method:blending:overlay}}\hspace{1em}
+    \subfloat[blending]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended2.pdf}\label{fig:method:blending:blending}}\hspace{1em}
+    \subfloat[contrast enhance]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
+    \Caption{Our multi-layer real-time rendering system.}
     {%
+    \subref{fig:method:blending:overlay} shows the original 3-layer image output overlaid.
+    \subref{fig:method:blending:blending} shows the result with cross-layer blending that enables smooth appearance on the edges.
+    \subref{fig:method:blending:contrast} shows the contrast-enhanced rendering pass in the periphery.
     }
-    \label{fig:pipeline}
+    \label{fig:method:blending}
 \end{figure*}
 
-%To meet the requirement of real-time, use fewer channels/layers, much fewer samples, predict a coarse fovea image, then refine using local view guidance
+\subsection{Latency-Quality Joint Optimization}
+\label{sec:method:optimization}
+As a view synthesis system based on sparse egocentric representation (the density of $\sphereNum$ spheres) and neural networks (the $\mlpLayerNum\times\mlpChannelNum$), the method inevitably introduces approximation errors.
+On the other hand, these variables also determine the online computational time that involves inferring function $\mlpFunc$ and ray marching the $\sphereNum$ spheres.
+However, VR strictly demands both quality and performance. We present a spatial-temporal model that analytically depicts the correlations and optimizes the variables for ideal user experience.
 
-\subsection{GazeNet}
-\label{sec:method:net}
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\linewidth]{TOG/figs/spher_net_flow.png}
-    \caption{GazeNet Network.}
-    \label{fig:my_label}
-\end{figure*}
-
-%Follow NeRF's idea, represent scene using a MLP but in spherical space because we are inside-out (viewer is located at center area and look around)
-
-\qisun{(Jan 9, 2021) Avoid usage of subsubsection.}
-\paragraph{Multi spherical representation}
-As reviewed in \autoref{sec:prior}, prior work contributed a lot to represent a static scene via object-focused manner.
-Inspired by that, we investigate on how to represent a static scene for VR environment to render real-timely without loading the entire scene.
-However, VR environment is user-centered.
-That is, unlike the object-focused ``outside-in'' way such as \cite{mildenhall2020nerf}, the views are highly varied in an ``inside-out'' fashion.
-Consequently, the overlaps among views are diversified, resulting in training challenges for large virtual scenes.
-To address this problem, we depict a VR scene as concentric spherical volumes with the center as the original point, as visualized in \Cref{fig:teaser:scene}.
-
-%\begin{figure}[htb]
-%    \centering
-%    \includegraphics[width=0.96\linewidth]{TOG/figs/spher_coord.pdf}
-%%    \caption{Visualization of our multi-spherical coordinate system.}
-    \label{fig:method:coordinate}
-%\end{figure}
-% different position (10cm working, 50cm not working),
-% different rotation
-
-\dnc{(Jan 11) Ray to spherical coord} For a sphere with radius $r$, the intersection of ray and sphere $\pt_r$ can be calculated by solving:
-
-\begin{equation}
-    \|\pt_r\|^2=\|\rayo + k\rayd\|^2=r^2
+\paragraph{Precision loss of a 3D scene}
+As shown in \Cref{fig:notations}, under the egocentric representation, a 3D point $\pt$ is re-projected as the nearest point on a sphere that connects it to the origin point:
+\begin{equation}\label{eq:closestPoint3D}
+{\pt^\prime}(\sphereNum,\mathbf{\sphereRadius},\SpatialPt) \triangleq \sphereRadius_k \frac{\pt}{\norm{\pt}}, \ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
 \end{equation}
 
-Convert $\pt_r$ to spherical coordinate $(r, \theta, \phi)$:
+Similar to volume-based representation, the multi-spherical system is also defined in discrete domain.
+The sparsity thus naturally introduces approximation error that compromises the synthesis quality.
+To analytically model the precision loss, we investigate the geometric relationship among the camera, the scene, and the representation.
+As illustrated in \Cref{fig:teaser:scene,fig:notations},  for a sphere (located at origin point) with radius $\sphereRadius$, its intersection (if exists) with a directional ray $\{\rayo,\rayd\}$ is
+\begin{equation}
+\intersectionFunc(\sphereRadius,\rayo,\rayd) = \rayo + \left({\left((\rayo\cdot\rayd)^2-\norm{\rayo}^2+\sphereRadius^2\right)}^{\frac{1}{2}}-\rayo\cdot\rayd\right)\rayd,
+\label{eq:raySphereIntersection}
+\end{equation}
+where $\rayo$ and $\rayd$ are the ray's origin point and normalized direction respectively.
+% \begin{equation}
+% \left\{\intersectionFunc(\sphereRadius_i,\rayo,\rayd) \right\}\ |\ i\in[1,2,\dots,\sphereNum].
+% \end{equation}
+Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt$ represented in spherical coordinates, the ray connect them is with the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$.
+This ray may intersect with more than one spheres.
+Among them, the closest one to $\SpatialPt$ is:
+\begin{equation}\label{eq:closestPointImg}
+\begin{aligned}
+    \hat{\pt}&(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt) \triangleq \\
+    &\intersectionFunc(\sphereRadius_k,\rayo,\rayd(\SpatialPt,\rayo))\ |\ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
+\end{aligned}
+\end{equation}
+In the 3D space, the offset distance $\norm{\pt^\prime-\hat{\pt}}$ indicates the precision loss at $\pt$ from the representation. By integrating over all view and scene points, we obtain:
+\begin{equation}
+\begin{aligned}
+\sparseError(\sphereNum, \mathbf{\sphereRadius})  &= \iint \norm{\pt^\prime(\sphereNum,\mathbf{\sphereRadius},\SpatialPt)-\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)} \mathbf{d}\SpatialPt \mathbf{d}\rayo,\\
+&\forall \{\rayo, \SpatialPt\} \text{ pair without occlusion in between}.
+\label{eq:sparseError}
+\end{aligned}
+\end{equation}
+$\sparseError$ depicts how the generic representation precision of a scene, given a coordinate system defined by $\sphereNum$ and $\mathbf{\sphereRadius}$.
 
+In comparison, a neural representation aims at predicting projected image given a camera position $\rayo$ and orientation $\camDir$.
+Thus, we further extend \Cref{eq:sparseError} to image space to analyze the error given a set of camera's projection matrix  $\projectionMatrix(\rayo,\camDir)$ as
+%However, typical 3D representations are not uniformly distributed points.
+Moreover, the network only predicts pixel colors/intensities but not depth.
+By extending \Cref{eq:sparseError} to the captured image space data, we obtain
+\begin{align}
+\imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo, \camDir)  = \int \norm{\projectionMatrix(\rayo,\camDir)\cdot\left(\SpatialPt-
+\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)\right)}\mathbf{d}\SpatialPt.
+\label{eq:imageError}
+\end{align}
+From \Cref{fig:notations,eq:imageError}, we observe:
+given a fixed min/max range of $\mathbf{\sphereRadius}$, $\sphereNum$ is negatively correlated to $\imgSpaceError$;
+with a fixed $\sphereNum$, the correlation between distribution of $\sphereRadius_j$ and scene content (i.e., distribution of $\SpatialPt$s) also determines $\imgSpaceError$.
+
+However, for neural scene representation, infinitely enlarging $\sphereNum$ would significantly raise the challenges in training precision and inference performance.
+Likewise, increasing prediction and rendering densities (i.e., higher $\sphereNum$ or $\mlpLayerNum$/$\mlpChannelNum$) naturally improves the image output quality for lower \Cref{eq:imageError} while significantly increases the computation during ray marching, causing quality drop stretched along time.
+In the performance-sensitive VR scenario, the latency breaks the continuous viewing experiment and may cause simulator sickness.
+Thus, we determine the coordinate variables with a content-aware optimization towards optimal quality-speed performance.
+
+\paragraph{Spatial-temporal modeling}
+Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform a spatial-temporal joint modeling to determine the optimal coordinate system ($\sphereNum$ and $\mathbf{\sphereRadius}$) for positional precision and MLP settings ($\mlpLayerNum$/$\mlpChannelNum$) for color precision that adapt to individual computational resources and scene content.
+This is achieved via a latency-precision modeling in the spatial-temporal domain.
+%Specifically, we model the prediction image $P$ with gaze position ($G$) at time $t$ as $P_G(t)$ and the ground truth retinal image (from a simulated local rendering) as $\hat{P}_G(t)$.
+%In practice, increasing the number of spheres $\sphereNum$ or number of MLP layers may affect the computational and transmissional latency but will improve the quality of individual frames (as in \Cref{eq:imageError}).
+Thus, we define the perceptual error as:
 \begin{equation}
-    \begin{cases}
-        \theta = \mathrm{atan2}(x_{\pt_r}/z_{\pt_r})\\
-        \phi = \mathrm{acos}(y_{\pt_r} / r)
-    \end{cases}
+\begin{aligned}
+\finalError(\sphereNum, &\mlpLayerNum, \mlpChannelNum) = 
+\sum_t\int \mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)\times\\ &\norm{\projectionMatrix(\rayo_t,\camDir_t)\cdot\SpatialPt-\projectionMatrix(\rayo_{t-\latency},\camDir_{t-\latency})\cdot\pt(\sphereRadius_k,\rayo_{t-\latency},\rayd_{t-\latency})}\mathbf{d}\SpatialPt,
+%&\int  \norm{ \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_t, \rayd_t) - \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_{t-l(\sphereNum, \mathbf{\sphereRadius})}, \rayd_{t-l(\sphereNum, \mathbf{\sphereRadius})})) } \mathbf{d} t
+\label{eq:error:image}
+\end{aligned}
 \end{equation}
+where $\latency\triangleq\latency(\sphereNum, \mlpLayerNum, \mlpChannelNum)$ is the system latency with a given coordinate and network setting.
+$\mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)$ is the L1-distance between the four channels ($r$,$g$,$b$,$a$) and0 the highest values of network layer settings.
+For simplicity, we assumed uniformly distributed $\mathbf{\sphereRadius}$ with a fixed range of the spherical coverage.
 
-\dnc{(Jan 11) A simple derivation for the requirement of depth range and samples} Take 2D as an example.
-Consider two view centers $\rayo_1$, $\rayo_2$ on X axis and a point $\pt$ in 3D space.
-The radius of nearest sampling sphere of $\pt$ is $r_l$.
-The intersections of the sphere and rays from $\rayo_1$, $\rayo_2$ to $\pt$ are $\pt'_{1,r_l}$ and $\pt'_{2,r_l}$, respectively.
-$\Delta\pt'_{r_l}=\|\pt'_{2,r_l}-\pt'_{1,r_l}\|$ is small enough, then we have:
-
+Suggested by Albert et al.
+\shortcite{albert2017latency}, the latency for a foveated system shall reach below \textasciitilde$50$ms for undetectable artifacts.
+Given our test device's eye-tracking latency \textasciitilde$12ms$ and the photon submission latency \textasciitilde$14ms$ (\cite{albert2017latency}), the synthesis and rendering latency shall be less than $L_0 = 24$ms.
+Thus, we determine the optimal $\{\sphereNum, \mathbf{\sphereRadius}\}$ to balance latency and precision as
 \begin{equation}
-    \Delta\pt'_{r_l} \approx \Delta\rayo \frac{|z_\pt - r_l|}{z_\pt}
+\begin{aligned}
+    &\argmin_{\sphereNum, \mlpLayerNum, \mlpChannelNum} \finalError(\sphereNum, \mlpLayerNum, \mlpChannelNum), \\
+    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L_0
+\end{aligned}
 \end{equation}
-
-\begin{equation}
-    hello
-\end{equation}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.96\linewidth]{TOG/figs/depth_disparity.pdf}
-    \caption{Depth disparity}
-    \label{fig:method:eccentricity}
-\end{figure}
-
-With this coordinate system, the input of individual element in the 3D spaces is defined as $p=(r, \theta, \phi)$ and the output is color information $c=(r,g,b)$ and volume density $d$.
-We train the scene representation with an multilayer perception (MLP) network first.
-For the training uniformity and run-time efficiency, we do not distinguish different eccentricities during the training stage.
-Instead, the spatially variant visual acuity is represented as varying field of views in the rendering, i.e., data creation.
-Specifically, we depict the whole visual field as three semantic layers the fovea ({0-10 deg}), mid-periphery {6-60 deg}, and far periphery {y-110 deg}.
-In each frame, The uniformized networks returns three images with a identical resolution yet, as seen from the separation range in deg, gradually larger areas along eccentricity.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.96\linewidth]{TOG/figs/layers_together_zh.pdf}% try not overlapping the text on the image since we want the reviewers to see each pixel clearly
-    \caption{Visualization of the output at three eccentricity ranges.}
-    \label{fig:method:eccentricity}
-\end{figure}
-
-%estimate the fovea, mid-peripheral, peripheral field of view for a novel view next, fine-tune the fovea result with local view streamed from the cloud, and lastly warp/blend the three results for VR display.
-
-\paragraph{Input encoding}
-Since the input of our representation network is only three dimensional, we map the inputs to a higher dimensional space via high-frequency functions.
-\qisun{(Jan 9, 2021) What is bad if we have low-dim input?
-Harder to converge, higher error etc.?} \zh{(Jan 9) similar to NeRF, three components are not sufficient for achieving state-of-the-art quality, as demonstrated in Section 6.4).
-We introduce a positional encoding of the input coordinates that assists the MLP in
-representing high-frequency functions.}
-Thus we encoded the input through function:
-\begin{align}
-\iota (p) = (\sin(2^0 p), \cos(2^0 p), ..., \sin(2^{L-1}p, \cos(2^{L-1}p))
-\label{eqn:inputencode}
-\end{align}
-The function $\iota (\cdot)$ is applied to each of the 3D location representation ($r$, $\theta$, and $\phi$), we choose $L$ to be 10 in practice.
-Hence, the input dimension becomes 3 x 20 = 60 after encoding.
-The MLP network is composed with $N_{layer}$ fully-connected layers, each uses ReLu as activation function and $N_{channel}$ channels per layer and output the color and volume density.
-
-\paragraph{Synthesis}
-% the description of the volume density is the same as chapter 4 in NeRF paper.
-Similar to \cite{mildenhall2020nerf}, the trained representation is then deployed to synethesis final images via ray marching.
-For accelerating the performance, we bounded the near/far range to $20$ and $50$ meters respectively.
-The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
-\begin{figure}
-    \centering
-    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
-    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
-    \caption{Visualization of ray marching bounding.}
-    \label{fig:method:bounding}
-\end{figure}
-%For each pixel, we perform a ray from the camera and accumulate all the intersection's color weighted by the density accordingly between near bound and far bound.
-In practice, we experimented with different far bound values: 20 (meters) and 50 (meters), the difference between output is minimal.
-(\note{add figure for comparison}).
-We also experimented with the sampling number ($N_{sample}$) for accumulation between near bound and far bound to balance time cost and effects.
-Hence, we have an output image for the camera perspective.
-However, as a gaze-contingent method, the compromised quality would be unnoticeable in periphery. Thus, we further refine the visual quality in foveal vision, as detailed in \Cref{sec:method:refine}.
-Since we aim to render a novel view in real time and with perceptually high quality, we estimate the foveated image (10 degree), middle peripheral image (60 degree), and full peripheral image (110 degree) from the same novel view for further synthesis.
-Regarding foveated rendering, we estimate an image with size of 64 x 64, $N_{layer}=12$, $N_{channel}=64$, and $N_{sample}=16$.
-Regarding middle peripheral and full peripheral rendering, we estimate the two images with size of 256 x 256, $N_{layer}=8$, $N_{channel}=64$, and $N_{sample}=4$.
-We next upsample middle peripheral to ???
-x ???
-and full peripheral to 1440 x 1440 for blending.
-\nothing{
-\subsection{Refinement for Fovea Layer}
-\label{sec:method:refine}
-Due to the depth bounding and prediction errors, it is particularly challenging to preserve pixel-wise quality without compromising real-time performance.
-However, the human foveal vision is remarkably sensitive to subtle details across all frequency ranges.
-To
-To maximize the quality of foveated image, we utilize the limited but consistent network bandwidth by retrieving a local peripheral image from the cloud.
-The retrieval was realized by cloud-based rendering using users' upstreamed head and gaze parameters.
-Due to the inevitable dual-way transmission, it is impossible to instantly obtain the image without latency. Displaying outdated frames have been studied as a major cause of simulator sickness.
-}%nothing
-\note{describe when to retrieve new full peripheral image.} We leverage local guidance views to fine tune the output foveated image.
-Local guidance view is decided according to current camera view.
-We chose the nearest 4 views (or 8 views) as reference views and estimate the color as well as volume density for all reference views.
-Every time we receive a new peripheral image from the cloud, we calculate the delta between the ground truth and our result for reference views for preparation.
-Then we warp the delta into target camera view, calculate the average and apply to the fovea result for each frame.
-
-\subsection{Real-time Rendering}
-% \subsection{Fovea-Peripheral Blending}
-\label{sec:method:blending}
-Lastly, we apply alpha blend on the fine-tuned fovea view and middle peripheral view, as well as on the middle peripheral view and full peripheral view via function smo/:
-
-\subsection{Latency-quality joint optimization}
-\label{sec:method:optimization}
-Increasing rendering sampling naturally improves the image output quality. However, it increases the latency, causing quality drop stretched along time, and more critically, simulator sickness.
-Inspired by \cite{Li:2020:TSP}, we perform a spatial-temporal joint optimization to determine the sampling that adapts to individual computational resources.
-This is achieved via a Pareto efficiency optimization.
-
-$E_{q}\triangleq $
-
-$E_{l}$ is defined as the loss introduced by latency. 
-\begin{align}
-    \int E_{l} \mathbf{d} t
-\end{align}
-
-\begin{figure*}
-    \centering
-    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
-    \caption{Latency-quality joint-optimization.}
-    \label{fig:optimization}
-\end{figure*}
-
-%\subsection{Cloud-Based System}
-%\label{sec:method:system}
-%Finally, we deploy our gaze-contingent synthesizer to a real-time cloud-based streaming system.
+\Cref{fig:optimization} visualizes an example of the optimization mechanism for in fovea.
+The optimized results ($\sphereNum, \mlpLayerNum, \mlpChannelNum$) for individual networks are detailed in the next section.
+The visual quality is validated by psychophysical study (\Cref{sec:study:user}) and objective analysis (\Cref{sec:study:quality}).
+The latency breakdown of our system is reported in \Cref{sec:study:intra}.
+\input{TOG/fig_optimization}
+\input{TOG/4.5-implementation}

### 2021-01-28 TOG/4.5-implementation.tex
last day

\section{Implementation}\label{sec:impl}
In this section, we elaborate on how we collect data for training, the specific parameters that we choose for two orthogonal neural networks, and system integration.

\paragraph{Data generation}
For each scene, we trained two independent neural networks: foveal and periphery. We sampled the rendered view for two networks accordingly. To train the foveal neural network, we rendered an image of resolution $256^2$ per 20 deg. For periphery neural network, we rendered an image of resolution $256^2$ per $60$ deg. We repeated the above every $5$ cm on each axis.
% \nothing{
%\paragraph{Input encoding}
Similar to \cite{mildenhall2020nerf}, we also extend the input dimensions ($d=3$) to a higher dimensional ($d=60$) for training. 
\nothing{This is achieved via coding the input as:
\begin{align}
\iota (p) = (\sin(2^0 p), \cos(2^0 p), ..., \sin(2^{L-1}p, \cos(2^{L-1}p))
\label{eqn:inputencode}
\end{align}
The function $\iota (\cdot)$ is applied to dimension of $\SpatialPt=(r_\SpatialPt,\theta_\SpatialPt, \phi_\SpatialPt)$.
In our experiment we choose $L=10$.\qisun{What is $L$?}\dnc{It's an encoding parameter for input. maps 3D input to $R^{6L}$ space, as \autoref{eqn:inputencode} shows}
}
\paragraph{Optimized network variables}
Following the optimization results from \Cref{sec:method:optimization}, we implemented both foveal and periphery neural network with $\mlpLayerNum=128$ and $\mlpChannelNum=4$ , and $\sphereNum = 32$ for foveal network and $\sphereNum = 16$ for periphery network. We will discuss the visual quality and performance difference in ~\Cref{sec:study:intra}. \nothing{We sampled the near foveal network and far foveal with $N_{sample}=16$ and $N_{sample}=16$, while we sampled the near ($1/distance < 0.5$) peripheral with $N_{sample}=8$,  $N_{channel}=64$,  $N_{layer}=4$ and far ($1/distance >= 0.5$) peripheral with $N_{sample}=8$,  $N_{channel}=128$,  $N_{layer}=4$.}
\nothing{
\dnc{Move to section 3.4 or just drop it (cause we have no detail analysis about the bound)? This small paragraph seems strange at here.}
\paragraph{Ray marching bounding}
For accelerating the performance and quality (as compared in \Cref{fig:method:bounding}), we bound the min/max of $\mathbf{\sphereRadius}$. The bounding range, together with $\sphereNum$, $\mlpLayerNum$ and $\mlpChannelNum$, are determined from a spatial-temporal optimization, as detailed in \Cref{sec:method:optimization}. %The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
\input{TOG/fig_ray_bounding}
}%nothing
\paragraph{Integration} \dnc{this paragraph seems similar to 3.3.}
The stereoacuity significantly drops beyond fovea \cite{mochizuki2012magnitude}.
Our system reaches interactive performance (around 30FPS) for the first design -- the same procedure is applied to each eye (the details are addressed in \Cref{sec:study:intra}). To archive real-time performance (around 50FPS), we implemented mono-periphery for better performance. For each frame, we infer the same mid- and far- peripheral images for left and right eyes from the camera pose of central eye. In \Cref{sec:study:user}, we evaluated mono-periphery for subjective and objective measurements as well as discuss the visual quality and performance difference in ~\Cref{sec:study:intra}.
% \warning{We conducted a preliminary study to investigate how users perceive mono-periphery.}

\paragraph{Environment}
The system was implemented through OpenGL framework using CUDA and accelerated by TensorRT with Intel(R) Xeon(R) Gold 6230R CPU @ 2.10GHz (256GB RAM) and one NVIDIA GTX 3090 graphics card.
For each scene, we trained around 200 epochs for experiment use (about a day).

### 2021-01-28 TOG/5-result.tex
last day

@@
 \label{sec:result}
-% use one net to represent a complicated scene
-% we still have other option for real-time, like streaming
-\note{
-Ablation studies for the network training/losses
-Target scene | Foveated target scene | NeRF | network predicted images ---> at least 4 different scenes.
-streaming time for 4-5 different scenes: full resolution vs our network prediction for “same data size”
-image quality of rendered scene vs our network prediction for “same amount of streaming time”
-User study results
-Table on performance comparison, streaming time, PSNR, SSIM, etc
-}%note
-With a variety of scenes as shown in \Cref{fig:results:comparison2}, we evaluate our method with subjective studies (\Cref{sec:study:user}), objective analysis (\Cref{sec:study:quality}), and intra-system efficiency (\Cref{sec:study:intra}).
 \input{TOG/fig_tile_results}
+With a variety of scenes as shown in \Cref{fig:results:comparison}, we first conduct subjective studies (\Cref{sec:study:user}) and objective measurements (\Cref{sec:study:quality}) to evaluate the perceptual quality.
+Further, we analyze the intra-system workings and compare inter-system performance in \Cref{sec:study:intra}.
 
@@
 % \zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
-
 % In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
-We first conducted a perceptual experiment to investigate how users perceive our solution ({\bf OURS}) compared to existing solution (\cite{mildenhall2020nerf}, {\bf NeRF}) and original mesh with foveated rendering (\cite{perry2002gaze}, {\bf F-GT}).
+We first conducted a psychophysical experiment to investigate how users perceive our solution ({\bf OURS}) compared with existing neural view synthesis methods (\cite{mildenhall2020nerf}, {\bf NeRF}) and foveated rendering with the full mesh data (\cite{perry2002gaze}, {\bf F-GT}).
 
@@
 \paragraph{Stimuli}
+For precise comparison and accommodating the low frame rate of alternative solutions (as to be compared in \Cref{sec:study:intra}), each group of stimuli were static images rendered via {\bf F-GT} or synthesized via  {\bf OURS}/{\bf NeRF} with the same views, as shown in \Cref{fig:results:comparison}.
+They were generated with the same and randomly defined gaze fixation across conditions.
+The resolution of the image per eye was $1440 \times 1600$. 
+During the study, the target gaze position was indicated as a green cross on the stimuli images.
+% The resolution of the images per eye was $1440 \times 1600$. 
+We used the {\it gas} and {\it minecraft} scenes for the diverse indoor/outdoor environment.
+Each condition from an individual scene consists of \nothing{$1$ view and} $3$ different gaze positions.
 
-The stimuli were the images rendered via {\bf F-GT} or synthesized via  {\bf OURS}/{\bf NeRF}, as shown in \Cref{fig:results:comparison}. The resolution of the image per eye was $1440 \times 1600$. 
-For fair comparison, each group of stimuli were generated with the same, randomly defined gaze p-osition. The position was indicated as a green cross on the stimuli images.
-The resolution of the images per eye was $1440 \times 1600$. 
-We used the {\it gas} and {\it minecraft} scenes.
-Each condition from individual scene consists of $1$ view and $3$ different gaze positions.
-
-Specifically, we controlled the foveal kernel size in {\bf F-GT} to match our display size and eye-display distances.
+The foveal kernel size in {\bf F-GT} was designed to match our display size and eye-display distances.
 For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $N_{sample}=16$, $N_{layers}=4$, $N_{channel}=128$.
-The last column of \Cref{fig:results:comparison} shows the result.
+The bottom right of each scene figure \Cref{fig:results:comparison} shows the resulting images.
 %We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.
@@
 \paragraph{Setup}
-Each participant was asked to wear a eye-tracked HTC Vive Pro Eye headset during the experiment. During the experiment, the participants wearing the headset examining  the stimuli. 
-Fourteen users participated in the study and 13 of them managed to complete (4 females and 9 males, $M=23.08$, $SD=1.75$).
-One participant (age=23, female) did not complete the experiment due to gaze drifting (please refer to the task description below).
-None of the subjects were aware of the research, the experimental hypothesis, or the number of rendering methods. All participants had normal or corrected-to-normal vision.
+Each participant wore an eye-tracked HTC Vive Pro Eye headset and remained seated to example the stimuli during the experiment.
+Ten users participated in and completed the study (3 females, $M=23.7$, $SD=1.49$). 
+% One participant (age=23, female) did not complete the experiment due to constant gaze drifting (please refer to the task description below).
+None of the subjects were aware of the research, the experimental hypothesis, nor the number of methods. All participants had a normal or corrected-to-normal vision.
 
@@
 The task was a \textit{two-alternative-forced-choice} (2AFC).
-Each trial consists of a pair of stimuli from the three conditions ({\bf OURS} / {\bf NeRF} / {\bf F-GT}).
-Each stimulus appeared for $300$ms on the the display.
-A forced $1.5$sec break was introduced between conditions.
+Each trial consists of a pair of stimuli generated from two of the three methods ({\bf OURS} / {\bf NeRF} / {\bf F-GT}) with a sampled view and gaze position.
+Each stimulus appeared for $300$ms on display.
+A forced $1.5$sec break (black screen) was introduced between conditions to clear the vision.
 During the study, the participants were instructed to fix their gazes on the green cross.
-To prevent fixation from shifting away, we tracked the users gaze during the whole experiment.
-Whenever the gaze is $1$deg away from the target, a trial was dropped immediately with black green informing the participant.
-After each trial, the participants were instructed to select which one of the two stimuli appeared with higher visual quality using keyboard.
-Before each experiment, a warm-up session with 1 trial was provided to have the participants familiarize the study procedure.
-The order of conditions cmong trials were randomized with counterbalancing.
+To prevent the fixation from shifting away and ensure accuracy, we tracked the users' gaze throughout the experiment.
+Whenever the gaze is more than $5$deg away from the target, a trial was dropped immediately with a black screen informing the participant.
+After each trial, the participants were instructed to select which of the two stimuli appeared with higher visual quality using a keyboard.
+Before each experiment, a warm-up session with 6 trials was provided to familiarize the participants with the study procedure.
+The orders of conditions among trials were randomized with counterbalancing.
+The entire experiment of each user consists of $36$ trials, $12$ trial per pair of conditions.
 %To start with, each participant was informed that they were required to choose an alternative with higher quality after watching each pair of stimuli.
@@
 After each pair, participants reported their opinions on which image has higher quality.
-To minimize the effect of accumulating discomfort during the experiment, we enforced at least a 60-second break between each trial.
-Meanwhile, the participants were instructed to take as much time as needed to recover.
+To minimize the effect of accumulated fatigue, we enforced breaks at least a 2-second between trials and 60-second after each scene.
+Meanwhile, the participants were able to take as much time as needed.
 
 \paragraph{Results}
-\Cref{fig:results:2afc} plots the study results. 
-Among all three conditions, we observed near to random guess among trails that compares  {\bf F-GT} and {\bf OURS} (50.0\% voted for {\bf OURS}).
-However, participants significantly prefers  {\bf F-GT} over {\bf NeRF} (98.7\% voted for NeRF, binomial test showed $p<0.005^{***}$).
+\input{TOG/fig_vote}
+\Cref{fig:results:2afc} plots the results considering both methods and their orders in the 2AFC experiments.
+Among all three conditions, we observed close-to-random-guess among trails that compares {\bf F-GT} and {\bf OURS} ($53.3\%$ voted for {\bf OURS}, $SD=14.9\%$).
+Meanwhile, a significantly higher ratio of voting {\bf F-GT} over {\bf NeRF} was observed ($92.5\%$, voted for {\bf F-GT}, $SD=11.4\%$, binomial test showed $p<0.005^{***}$).
 The preference applies to {\bf OURS} vs.
-{\bf F-GT} as well (97.4\% voted for NeRF, binomial test showed $p<0.005^{***}$).
+{\bf NeRF} as well ($91.7\%$ voted for {\bf OURS}, $SD=14.8\%$, binomial test showed $p<0.005^{***}$).
+% \qisun{(Jan 23, 2021) @Zhenyi: consider adding stdev to the average numbers above?
+This shall be reflected in \Cref{fig:results:2afc} as well.
+Also, double check the numbers between here and the plot as they are not matching now.}
 
@@
 % nerf/gt 2/156 1.3407579916082842e-43
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.96\linewidth]{TOG/figs/vote.pdf}
-    \caption{The users' preference votes from our evaluation experiment. X axis shows all the pair combination and Y axis shows the ratio of voting for the second one.}
-    \label{fig:results:2afc}
-\end{figure}
+% our/gt 64/120 0.26149549194499716
+% our/nerf 110/120 9.586645257178695e-23
+% nerf/gt 9/120 8.546460059714397e-24
 
@@
 \paragraph{Discussion}
-The near-to-random-guess indicated the statistically similar perceptual quality between {\bf OURS} and foveated ground truth ({\bf F-GT}). 
-Given the perceptual identicality between foveated and full-resolution rendering, the subjective study indicates that {\bf OURS} can achieve similar quality as a rendering with a locally fully stored high-quality mesh.
-Meanwhile, both of the two conditions showed significant subjective quality preference than {\bf NeRF}.
-That is, under an immersive, high FoV, and first-person view, the current volume-based scene representation may not fully reproduce the perceptually identical retinal images.
-To further validate the perceptual quality by breakdown across the whole visual field, we conducted measurements with objective metrics in the following section.
+The close-to-random-guess indicated the statistically similar perceptual quality between {\bf OURS} and foveated ground truth ({\bf F-GT}). 
+Given the perceptual identicality between foveated and full-resolution rendering, the subjective study reveals that {\bf OURS} can achieve similar quality as a rendering with locally fully stored high-quality 3D data.
+
+Meanwhile, both of the two conditions showed significant quality preference than {\bf NeRF}.
+That is, with immersive, high FoV, and first-person viewing constraints, {\bf OURS} synthesizes gaze-contingent retinal image with superior perceptual quality than alternative synthesis solutions.
+The latter typically aims to reconstruct the full image over the entire field-of-view.
+\nothing{the current volume-based scene representation may not fully reproduce the perceptually identical retinal images.}
+To further validate the perceptual quality at individual eccentricities across the whole visual field, we analyze with objective metrics in the following section.
 
@@
 \label{sec:study:quality}
+\input{TOG/fig_lpips}
 Complementary to the subjective measurement (\Cref{sec:study:user}), we evaluate the perceived quality in an objective fashion.
-Specifically, under four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT} and {\bf NeRF} at individual eccentricity ranges across the whole visual field (up to $110$ deg, the capability of the VR HMD).
-%\zh{gonna comment this: To validate the non-uniform perceptual quality on the retina, we apply a foveated filtering to {\bf GT} \cite{perry2002gaze,jiang2015salicon}.}
+Specifically, under all four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT} and {\bf NeRF} at individual eccentricity ranges (up to $110$ deg, the capability of the VR HMD) via partial images.
 For each eccentricity range, we compare the deep perceptual similarity (LPIPS) \cite{zhang2018unreasonable} across all scenes.
-LPIPS uses deep neural networks to evaluate perceptual similarity between the image provided and a reference image.
+LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image.
 Smaller values indicate higher perceptual similarity.
 For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $420$ data per scene.
-We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity value and together.
+We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together.
 Paired t-tests with Holm correction were used for all pairwise comparisons between stimuli.
@@
 \paragraph{Results} 
-\Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg per sample). 
-From foveal to near periphery (<=$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.84$).
-That is, \textbf{OUR} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), while significantly higher LPIPS than \textbf{F-GT} ($p<.001^{***}$).
-For example, the main effects of stimuli ($F(2,38)=42.023, p=.001^{***}$) was significant on eccentricity $=25$  deg  in scene \textit{gas} (\Cref{fig:lpips:gas}).
-\textbf{OUR} was significantly lower than \textbf{NeRF} ($t(19)=-18.557, p<.001^{***}$) and \textbf{F-GT} ($t(19)=-10.342, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
+% \warning{need to update}
+\Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg step size). 
+From foveal to near eccentricity (<=$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.26$).
+That is, \textbf{OURS} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), although higher than \textbf{F-GT} ($p<.001^{***}$).
+For example, the main effects of stimuli ($F(2,38)=210.76, p<.001^{***}$) was significant on eccentricity $=[0,25]$ deg in scene \textit{gallery} (\Cref{fig:lpips:gallery}).
+\textbf{OURS} was significantly lower than \textbf{NeRF} ($t(19)=-19.54, p<.001^{***}$) and \textbf{F-GT} ($t(19)=-4.32, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
+The example observation generally applies to all 4 scenes being validated.
 
-From mid- to far- periphery (>$40$ deg ), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.19$).
-\textbf{OUR} shows significantly higher LPIPS than \textbf{NeRF} due to the foveated nature of our design.
+From mid- to far- periphery (>$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.16$).
+\textbf{OURS} shows significantly higher LPIPS than \textbf{NeRF}.
 Whereas, comparing with \textbf{F-GT}, we observed significant lower scores ($p<.001^{***}$).
-For instance, the main effects of stimuli was significant on eccentricity $=60$ deg in scene \textit{gas} ($F(2,38)=493.51, p<.001^{***}$, \Cref{fig:lpips:gas}).
-\textbf{OUR} was significantly lower than \textbf{F-GT} ($t(19)=-12.28, p<.001^{***}$), and higher than \textbf{NeRF} ($t(19)=18.90, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
-
-The general trend applies to all 4 scenes being validated.
+For instance, the main effects of stimuli was significant on eccentricity $=60$ deg in scene \textit{bedroom} ($F(2,38)=411.93, p<.001^{***}$, \Cref{fig:lpips:bedroom}).
+\textbf{OURS} was significantly lower than \textbf{F-GT} ($t(19)=-3.78, p<.001^{***}$), and higher than \textbf{NeRF} ($t(19)=22.75, p<.001^{***}$) both with a ``large'' effect size (Cohen's d $>0.8$).
+The example observation generally applies to all 4 scenes being validated.
 
 \paragraph{Discussion}
-The observation revealed our method's over-performance than alternative view synthesis solutions in the foveal and near peripheral vision.
-This is evidenced by the significantly increased perceptual similarity to \textbf{GT} while comparing between \textbf{OUR} and \textbf{NeRF}.\zh{``significantly increased LPIPS'' requires analysis}
+In the fovea and near-periphery, the observation revealed our method's significantly higher perceptual quality than alternative neural representation solutions.
+This is evidenced by the significantly stronger perceptual similarity to \textbf{F-GT} by comparing between \textbf{OURS} and \textbf{NeRF}.
+The subjective study (\Cref{sec:study:user}) evidenced the marginally lower similarity than \textbf{F-GT} as statistically unnoticeable among users.
 
-Meanwhile, \textbf{OUR}'s increased perceptual similarity comparing with \textbf{F-GT} beyond $40$ deg indicated their similar visual quality.
-This shows that \textbf{OUR} doesn't compromise the peripheral vision's quality with its significantly enhanced foveal image.
+In the far-periphery, \textbf{OURS} showed increased LIPIPS than \textbf{F-GT} beyond $40$ deg.
+The latter has been shown to display identical perceptual similarity than full resolution rendering in VR \cite{Patney:2016:TFR}.
+Thus, the findings show that \textbf{OURS} doesn't compromise the peripheral vision's quality with its significantly enhanced synthesis acuity in fovea.
 
-The discoveries also agrees with the trend from \Cref{sec:study:user}.
-That is, in addition to the faster performance, our method showed remarkably perceptual quality gain than \textbf{NeRF} for first-person 6DoF immersive viewing of 3D scenes: our synthesized views are perceptually identical to the traditionally rendered images from complete mesh.
-It yet achieves significant data storage saving (\warning{xxx\%-xxx\%}).
+Note that the discoveries also agree with the observations from \Cref{sec:study:user}.
+That is, in addition to the significantly faster performance  ($99.7\%$ time reduction per frame as in \Cref{sec:study:intra}\nothing{\qisun{Check if I noted it right.}}), our method showed superior perceptual quality than \textbf{NeRF} under first-person, high resolution, and immersive viewing of 3D scenes.
+Meanwhile, our synthesized views are perceptually identical to the traditionally rendered images with a complete 3D mesh.
+It yet achieves significant data storage-saving with only about $0.5$mb (up to $99.5\%$ in our experiment).
 
-% ZH: why \Cref{fig:lpips:gas} shows 'section 5.2'
-\begin{figure*}
+\subsection{Performance}
+\label{sec:study:intra}
+% apply objective metric to |cref{fig:optimization}, aka different sample numbers
+Virtual and augmented reality demands high frame rates along with high quality to ensure immersive and comfortable experience.
+Our neural synthesis method introduces both spatial and angular (stereo) visual perception for real-time performance (\Cref{sec:method:representation}).
+Further, our precision-latency joint optimization for the representation and network also balances quality and performance.
+Here, we evaluate the performance of each component in our system and compare with existing neural synthesis solutions.
+
+For high resolution ($1440\times1600$), high field-of-view ($110$deg) stereo images required by VR display, our system completes all computation (including gaze-contingent neural-inference and elemental images composition) in $31.8$ms per frame without stereo foveation, \dncR{as shown in \Cref{tbl:ablation} \st{shows a time consumption breakdown of our method, together with {\bf NeRF}}}.
+Compared with {\bf NeRF} \dncR{(the bottom row of \Cref{tbl:ablation})}, our spatial foveation \dncR{\st{for adaptive resolution}} significantly improves the system performance from offline computation (in seconds) to interactive speed (\dncR{about 30 frames per second\st{$31.8$ ms per frame}}).
+
+Tailored for VR displays, our \dncR{stereo foveation\st{adaptive stereo synthesis}} further reduces the computation time \dncR{and achieves high performance of more than 50 frames per second \st{by more than $10$ms/frame}}, contributing to a temporally continuous viewing experience without loss of perceived resolution and quality.
+
+%Specifically, the significant performance gain benefits from the spatial and stereo foveation that do not compromise the visual quality.
+\begin{table}[!htb]
     \centering
-    \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenegas.pdf}}\label{fig:lpips:gas}
-    \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}
+    \begin{tabular}{ l|r } 
+        \toprule
+         foveal infer (per eye)  & 3.8 \\
+         \midrule
+         periphery infer (per eye) & 11.9 \\
+          \midrule
+         blending \& contrast enhancement & 0.2 \\
+          \midrule\midrule
+         {\bf OUR} method without stereo foveation & 31.8\\
+         {\bf OUR} full method & {\bf 19.7}\\
+         \midrule
+         {\bf NeRF} \cite{mildenhall2020nerf} & $9\times10^3$ \\
+         \bottomrule
+    \end{tabular}
+    \Caption{Time consumption breakdown and comparison.}
+    {%
+    The numbers show the time consumption of each component and the overall system per frame. All units are in millisecond.
+    \nothing{\qisun{(Jan 25) I have update the time consumption of {\bf NeRF}. See if that is correct.}}
+    }
+    \label{tbl:ablation}
+\end{table}
 
-    \subfloat[gallery]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenegallery.pdf}}
-    \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
-    \caption{LPIPS analysis of all scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity.}
-    {\zh{Add intersection lines later.}}
-    \label{fig:lpips}
-\end{figure*}
+% \begin{table*}[!htb]
+%     \centering
+%     \begin{tabular}{ |c|c|c|c|c|c|c| } 
+%         \hline
+%          & foveal infer & periphery infer & blending & contrast & system without mono-periphery & system with mono-periphery \\
+%         \hline
+%         without sub-net & \warning{3.2\times2} & \warning{10} & \multirow{2}{3em}{\warning{2}} & \multirow{2}{3em}{\warning{2}} & 30.4 & 20.4 \\ 
+%         with sub-net& 3.5 & 12 & & & 35 & 23 \\ 
+%         \hline
+%     \end{tabular}
+%     \caption{Performance.}
+%     \label{tbl:ablation}
+% \end{table*}
 
-\subsection{Intra-System Workings}
-\label{sec:study:intra}
-\paragraph{Spatial-Temporal Optimality}
+% \paragraph{sub-net}
+% As described in \cref{sec:method:optimization} and \Cref{sec:impl}, we find the optimal parameters for our network and further split the network into two sub-nets based on the distance.
+Without sub-net, it takes \warning{3.2}ms to infer an image from foveal neural network and \warning{10}ms to infer mid- and far- peripheral images.
+With sub-net, it takes \warning{3.5}ms to infer an image from foveal neural network and \warning{12}ms to infer mid- and far- peripheral images.
+The results are shown in \warning{Fig.xxx}.
+We calculated the LPIPS value and can see it is increased by \warning{xxx} with sub-net design.
 
-% apply objective metric to |cref{fig:optimization}, aka different sample numbers
-
-\paragraph{Ablation Study}
+%\paragraph{Ablation Study}
 % the performance of an AI system by removing certain components, to understand the contribution of the component to the overall system

### 2021-01-28 TOG/6-conclusion.tex
last day

\section{Conclusion}
\label{sec:conclusion}
\zh{consistently using high eccentricity rather far periphery?}
In this paper, we present a gaze-aware neural scene representation and view synthesis method tailored for future portable virtual reality viewing experiences. Especially, we overcome the limitations of existing neural rendering approaches for high performance, resolution, and fidelity. Our network individually synthesizes foveal, mid- and far-periphery retinal images which are then blended to form a wide field of view image matching retinal acuity. This is achieved via adapting the neural network model to the foveated visual acuity and stereopsis. 
Comparing with traditional rendering, our method requires less than $1\%$ data storage yet delivering identically high perceptual quality.
Comparing with alternative neural view synthesis approaches, our method creates significantly faster and higher fidelity viewing for high resolution, high FoV, and stereo VR head-mounted-displays.
%The experiments with user studies, objective measurements, and intra-system analysis show our significant effectiveness of delivering perceptually identical visual contents in real-time.

\paragraph{Limitations and future work}
While our method achieves unprecedented performance compared to existing approaches, it suffers from a few limitations. Our multi-spherical representation introduces higher quality when the VR camera is within the first sphere. As shown in \Cref{fig:large_translation}, when the virtual camera translates largely within a highly occluded scene, our method still synthesizes the accurate views but with a decline in foveal image quality. This is due to the accumulated error of the concentric spherical representation (\Cref{eq:imageError}).
Multiscale coordinates and networks that consider various level-of-details of the 3D space have shown their effectiveness of interpolating geometries \cite{winkler2010multi}. Developing a multiscale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, to extend the applicability of our method to large scale scenes with strong occlusions.
\begin{figure}[!ht]
    \centering
    \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.pdf}\label{fig:large_translation:small}}\hspace{1em} % or 2_gas_0.3.png
    \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.pdf}\label{fig:large_translation:large}} % or 2_gas_0.9.png
    \Caption{Comparing foveal quality with regards to translation range. }
    {%
    Using our method, \subref{fig:large_translation:small} shows a fovea image \dnc{from training dataset} with a small-scale translation box (0.3 meters); \subref{fig:large_translation:large} shows a fovea image with a large-scale translation box (0.9 meters). In the 2D illustrative inset: the blue circles indicates the distribution of the spheres; the orange arrow shows the eye's translation range.
    }
    \label{fig:large_translation}
\end{figure}

Currently, we sample the scene fully based on eccentricity, considering acuity and stereopsis. However, we envision that fine grained visual sensitivity analysis, such as luminance \cite{Tursun:2019:LCA} or depth \cite{Sun:20:OE}, would provide more insights on achieving further higher quality and/or faster performance. Also, our multi-spherical representation is specialized to the scene for optimal quality. Thus, our network is trained per-scene. Exploring potential means of generalized representation may further reduce storage.

For simplicity, the spatial-temporal joint optimization connects the output precision and latency to the number of the spheres ($\sphereNum$) but not their radii $\mathbf{\sphereRadius}$. This is due to the significant amount of training with parameter sampling. Incorporating the parameters into single training process may significantly reduce the time consumption for the optimization. With the adaptive training process, a content-aware distribution (i.e., $\mathbf{\sphereRadius}$) of the spheres would further improve the synthesis quality and performance.
%\qisun{note to myself: Right now we use did not optimize for radii but only number of spheres}

### 2021-01-28 TOG/ccs.tex
last day

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010371</concept_id>
<concept_desc>Computing methodologies~Computer graphics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003121.10003124.10010866</concept_id>
<concept_desc>Human-centered computing~Virtual reality</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept_id>10010147.10010371.10010387.10010393</concept_id>
<concept_desc>Computing methodologies~Perception</concept_desc>
<concept_significance>500</concept_significance>
</concept>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Perception}
\ccsdesc[500]{Human-centered computing~Virtual reality}

\keywords{virtual reality; gaze-contingent graphics; neural representation}

### 2021-01-28 TOG/fig_lpips.tex
last day

@@
-\begin{figure*}
+% \begin{figure*}[!ht]
+%     \centering
+%     \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips-fov-gas-group.pdf}}\label{lpips:gas}
+%     \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}
+
+%     \subfloat[studyroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenestudyroom.pdf}}
+%     \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
+%     \caption{LPIPS analysis of four scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity.}
+%     {We found a significant difference in \warning{ecc range} between \warning{stimulus} and \warning{stimulus};}
+%     \label{fig:lpips}
+% \end{figure*}
+
+% ZH: why \Cref{fig:lpips:gas} shows 'section 5.2'
+\begin{figure*}[ht]
     \centering
-    \subfloat[gas]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips-fov-gas-group.pdf}}\label{lpips:gas}
-    \subfloat[minecraft]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenemc.pdf}}
+    \subfloat[gas]{\label{fig:lpips:gas}\includegraphics[width=0.48\linewidth]{TOG/figs/lpips/lpips-gas-84.pdf}}
+    \subfloat[minecraft]{\label{fig:lpips:mc}\includegraphics[width=0.48\linewidth]{TOG/figs/lpips/lpips-mc-84.pdf}}
 
-    \subfloat[studyroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenestudyroom.pdf}}
-    \subfloat[bedroom]{\includegraphics[width=0.48\linewidth]{TOG/figs/lpips_scenebedroom.pdf}}
-    \caption{LPIPS analysis of all scenes as well as comparison of our method, NeRF, and foveated GT on each eccentricity. $^*$: $p<.05$, $^{**}$: $p<.01$.}
-    {We found a significant difference in \warning{ecc range} between \warning{stimulus} and \warning{stimulus};}
+    \subfloat[gallery]{\label{fig:lpips:gallery}\includegraphics[width=0.48\linewidth]{TOG/figs/lpips/lpips-gallery-84.pdf}}
+    \subfloat[bedroom]{\label{fig:lpips:bedroom}\includegraphics[width=0.48\linewidth]{TOG/figs/lpips/lpips-bedroom-84.pdf}}
+    \Caption{LPIPS visualization of all scenes as well as comparison of {\bf OURS}, {\bf NeRF}, and {\bf F-GT} over the visual field.}
+    {%
+    X-axis indicate eccentricity range ($0$ to 110).
+Y-axis shows the LPIPS loss \cite{zhang2018unreasonable} with a full resolution rendering as reference.
+Lower values mean more similar perceptual quality to the reference.
+The semi-transparent layer indicate first and third quarterlies.
+The inset figures' corresponding eccentricity ranges are indicated by the dashed lines.
+    }
     \label{fig:lpips}

### 2021-01-28 TOG/fig_mono_comparison.tex
last day

\begin{figure}[hb]
    \centering
    %\subfloat[w/o gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/stereo_periphery/bedroom_view0000_blended_stereo.png}\label{fig:mono:wo}}
    %\subfloat[w/ gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/mono_periphery/bedroom_view0000_blended_stereo.png}%\label{fig:mono:w}}
    %\subfloat[w/o gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/stereo_periphery/gallery_view0000_blended_stereo.png}\label{fig:mono:wo}}
    %\subfloat[w/ gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/mono_periphery/gallery_view0000_blended_stereo.png}%\label{fig:mono:w}}

    \subfloat[w/o gaze-adaptive stereo]{\includegraphics[width=0.48\linewidth]{TOG/figs/stereo_periphery/stereo_periphery.pdf}\label{fig:mono:wo}}%\hspace{0.5em}
    \subfloat[w/ gaze-adaptive stereo]{\includegraphics[width=0.48\linewidth]{TOG/figs/stereo_periphery/mono_periphery.pdf}\label{fig:mono:w}}
    %\subfloat[w/o gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/stereo_periphery/gas_view0000_blended_stereo.png}\label{fig:mono:wo}}
    %\subfloat[w/ gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/mono_periphery/gas_view0000_blended_stereo.png}\label{fig:mono:w}}

    %\subfloat[w/o gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/stereo_periphery/gas_view0001_blended_stereo.png}\label{fig:mono:wo}}
    %\subfloat[w/ gaze-adaptive stereo]{\includegraphics[width=0.248\linewidth]{TOG/figs/mono_periphery/gas_view0001_blended_stereo.png}\label{fig:mono:w}}
    \Caption{Visualization of the adaptive stereo-acuity with anaglyph.}
    {%
    \subref{fig:mono:wo} shows the rendered image with 6 retinal sub-images ($\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$).
    \subref{fig:mono:w} shows the rendered image with our adaptive and accelerated inference considering foveated stereoacuity ($\imageFoveal^{\{l,r\}}$, $\imageMid^{\{m\}}$, and $\imageFar^{\{m\}}$). Our method preserves full stereopsis in the fovea while reducing the angular resolution in the periphery for accelerated inference.
    }
    \label{fig:mono}
\end{figure}

### 2021-01-28 TOG/fig_notations.tex
last day

\begin{figure}
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/parameters.pdf}
    % \subfloat[Notations]{\includegraphics[width=0.96\linewidth]{TOG/figs/parameters.pdf}\label{fig:variables}}

    % \subfloat[Correlation between loss and parameters]{\includegraphics[width = 0.96\linewidth]{example-image-a}}
    \caption{Coordinate system and variable annotations.}
    \label{fig:notations}
\end{figure}

### 2021-01-28 TOG/fig_optimization.tex
last day

\begin{figure}[htb]
    \centering
    \subfloat[quality]{\includegraphics[width=0.48\linewidth]{TOG/figs/latency_vs_quality/heat_quality2.pdf}\label{fig:optimization:heat:quality}}
    \subfloat[latency]{\includegraphics[width=0.48\linewidth]{TOG/figs/latency_vs_quality/heat_time2.pdf}\label{fig:optimization:heat:time}}

    \subfloat[quality-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/quality_only.pdf}\label{fig:optimization:quality}}\hspace{0.5em}
    \subfloat[latency-only]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/latency_only.pdf}\label{fig:optimization:time}}\hspace{0.5em}
    \subfloat[ours]{\includegraphics[width=0.3\linewidth]{TOG/figs/latency_vs_quality/our.pdf}\label{fig:optimization:our}}
    \Caption{Latency-quality joint-optimization.}
    {%
    \protect\subref{fig:optimization:heat:quality}/\protect\subref{fig:optimization:heat:time} shows \protect\Cref{eq:imageError}/latency(in millisecond under pytorch implementation) of an example foveal network.
    The values are computed with various settings of $\mlpLayerNum\times\mlpChannelNum$ (x-axis) and $\sphereNum$ (y-axis). The second row indicates a corresponding foveal image under different settings. Our method balances both perceptual quality and latency.\nothing{\qisun{(Jan 24) What is the unit of \subref{fig:optimization:heat:time}? It shouldn't be second, right?}\dnc{It's millisecond.}}%nothing
    }
    \label{fig:optimization}
\end{figure}

### 2021-01-28 TOG/fig_ray_bounding.tex
last day

\begin{figure}
    \centering
    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
    \Caption{Comparison of ray marching bounding.}
    {%
    }
    \label{fig:method:bounding}
\end{figure}

### 2021-01-28 TOG/fig_system.tex
last day

\begin{figure}[bht]
    \centering
    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_fovea.png}\label{fig:system:fovea}}\vspace{1em}
    \subfloat[mid-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_mid.png}\label{fig:system:mid}}\vspace{1em}
    \subfloat[far-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_periph.png}\label{fig:system:far}}

    \subfloat[rendering]{\includegraphics[width=0.9\linewidth]{TOG/figs/layer_blend/lobby_view0000_blendmask.png}\label{fig:system:blending}}
    \Caption{Synthesis output and rendering.}
    {%
    \subref{fig:system:fovea}/\subref{fig:system:mid}/\subref{fig:system:far} shows our individual gaze-contingent synthesis for fovea (within $20$deg)/mid-periphery (within $45$deg)/far-periphery(within $110$deg, the capability of our VR display), respectively. 
    \subref{fig:system:blending} illustrates our real-time rendering system by dynamically blending the individual images to the final displayed frame.
    }
    \label{fig:system}
\end{figure}

### 2021-01-28 TOG/teaser_fig.tex
snapshot
last day

@@
     % \includegraphics[height=4.5cm]{TOG/figs/scene.pdf}
-    \includegraphics[height=4.5cm]{TOG/figs/teaser_1a.pdf} %try other colors a little bit
+    \includegraphics[trim=100 0 120 0,clip,height=5.5cm]{TOG/figs/teaser/teaser_1a_wc.pdf}
   }%subfloat
-  \subfloat[latency]{
+  \subfloat[overall quality]{
     \label{fig:teaser:latency}
-    \includegraphics[height=4.5cm]{TOG/figs/teaser_1b_small.pdf}
-    % \includegraphics[height=4.5cm]{TOG/figs/teaser.pdf}
+    % \includegraphics[height=4.5cm]{TOG/figs/teaser_1b_update.pdf}
+    % text version
+    % \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_crop_zh_large.pdf}
+    % color version
+    \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_crop_color_s.pdf}
     % nerf: 15+s/per infer
@@
     \label{fig:teaser:quality}
-    \includegraphics[height=4.5cm]{TOG/figs/teaser_1c.pdf}
+    % \includegraphics[height=4.5cm]{TOG/figs/teaser_1c.pdf}
+    % text version
+    % \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_flat_zh_large.pdf}
+    % color version
+    \includegraphics[height=5.5cm]{TOG/figs/teaser/overall_flat_color_s.pdf}
   }%subfloat
- \Caption{Illustration of our system.}
+ \Caption{Illustration of our gaze-contingent neural synthesis method.}
  {% 
-\subref{fig:teaser:scene} shows visualizes our active-viewing tailored scene representation.
-Comparing with \protect\cite{mildenhall2020nerf}, \subref{fig:teaser:latency} shows our benefits of reduced latency.
-As a zoom-in, \subref{fig:teaser:quality} compares our foveal quality preservation.
-\qisun{Cross out half-half in  \subref{fig:teaser:latency}. TODO for myself: I should make the background scene semi-transparent. Will do so soon.}
+\subref{fig:teaser:scene} Visualization of our active-viewing-tailored egocentric neural scene representation.
+\subref{fig:teaser:latency} Our method allows for extremely low data storage of high quality 3D scene assets and achieves high perceptual image quality and low latency for interactive virtual reality.
+Here, we compare our method with the state-of-the-art alternative neural synthesis \protect\cite{mildenhall2020nerf} and foveated rendering solutions \cite{perry2002gaze,Patney:2016:TFR} and demonstrate that our method significantly reduces time consumption and data storage (more than $99\%$ reduction for both) for first-person immersive viewing.
+The orange circle indicates viewer's gaze position.
+We also show zoomed-in views of the foveal images generated by various methods in \subref{fig:teaser:quality} and show that our method is superior compared to existing methods in achieving high perceptual image fidelity.
+%\qisun{(Jan 24, 2021) Let's add some latency numbers to
+%\zh{15s/infer for NeRF}\subref{fig:teaser:latency}?}
+%\qisun{(Jan 25) Can I get your help on adding the numbers with the fancy font to the figures directly in \subref{fig:teaser:latency}?}
  }

### 2021-03-09 TOG/rebuttal.md
submission

# Grouped and Organized
## Shared
1. About hardware requirement (r1.1, r2.2):
   (r1) experiment is done on high-end machine but paper is targeting for mobile device?
   (r2) distance towards low latency, and low-storage VR/AR platforms?
2. About static scene (r1.2, r1.3, r3.5):
   (r1) no interaction, mobile devices can store
   (r3) scope should be clarified
3. About encoding performance (r3.5, r4.3):
   (r3) scope should be clarified
   (r4) viewing is instant, but creation is not (encoding/training is slow)
4. About user study (r1.4, r3.4, r4.6):
   (r1, r4) short, gaze fixed / static
   (r3) narrow foveation methods selection
5. About constraint (r1.6, r2.5, r2.6, r4.4, r4.7):
   (r1) restricted position and direction?
   (r2) low (angular) resolution // head volume? frequency of transmiting network?
   (r4) occlusion (motion parallel): not clearly demonstrated
6. Publish source code (r3.3, r5.1):
   (r5) to help understanding
7. About written (r3.1, r3.2, r4.14):
   (r3) improve clarity / bib consistency
   (r4) typo / style

## Individual
Review 1
* Foveation rendering as ground truth (r1.5)

Review 2
* Concern about the comparison with MSI ([Broxton2020] or [Attal2020]):
  MSI can achieve same low storage under constraint of no view-dependent effects, small motion range, low resolution (r2.4, r2.6, r2.7)
* Angular resolution: 6.4 PPD vs ground truth 13 PPD? (r2.1)
* Experiment about storage? (r2.3)

Review 3 (None)

Review 4
* not comparable between nerual and mesh representation (r4.2)
* A fairer comparison: allow NeRF an additional 96 channels (r4.8)
* Why our is better? foveat network: still represents high-res whole scene, why baseline is worse? (4.9)
* Artifacts in video:
  1m30: double image artifacts (r4.5)
  1m03: black splotches, where they come from? (r4.10)
* Minor questions:
  Misunderstanding: need depth reconstruction for 360 photos application. (r4.1)
  Not understande Fig 7 (r4.11)
  Isn't the representation just alpha composited front to back? (r4.12)
  Is NeRF here also running on a TensorRT implementation? (r4.13)

Review 5
* Hard to read (Not understand)
  "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions."
  --> how did this come about? 3 pairs of conditions x 2 scenes? (r5.2)
  "Each condition from an individual scene consists of 3 different gaze positions."
  --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped? (r5.3)
  "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image."
  --> What is meant by perceptual similarity? color similarity? (r5.4)
  "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together."
  --> What is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to? (r5.5)

# Raw Collections
## Review 1
1. Target mobile device but experiment on high-end machine?
2. Static scene: mobile devices can store. how about dynamic scene?
3. No interaction
4. user study: fix gaze, short
5. foveation rendering as ground truth
6. restricted position and direction?

## Review 2
1. Angular resolution: 6.4 PPD vs ground truth 13 PPD?
2. Performance: hardware? distance from low latency, and low-storage VR/AR platforms?
3. experiment about storage?
4. MSI: 10PPD - 10 MB -> 6PPD - 0.06MB?
5. head volume? frequency of transmiting network?
6. powerful computer but low resolution. MSI can achieve same storage real-time on mobile
7. why ray-caster? why not bake to MSI? (Under constraint of no view-dependent effects, small motion range, low resolution)

## Review 3
1. Improvement of clarity: the introduction of the egocentric coordinates might be a bit dense/slow to read
2. Reference: make bibliography more consistent
3. Publish source code
4. User study: the selection of foveation methods to compare is somewhat narrow
5. The scope should be clearly stated: static and receive side (decoder) only

## Review 4
1. Misunderstanding: 360 photos application: need depth reconstruction, e.g., Serrano et al. 2019 IEEE VR.
2. not comparable between nerual and mesh representation
3. Encode is slow, viewing is instant, but creation is not
4. Occlusion: small.
5. double image artifacts at 1m30s in video
6. Static user study: only a subset of experience
7. Translation and occlusion (motion parallax) are not clearly demonstrated and validated
8. A fairer comparison: allow NeRF an additional 96 channels
9. Why our is better? foveat network: still represents high-res whole scene, why baseline is worse?
10. Artifacts: 1m03 in video: black splotches, where they come from?
11. (Minor) Not understande Fig 7
12. (Minor) Isn't the representation just alpha composited front to back?
13. (Minor) Is NeRF here also running on a TensorRT implementation?
14. Typos/style:
    L154: "remarkably" <- can remove this; the reader can determine whether it is remarkable or not.
    L324: using 'd' for density here is not a huge problem but might be confused for 'depth'. Later on (L614) this is refered ot as 'a'.
    L1064: "unprecedented performance" <- please tone it down a bit; every technique is a trade-off somewhere.
    Duplicated citation of Attal et al.

## Review 5
1. Hard to read
2. "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions." --> how did this come about? 3 pairs of conditions x 2 scenes?
3. "Each condition from an individual scene consists of 3 different gaze positions." --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped?
4. "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image." What is meant by perceptual similarity? color similarity?
5. "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together." --> Unfortunately this is too concise an explanation to make sense -- what is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to?

### 2021-03-09 TOG/0-abstract.tex
submission

@@
 Tailored for the future portable, low-storage, and energy-efficient VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis method.
-We incorporate the human psychophysics of visual- and stereo-acuity into egocentric neural representation of 3D scenery.
-Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high quality immersive interaction.
+We incorporate the human psychophysics of visual- and stereo-acuity into an egocentric neural representation of 3D scenery.
+Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high-quality immersive interaction.
 Both objective analysis and subjective study demonstrate the effectiveness of our approach in significantly reducing local storage volume and synthesis latency (up to 99\% reduction in both data size and computational time), while simultaneously presenting high-fidelity rendering, with perceptual quality identical to that of fully locally stored and rendered high-quality imagery.

### 2021-03-09 TOG/1-introduction.tex
submission

@@
 }%note
-Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
-Thus, streaming services and remote cloud-based rendering (such as with Google Stadia and Nvidia Shield) have rapidly emerged in recent years.
+Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high-resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
+Thus, streaming services and remote cloud-based rendering (such as Google Stadia and Nvidia Shield) have rapidly emerged in recent years.
 These storage-friendly systems have shown an unprecedented potential to reshape the traditional computer graphics pipeline by completely relieving the end user's computational workload and storage pressure, thus advancing device portability.
@@
 
-However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles, due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}).
-In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for sickness-free and natural experience.
+However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}).
+In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for a sickness-free and natural experience.
 As a consequence, there is a demand for massive high-quality spatial-temporal data.
@@
 Recent years have witnessed a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets.
-Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf} and many more.
+Application of neural representations includes voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf}, and many more.
 However, existing solutions suffer from either high time consumption or low image fidelity with large scenes.
@@
 Acceleration without compromising perceptual quality is also an ultimate goal of real-time rendering.
-Gaze-contingent rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}.
+Gaze-contingent rendering approaches have shown remarkable effectiveness in presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}.
 This is achieved by harnessing high field-of-view(FoV) head-mounted displays and high precision eye-tracking technologies.
@@
 To overcome these open challenges mentioned above, we present the first gaze-contingent and interactive neural synthesis for high-quality 3D scenes.
-Aiming at the next-generation wearable, low-latency and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions.
-The predicted frames are perceptually identical to traditional real-time rendering of high-quality scenes.
-With wide FoV and high resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB to 0.5MB) thus enables instant viewing (without long loading/transmission waits) compared to traditional rendering pipelines; preserves visual quality, and improves responsiveness (from 9s to as low as 20 ms) compared with alternative view neural synthesis methods.
-Tailoring to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.
+Aiming at the next-generation wearable, low-latency, and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions.
+The predicted frames are perceptually identical to the traditional real-time rendering of high-quality scenes.
+With wide FoV and high-resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB mesh to 0.5MB neural model) compared with traditional rendering pipelines; it also remarkably improves responsiveness (from 9s to 20 ms) compared with alternative view neural synthesis methods.
+That is, we enable instant immersive viewing without long loading/transmission waits while preserving visual fidelity.
+Tailored to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.
 %\praneeth{---- somoene please check the text in the brackets and make sure it is ok :)}
@@
 
-To achieve immersive viewing with low storage, high visual quality as well as low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis.
+To achieve immersive viewing with low storage, high visual quality and low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis.
 %To this end, we build both the retina-matched neural models and VR-tailored representation coordinates, towards high visual quality as well as low systematic latency.
@@
 Specifically, we represent 3D scenes with concentric spherical coordinates.
-Our representation also allows for depicting the foveated color sensitivity and stereopsis, during both the training and inference phases.
+Our representation also allows for depicting the foveated color sensitivity and stereopsis during both the training and inference phases.
 Furthermore, we derive an analytical spatial-temporal perception model to optimize our neural scene representation towards imperceptible loss in image quality and latency.\qisun{better?
@@
 
-We validate our approach by conducting numerical analysis and user evaluations on commercially available AR/VR display devices.
+We validate our approach by conducting numerical analysis and user evaluation on commercially available AR/VR display devices.
 The series of experiments reveal our method's effectiveness in delivering an interactive VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, with significantly lowered data storage.
+We will open source our implementation and dataset.
 We make the following major contributions:

### 2021-03-09 TOG/2-prior.tex
submission

@@
 \subsection{Implicit Scene Representation for Neural Rendering}
-To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attentions in both computer graphics and computer vision communities. 
-With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf}.
-Prior arts also investigated on implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects.
-Inspired by \cite{park2019deepsdf}, signed distance function has been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}.
-The input information was varied from 2D images ~\cite{lin2020sdf, yariv2020multiview}, 3D context ~\cite{saito2019pifu, oechsle2019texture}, time-varying vector field~\cite{niemeyer2019occupancy}, and local features~\cite{tretschk2020patchnets, liu2020neural}.
-Current implicit representations primarily focus on locally ``out-side-in'' viewing of individual objects, with low field-of-view, speed and resolution.
-For large scenes, the lack of coverage may cause quality drop.
+To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attention in computer graphics and computer vision communities. 
+With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf,park2017transformation}.
+Prior arts also investigated implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects.
+Inspired by \cite{park2019deepsdf}, signed distance functions have been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}.
+The input information ranged from 2D images ~\cite{lin2020sdf, yariv2020multiview,choi2019extreme}, 3D context ~\cite{saito2019pifu, oechsle2019texture, zhang2020deep}, time-varying vector field~\cite{niemeyer2019occupancy} to local features~\cite{tretschk2020patchnets, liu2020neural}.
+Current implicit representations primarily focus on locally ``outside-in'' viewing of individual objects, with low field-of-view, speed, and resolution.
+For large scenes, the lack of coverage may cause the quality to drop.
 However, in virtual reality, the cameras are typically first-person and highly dynamic, requiring low latency, high resolution, and 6DoF coverage. 
@@
 %On the other hand, most view synthesizing methods are designed for offline viewing or processing, other than real-time or interactive consumption, especially with high resolution and FoV.  
-Our method bridges human vision to deep synthesis.
+Our method bridges human vision and deep synthesis.
 It accelerates neural synthesis beyond local rendering, enabling both computational and storage efficiency.
-It is suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
+It is also suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
 %\zh{seems not sufficient?}
@@
 \subsection{Panorama-Based 6DoF Immersive Viewing}
-Panorama images and videos greatly fit VR applications thanks to their 360 FoV coverage.
+Panorama images and videos much fit VR applications thanks to their 360 FoV coverage.
 However, the main challenge is their single projection center in each frame, limiting free-form camera translation, thus 6DoF immersive viewing experience.
@@
 As a spherical image, panoramas are designed for capturing physical worlds with flexible reprojections to displays.
-The 6DoF viewing typically suffer from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
+The 6DoF viewing typically suffers from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
 With an orthogonal mission, our model represents and reproduces arbitrary 3D virtual worlds for low storage VR.

### 2021-03-09 TOG/3-method.tex
submission

--- 2021-01-28 TOG/4-method.tex
@@
 \label{sec:method}
-In this section, we present our perception-guided neural scene synthesis for virtual reality.
-Based on a concentric spherical representation (\Cref{sec:method:representation}), our system predominantly comprises of two main steps: visual- and stereo- acuity adaptive view synthesis (\Cref{sec:synthesis}), and image-based real-time rendering to composite stereo display frames (\Cref{sec:method:blending}).
-We then craft an analytical spatial-temporal model to guide the optimization of our intra-system variables for desired precision-performance balance (\Cref{sec:method:optimization}).
+\nothing{In this section, we present our perception-guided neural scene synthesis for virtual reality.
+\zh{As a real-time overview: we first predict the the color as well as corresponding density for each pixel in the fovea, mid- and far- periphery (or designed 3D point?) given the camera position.
+Next we apply ray marching to generate the elemental image.
+Afterwards, we blend the results from each part and tune the contrast for each eye.}\qisun{merged with the following sentences.
+see if happy}} Based on a concentric spherical representation and trained network predicting RGBDs on the spheres (\Cref{sec:method:representation}), our system predominantly comprises two main steps in real-time: visual-/stereo-acuity adaptive elemental images synthesis with ray marching for fovea, mid-, and far- peripheries; followed by image-based rendering to composite displayed frames (\Cref{sec:synthesis}).
+For desired precision-performance balance, we further craft an analytical spatial-temporal model to optimize the determination of our intra-system variables, including coordinates and neural-network parameters (\Cref{sec:method:optimization}).
 \input{TOG/fig_notations}
@@
 However, immersive VR environments introduce open challenges for such parameterization due to the rapid variation of the egocentric (first-person) viewing (e.g., \Cref{fig:teaser:scene}).
-As a consequence, the neural representation on large virtual environments typically suffer from ghosting artifacts, low resolution, or slow speed (e.g., the right triangle in \Cref{fig:teaser:quality}).
+As a consequence, the neural representation on large virtual environments typically suffer from ghosting artifacts, low resolution, or slow speed (e.g., the white cube and red wall are blurred in \Cref{fig:teaser:latency} and \Cref{fig:teaser:quality}).
 %We tailor our method specifically for egocentric (first-person) views in immersive VR environments.
@@
 \paragraph{Egocentric coordinates}
-To tackle this problem we are inspired by the recent panorama dis-occlusion methods \cite{Lin:DeepPanorama,Benjamin:2020:RTV,Broxton:immersiveLF} to depict the rapidly varying first-person views with concentric spheres.
+To tackle this problem, we are inspired by the recent panorama dis-occlusion methods \cite{Lin:DeepPanorama,Benjamin:2020:RTV,Broxton:immersiveLF} to depict the rapidly varying first-person views with concentric spherical coordinates.
 This representation enables robust rendering at real-time rates and allows for 6DoF interaction to navigate complex immersive environments.
@@
 What do you think?}\dnc{Then perhaps we can use $\hat{q}$ as shown in \Cref{fig:notations} I think the alphabet 'q' means it's in spherical system, 'p' never exists anywhere else}
+%\zh{considering denoting $\sphereNum$ and $\mathbf{\sphereRadius}=\{\sphereRadius_i\},i\in [1,\sphereNum]$ in \Cref{fig:notations}}
 Here, $(r,g,b)$ and $d$ are the predicted color and density, respectively. 
@@
 %\dnc{Perhaps we can merge this paragraph with above section? And move the last paragraph of above section to here as a intro. Then this section is more focusing on GAZE-AWARE}
-For each scene, we approximate the function $\mlpFunc$ by training a multilayer perceptron (MLP) network, similar to \cite{mildenhall2020nerf}.
+For each scene, we approximate the function $\mlpFunc$ by a multilayer perceptron (MLP) network, similar to \cite{mildenhall2020nerf}.
 The network is composed of $\mlpLayerNum$ fully-connected layers of $\mlpChannelNum$ channels, each using ReLu as the activation function.
-We will further determine $\mlpLayerNum$, $\mlpChannelNum$ and $\sphereNum$ via a spatial-temporal optimization in \Cref{sec:method:optimization}.
-Once trained, the network achieves 4D prediction on the spherical voxels, on which a ray marching rendering is performed to synthesize the final frames for display.% based on users' frame-wise views. 
+We further determine $\mlpLayerNum$, $\mlpChannelNum$ and $\sphereNum$ via a spatial-temporal optimization in \Cref{sec:method:optimization}.
+Given rays of pixels in an image, the network predicts the colors and densities along these rays on every sphere.
+The ray marching weighted integrates predictions in back-to-front order to obtain actual colors of corresponding pixels.% based on users' frame-wise views.
 Similar to \cite{mildenhall2020nerf}, we apply a trigonometric input encoding and compute the L2-distance between the volume-rendered and mesh-rendered images as the training loss function.
-We discuss the specifics of training data creation and implementations in \Cref{sec:impl}.
+We discuss the specifics of training data creation and implementations in \Cref{sec:impl}.\nothing{\zh{Considering removing the last sentence.
+I think it is clear that this paragraph is not about details, guess reviewers are not expected to see the details.
+Or we should lead them not to expect details here.}}
 
-\subsection{Gaze-Aware Acceleration}
+\subsection{Gaze-Aware Synthesis}
 \label{sec:synthesis}
@@
 %the foveated rendering manner of VR display through training two orthogonal neural networks that are tailored for the perceptual acuity variances.
-Our concentric spherical coordinate system, as described in \Cref{sec:method:representation}, addresses the view variance problem in VR.
-However, it sill suffers from significant rendering latency (seconds per frame).
+%\zh{shall we explicitly name the representation with exactly `concentric spherical coordinate system' earlier?}
+Our concentric spherical coordinate, as described in \Cref{sec:method:representation}, addresses the view variance problem in VR.
+However, it still suffers from significant rendering latency (several seconds per frame).
 This is one of the core challenges causing neural representation to be unsuitable yet for VR.
-To combat the common performance problem of view synthesis, we predict multiple elemental images based on the viewer's real-time head and gaze motions\nothing{ instead of synthesizing a globally uniform image}.
-The elemental images are tailored for the visual acuity variances.
-We detail our network design and elemental image synthesis in the following section.
-We enable real-time responsiveness via introducing gaze-adaptive synthesis mechanisms.
+Instead of the typical single image prediction, we synthesize multiple elemental images to enable real-time responsiveness.
+The elemental images are generated based on the viewer's head and gaze motions and are adapted to the retinal acuity in resolution and stereo.
+%We detail our network design and elemental image synthesis in the following section.
+%We enable real-time responsiveness via introducing gaze-adaptive synthesis mechanisms.
 \input{fig_system}
-\paragraph{Adaptive visual acuity} 
-%the user's current view $[\rayo,\camDir]$ and gaze position $\gazeDir$ that are tracked in real-time by a VR headset.
-Existing neural scene representation networks predict uniform resolution images for offline third-person viewing.
-Due to the trade-off between quality and latency during inference, they are typically executed with low quality and/or high latency when it comes large field-of-view and high resolution imagery.
-\praneeth{this line is confusing}
-Our method, without compromising performance, generates {\it perceptually} high quality frames matching the retinal acuity, i.e. how the retina perceives dynamic and stereo frames.
-% Note that foveal/peripheral retinal images requires high/low pixel-wise precision but low/high FoV (number of pixels).
-Thus, we train two orthogonal neural networks that are tailored for the perceptual acuity variances.
-Because of foveation of imagery by the human visual system \cite{Guenter:2012:F3G}, we train two orthogonal networks (i.e., $\mlpFunc$s) to synthesize foveal ($\imageFoveal(\rayo,\camDir,\gazeDir)$, $0-20$deg), and peripheral elemental images (i.e., two $\mlpFunc$s to deploy for ray marching) for each eye.
-The periphery includes mid- ($\imageMid(\rayo,\camDir,\gazeDir)$, $0-45$deg) and high- eccentricity visual fields ($\imageFar(\rayo,\camDir$), $0-110$deg), see \Cref{fig:system}.
+\subsubsection{Adaptive visual acuity}
+% The human vision is foveated.
+With the high field-of-view VR displays, we exploit the spatially-adaptive visual acuity to significantly accelerate the computation without compromising {\it perceptual} quality.
+The human vision is foveated.
+Inspired by this, we significantly accelerate the computation by integrating the characteristic of spatially-adaptive visual acuity without compromising {\it perceptual} quality.
+
+Specifically, given device-tracked camera position ($\rayo$), direction ($\camDir$), and gaze position ($\gazeDir$),
+we train two orthogonal networks (i.e., function $\mlpFunc$s) to synthesize foveal ($\imageFoveal(\rayo,\camDir,\gazeDir)$, $0-20$ deg), and peripheral elemental images (i.e., two $\mlpFunc$s to deploy for ray marching) for each eye.
+The periphery includes mid- ($\imageMid(\rayo,\camDir,\gazeDir)$, $0-45$ deg) and high- eccentricity visual fields ($\imageFar(\rayo,\camDir$), $0-110$ deg), see \Cref{fig:system}.
 Note that $\imageFar$ is independent from the gaze direction $\gazeDir$. 
 %The images are defined as functions that maps individual camera rays ($\rayo,\rayd$) to an RGB color.
-\dnc{Could move to section 4?
-I think it's implementation related}\qisun{I think the PDD numbers give readers a rough impression of the varied quality across each of the image, right?} Specifically, to incorporate the display hardware properties and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $128^2$/$256^2$/$230\times256$ resolutions respectively.
-That is, the $\imageFoveal$ has the highest spatial resolution of $6.4$ pixels per degree (PDD), higher than those of $\imageMid$ ($5.7$ PDD) and $\imageFar$ ($2.33$ PDD).
+Specifically, to incorporate the display capabilities and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $128^2$/$256^2$/$230\times256$ resolutions respectively.
+That is, the $\imageFoveal$ has the highest spatial resolution of $6.4$ pixels per degree (PPD), higher than those of $\imageMid$ ($5.7$ PPD) and $\imageFar$ ($2.33$ PPD).
 %In each frame, the two networks independently return three elemental images, as seen from the separation range in deg, gradually larger areas along eccentricity.
 An example is shown in \Cref{fig:system}.
+% ZH: add conclusive or ending sentences here.
+%In each frame, the two networks independently return three elemental images, as seen from the separation range in deg, gradually larger areas along eccentricity.
+An example is shown in \Cref{fig:system}.
+% ZH: add conclusion or ending sentences here.
 
-\paragraph{Adaptive stereo-acuity}
-{
-To be rendered on head-mounted displays, stereo elemental images $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$ ($l$ and $r$ represent left and right eyes respectively) are required.
-This doubles the computational cost that is critical for latency- and frame-rate- sensitive VR experience
+\subsubsection{Adaptive stereo-acuity}
+Head-mounted VR displays require stereo rendering, thus the elemental images $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$ for each ($l$eft and $r$ight) eye.
+This, however, doubles the inference computation cost that is critical for latency- and frame-rate- sensitive VR experience
 \nothing{The separation of the three elemental images considers the varied visual acuity.
@@
 Whereas, because $\imageMid^{\{l,r\}}$ and $\imageFar^{\{l,r\}}$ demand high spatial resolution due to their large eccentricity coverage, }
-}
-(please refer to \Cref{sec:study:intra} for detailed comparisons).
+(please refer to \Cref{sec:study:intra} for breakdown comparisons).
 
-To tackle this problem towards real-time experience, we accommodate the inference process with the adaptive stereo-acuity in the periphery.
-Specifically, besides the visual acuity, psychophysical studies have revealed human's significantly declined stereopsis while receding from the gaze point \cite{mochizuki2012magnitude}.
-Inspired this characteristic, instead of inferring $6$ elemental images, we perform the computation with $\imageFoveal^{\{l,r\}}$, $\imageMid^m$, and $\imageFar^m$, where $m$ indicates the view at the midpoint of the left and right eyes.
-The individual images are further blended for display, as detailed in \Cref{sec:method:blending}.
-\Cref{fig:mono} visualizes the stereo changes from the adaptation.
+To further accelerate stereo rendering towards real-time experience, we accommodate the inference process with the adaptive stereo-acuity in the periphery.
+In fact, besides the visual acuity, psychophysical studies have also revealed human's significantly declined stereopsis while receding from the gaze point \cite{mochizuki2012magnitude}.
+Inspired by this characteristic, we perform the computation with $\imageFoveal^{\{l,r\}}$, $\imageMid^c$, and $\imageFar^c$ instead of inferring $6$ elemental images, where $c$ indicates the view at the central eye (the midpoint of the left and right eyes).
+\Cref{fig:mono} visualizes the stereopsis changes from the adaptation using an anaglyph.
 \input{TOG/fig_mono_comparison}
-
-\subsection{Real-Time Frame Composition}\dnc{May be moved to section 4 (merge with the paragraph 'Integration').
+\subsubsection{Real-time frame composition}%\dnc{May be moved to section 4 (merge with the paragraph 'Integration').
 I think it's not closely related to our methodology and we have no key contributions about that.
@@
 Intentionally kept it short according to contribution level here.}
-\label{sec:method:blending}
-In real-time, we perform a image-based rendering that blends the elemental images. The output frames are rendered on the stereo VR displays.
-To accommodate the uniformized $\imageMid^m$, we shift it with $-/+ \frac{\rayo^r-\rayo^l}{2}$ for left/right eye respectively. 
-\dnc{Not accurate. shifting is related to mean of fovea depthmap inferred in actual implementation to minimize the misalignment between fovea and mid}
+%\label{sec:method:blending}
+%The individual elemental images are further blended as a single frame via a real-time image-based rendering. 
+\input{TOG/fig_blending}
+With the obtained elemental images as input, an image-based rendering in the fragment shader is then executed to generate final frames for each eye (\Cref{fig:system:blending}).
+%In real-time, we perform a image-based rendering that blends the elemental images. 
+The output frames are displayed on the stereo VR HMDs.
 {
-With the obtained elemental images as input, a fragment shader is executed to generate final images for each eye.
-Two adjunct layers are blended using a smoothstep function across $40\%$ of inner layer.
-\nothing{After obtaining the individual images, our fragment shader retrieves the three frames and applied a pixel wise bilinear interpolation at each $10\%$ connecting areas between two adjunct layers.}
+Two adjunct layers are blended using a smooth-step function across $40\%$ of the inner layer.
+This enhances visual consistency on the edges between layers \cite{Guenter:2012:F3G}.
 }
-This enhances visual consistency on the edges between layers \cite{Guenter:2012:F3G}.
-Moreover, to further preserve peripheral elemental images' visual fidelity due to its low PDD, we enhance the contrast following the mechanism of \cite{Patney:2016:TFR}, the details are visualized in \Cref{fig:method:blending}.
+To accommodate the mono-view $\imageMid^c$ and $\imageFar^c$, they are shifted towards each eye according to approximated foveal depth range.
+%\nothing{we shift it with $-/+ \frac{\rayo^r-\rayo^l}{2}$ for left/right eye respectively. }
+Lastly, we enhance the contrast following the mechanism of \cite{Patney:2016:TFR} to further preserve peripheral elemental images' visual fidelity due to its low PDD.
+The details are visualized in \Cref{fig:method:blending}.
 %\dnc{Note: the contrast enhancement is performed on elemental images, before blending.
@@
 
-\begin{figure*}[htb]
-    \centering
-    \subfloat[network overlay]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/overlap2.pdf}\label{fig:method:blending:overlay}}\hspace{1em}
-    \subfloat[blending]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended2.pdf}\label{fig:method:blending:blending}}\hspace{1em}
-    \subfloat[contrast enhance]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
-    \Caption{Our multi-layer real-time rendering system.}
-    {%
-    \subref{fig:method:blending:overlay} shows the original 3-layer image output overlaid.
-    \subref{fig:method:blending:blending} shows the result with cross-layer blending that enables smooth appearance on the edges.
-    \subref{fig:method:blending:contrast} shows the contrast-enhanced rendering pass in the periphery.
-    }
-    \label{fig:method:blending}
-\end{figure*}
-
 \subsection{Latency-Quality Joint Optimization}
 \label{sec:method:optimization}
-As a view synthesis system based on sparse egocentric representation (the density of $\sphereNum$ spheres) and neural networks (the $\mlpLayerNum\times\mlpChannelNum$), the method inevitably introduces approximation errors.
+As a view synthesis system based on sparse egocentric representation (the $\sphereNum$ spheres) and neural networks (the $\mlpLayerNum, \mlpChannelNum$), the method inevitably introduces approximation errors.
 On the other hand, these variables also determine the online computational time that involves inferring function $\mlpFunc$ and ray marching the $\sphereNum$ spheres.
-However, VR strictly demands both quality and performance. We present a spatial-temporal model that analytically depicts the correlations and optimizes the variables for ideal user experience.
+However, VR strictly demands both quality and performance. We present a spatial-temporal model that analytically depicts the correlations and optimizes the variables for an ideal user experience.
 
@@
 
-Similar to volume-based representation, the multi-spherical system is also defined in discrete domain.
+Similar to volume-based representation, the multi-spherical system is also defined in the discrete domain.
 The sparsity thus naturally introduces approximation error that compromises the synthesis quality.
@@
 \end{equation}
-where $\rayo$ and $\rayd$ are the ray's origin point and normalized direction respectively.
+where $\rayo$ and $\rayd$ are the ray's origin point and normalized direction, respectively.
 % \begin{equation}
@@
 % \end{equation}
-Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt$ represented in spherical coordinates, the ray connect them is with the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$.
-This ray may intersect with more than one spheres.
+Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt$, the ray connecting them has the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$.
+This ray may intersect with more than one sphere.
 Among them, the closest one to $\SpatialPt$ is:
@@
 \end{equation}
-In the 3D space, the offset distance $\norm{\pt^\prime-\hat{\pt}}$ indicates the precision loss at $\pt$ from the representation. By integrating over all view and scene points, we obtain:
+In the 3D space, the offset distance $\norm{\pt^\prime-\hat{\pt}}$ indicates the precision loss at $\pt$ from the representation. By integrating over all views and scene points, we obtain:
 \begin{equation}
@@
 \end{equation}
-$\sparseError$ depicts how the generic representation precision of a scene, given a coordinate system defined by $\sphereNum$ and $\mathbf{\sphereRadius}$.
+By integrating all 3D vertices $\SpatialPt$ and camera positions $\rayo$ in our dataset sampling, $\sparseError$ depicts how the generic representation precision of a scene, given a coordinate system defined by $\sphereNum$ and $\mathbf{\sphereRadius}$.
 
-In comparison, a neural representation aims at predicting projected image given a camera position $\rayo$ and orientation $\camDir$.
+In comparison, a neural representation aims at predicting projected image given a $\rayo$ and $\camDir$.
 Thus, we further extend \Cref{eq:sparseError} to image space to analyze the error given a set of camera's projection matrix  $\projectionMatrix(\rayo,\camDir)$ as
@@
 However, for neural scene representation, infinitely enlarging $\sphereNum$ would significantly raise the challenges in training precision and inference performance.
-Likewise, increasing prediction and rendering densities (i.e., higher $\sphereNum$ or $\mlpLayerNum$/$\mlpChannelNum$) naturally improves the image output quality for lower \Cref{eq:imageError} while significantly increases the computation during ray marching, causing quality drop stretched along time.
+Likewise, increasing representation densities (i.e., higher $\sphereNum$) and/or network complexities (i.e., higher $\mlpLayerNum$/$\mlpChannelNum$) naturally improves the image output quality (lower \Cref{eq:imageError}).
+However, this significantly increases the computation during ray marching, causing quality drop stretched along time.
 In the performance-sensitive VR scenario, the latency breaks the continuous viewing experiment and may cause simulator sickness.
-Thus, we determine the coordinate variables with a content-aware optimization towards optimal quality-speed performance.
+Thus, with content-aware optimization, we further optimize the system towards an ideal quality-speed balance.%\zh{did not specify coordinate variables before.
+As well as consistent with the following ``optimal coordinate system ($\sphereNum$ and $\mathbf{\sphereRadius}$) and MLP settings ($\mlpLayerNum$/$\mlpChannelNum$)''}
 
 \paragraph{Spatial-temporal modeling}
-Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform a spatial-temporal joint modeling to determine the optimal coordinate system ($\sphereNum$ and $\mathbf{\sphereRadius}$) for positional precision and MLP settings ($\mlpLayerNum$/$\mlpChannelNum$) for color precision that adapt to individual computational resources and scene content.
-This is achieved via a latency-precision modeling in the spatial-temporal domain.
+\input{TOG/fig_optimization}
+Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform spatial-temporal joint modeling to determine the optimal coordinate system ($\sphereNum$) for positional precision and network complexity ($\mlpLayerNum$, $\mlpChannelNum$) for color precision that adapt to individual computational resources and scene content.
+This is achieved via latency-precision modeling in the spatial-temporal domain:
 %Specifically, we model the prediction image $P$ with gaze position ($G$) at time $t$ as $P_G(t)$ and the ground truth retinal image (from a simulated local rendering) as $\hat{P}_G(t)$.
@@
 where $\latency\triangleq\latency(\sphereNum, \mlpLayerNum, \mlpChannelNum)$ is the system latency with a given coordinate and network setting.
-$\mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)$ is the L1-distance between the four channels ($r$,$g$,$b$,$a$) and0 the highest values of network layer settings.
-For simplicity, we assumed uniformly distributed $\mathbf{\sphereRadius}$ with a fixed range of the spherical coverage.
+$\mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)$ is the four ($r,g,b,a$) output channels' L1-distance between a given network setting and the highest values $\sphereNum=64,\mlpLayerNum=8,\mlpChannelNum=256$).
+For simplicity, we assumed uniformly distributed $\mathbf{\sphereRadius}$ with a fixed range of the spherical coverage.%\zh{``and0 the highest values'' ??}
 
-Suggested by Albert et al.
+As suggested by Albert et al.
 \shortcite{albert2017latency}, the latency for a foveated system shall reach below \textasciitilde$50$ms for undetectable artifacts.
@@
     &\argmin_{\sphereNum, \mlpLayerNum, \mlpChannelNum} \finalError(\sphereNum, \mlpLayerNum, \mlpChannelNum), \\
-    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L_0
+    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L_0.
 \end{aligned}
 \end{equation}
-\Cref{fig:optimization} visualizes an example of the optimization mechanism for in fovea.
-The optimized results ($\sphereNum, \mlpLayerNum, \mlpChannelNum$) for individual networks are detailed in the next section.
+\Cref{fig:optimization} visualizes an example of the optimization mechanism for an foveal image $\imageFoveal$.
+The optimized results ($\sphereNum, \mlpLayerNum, \mlpChannelNum$) for individual networks are used for training.
+The optimization outcomes are detailed in \Cref{sec:impl}.
 The visual quality is validated by psychophysical study (\Cref{sec:study:user}) and objective analysis (\Cref{sec:study:quality}).
 The latency breakdown of our system is reported in \Cref{sec:study:intra}.
-\input{TOG/fig_optimization}
-\input{TOG/4.5-implementation}

### 2021-03-09 TOG/4-implementation.tex
submission

\section{Implementation}\label{sec:impl}
In this section, we elaborate on how we collect data for training, the specific parameters we choose for the two orthogonal neural networks, and the software/hardware environment we use.

\paragraph{Data generation}
For each scene, we generate two datasets separately for training foveal and periphery networks. The foveal dataset is composed of $256\times256$ images rendered at different views with 20 deg field-of-view, while the periphery dataset uses 45 deg field-of-view to rendering. The view translations ($\rayo$) are sampled every $5$ cm on each axis. The rotations ($\camDir$) are sampled so that adjacent samples are seamlessly connected with no overlap.
% \nothing{
% %\paragraph{Input encoding}
% Similar to \cite{mildenhall2020nerf}, we also extend the input dimensions ($d=3$) to a higher dimensional ($d=60$) for training. 
% \nothing{This is achieved via coding the input as:
% \begin{align}
% \iota (p) = (\sin(2^0 p), \cos(2^0 p), ..., \sin(2^{L-1}p, \cos(2^{L-1}p))
% \label{eqn:inputencode}
% \end{align}
% The function $\iota (\cdot)$ is applied to dimension of $\SpatialPt=(r_\SpatialPt,\theta_\SpatialPt, \phi_\SpatialPt)$.
% In our experiment we choose $L=10$.\qisun{What is $L$?}\dnc{It's an encoding parameter for input. maps 3D input to $R^{6L}$ space, as \autoref{eqn:inputencode} shows}
% }}
\paragraph{Optimized network parameters}
Guided by the latency-quality joint optimization described in \Cref{sec:method:optimization}, we implemented the foveal network with $\mlpLayerNum=4$, $\mlpChannelNum=128$ and $\sphereNum = 32$. For the periphery neural network, the values are $\mlpLayerNum=4$, $\mlpChannelNum=96$ and $\sphereNum = 16$. These optimized parameters achieve a proper balance between quality and latency, as shown in \Cref{fig:results:comparison} and \Cref{tbl:ablation}.

\paragraph{Ray marching bounding}
We bound the ray marching within region that $\mathbf{\sphereRadius}$ are sampled in, thus has impact on $\sparseError$ in \Cref{eq:sparseError}. As shown in \Cref{fig:impl:bounding:off}, results with infinite upper bounding ($\sphereRadius\in[\sphereRadius_{near},+\infty)$ incur blurry and ghosting artifacts due to larger $\sparseError$. In contrast, limited bounding significantly improves the quality (\Cref{fig:impl:bounding:on}). %The bounding range, together with $\sphereNum$, $\mlpLayerNum$ and $\mlpChannelNum$, are determined by a spatial-temporal optimization, as detailed in \Cref{sec:method:optimization}.
\input{TOG/fig_ray_bounding}
\input{TOG/fig_tile_results}
\paragraph{Environment}
The system was implemented through OpenGL framework using CUDA and accelerated by TensorRT with Intel(R) Xeon(R) Gold 6230R CPU @ 2.10GHz (256GB RAM) and one NVIDIA GTX 3090 graphics card.
For each scene, we trained around 200 epochs for experiment use (about a day).

### 2021-03-09 TOG/5-result.tex
submission

@@
 \label{sec:result}
-\input{TOG/fig_tile_results}
-With a variety of scenes as shown in \Cref{fig:results:comparison}, we first conduct subjective studies (\Cref{sec:study:user}) and objective measurements (\Cref{sec:study:quality}) to evaluate the perceptual quality.
-Further, we analyze the intra-system workings and compare inter-system performance in \Cref{sec:study:intra}.
+% \input{TOG/fig_tile_results}
+With a variety of scenes as shown in \Cref{fig:results:comparison}, we conduct a subjective study (\Cref{sec:study:user}), and calculate objective measurements (\Cref{sec:study:quality}) to evaluate the perceptual quality.
+Further, we analyze the intra- and inter-system performance in \Cref{sec:study:intra}.
 
@@
 \label{sec:study:user}
-% \zh{(Jan 10) all the Qs are answered offline, will update the section soon.}
 % In this section, we report two perceptual experiments to investigate 1) how users perceive our solution and the ground truth and 2) how our solutions behaves compared to other alternatives.
-We first conducted a psychophysical experiment to investigate how users perceive our solution ({\bf OURS}) compared with existing neural view synthesis methods (\cite{mildenhall2020nerf}, {\bf NeRF}) and foveated rendering with the full mesh data (\cite{perry2002gaze}, {\bf F-GT}).
+We conducted a psychophysical experiment to investigate how users perceive our solution ({\bf OURS}) compared with existing neural view synthesis methods (\cite{mildenhall2020nerf}, {\bf NeRF}) and foveation rendering of the original mesh scene (\cite{perry2002gaze}, {\bf F-GT}).
 
@@
 \paragraph{Stimuli}
-For precise comparison and accommodating the low frame rate of alternative solutions (as to be compared in \Cref{sec:study:intra}), each group of stimuli were static images rendered via {\bf F-GT} or synthesized via  {\bf OURS}/{\bf NeRF} with the same views, as shown in \Cref{fig:results:comparison}.
+For precise comparison and accommodating the low frame rate of alternative solutions (as to be compared in \Cref{sec:study:intra}), each group of stimuli consisted of static stereo images rendered via {\bf F-GT} or synthesized via {\bf OURS}/{\bf NeRF} with the same views, as shown in \Cref{fig:results:comparison}.
 They were generated with the same and randomly defined gaze fixation across conditions.
@@
 The foveal kernel size in {\bf F-GT} was designed to match our display size and eye-display distances.
-For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $N_{sample}=16$, $N_{layers}=4$, $N_{channel}=128$.
-The bottom right of each scene figure \Cref{fig:results:comparison} shows the resulting images.
+For generating {\bf NeRF} condition, we retrained the model from \cite{mildenhall2020nerf} on our dataset with $\mlpLayerNum=4$, $\mlpChannelNum=128$, and $\sphereNum = 16$.
+The bottom right of each scene figure \Cref{fig:results:comparison} shows the resulting {\bf NeRF} images.
 %We predict the image from the camera transformation, fine-tune the foveated layer, and render the image by alpha blending the three layers.
@@
 \paragraph{Setup}
-Each participant wore an eye-tracked HTC Vive Pro Eye headset and remained seated to example the stimuli during the experiment.
+Each participant wore an eye-tracked HTC Vive Pro Eye headset and remained seated to examine the stimuli during the experiment.
 Ten users participated in and completed the study (3 females, $M=23.7$, $SD=1.49$). 
 % One participant (age=23, female) did not complete the experiment due to constant gaze drifting (please refer to the task description below).
-None of the subjects were aware of the research, the experimental hypothesis, nor the number of methods. All participants had a normal or corrected-to-normal vision.
+No participants were aware of the research, the experimental hypothesis, nor the number of methods. All participants had a normal or corrected-to-normal vision.
 
@@
 To prevent the fixation from shifting away and ensure accuracy, we tracked the users' gaze throughout the experiment.
-Whenever the gaze is more than $5$deg away from the target, a trial was dropped immediately with a black screen informing the participant.
+Whenever the gaze is more than $5$ deg away from the target, a trial was dropped immediately with a black screen informing the participant.
 After each trial, the participants were instructed to select which of the two stimuli appeared with higher visual quality using a keyboard.
 Before each experiment, a warm-up session with 6 trials was provided to familiarize the participants with the study procedure.
-The orders of conditions among trials were randomized with counterbalancing.
-The entire experiment of each user consists of $36$ trials, $12$ trial per pair of conditions.
+The orders of conditions among trials were randomized and counter-balanced.
+The entire experiment of each user consists of $36$ trials, $6$ trials per pair of ordered conditions.
+% \zh{Regarding ``$12$ trial per pair of conditions'', here you treat pair(a-b) and pair(b-a) the same?}\qisun{I did this before knowing your plot on treating them differently, which was better.
+Tweaked.}
 %To start with, each participant was informed that they were required to choose an alternative with higher quality after watching each pair of stimuli.
@@
 After each pair, participants reported their opinions on which image has higher quality.
-To minimize the effect of accumulated fatigue, we enforced breaks at least a 2-second between trials and 60-second after each scene.
-Meanwhile, the participants were able to take as much time as needed.
+To minimize the effect of accumulated fatigue, we also enforced breaks between trials (at least 2 seconds) and after each scene (60 seconds).
+Meanwhile, the participants were allowed to take as much time as needed.
 
@@
 \Cref{fig:results:2afc} plots the results considering both methods and their orders in the 2AFC experiments.
-Among all three conditions, we observed close-to-random-guess among trails that compares {\bf F-GT} and {\bf OURS} ($53.3\%$ voted for {\bf OURS}, $SD=14.9\%$).
+Here we analyzed and reported the results regardless of the orders.
+Among all three conditions, we observed close-to-random-guess among trials that compared {\bf F-GT} and {\bf OURS} ($53.3\%$ voted for {\bf OURS}, $SD=14.9\%$).
 Meanwhile, a significantly higher ratio of voting {\bf F-GT} over {\bf NeRF} was observed ($92.5\%$, voted for {\bf F-GT}, $SD=11.4\%$, binomial test showed $p<0.005^{***}$).
@@
 Meanwhile, both of the two conditions showed significant quality preference than {\bf NeRF}.
-That is, with immersive, high FoV, and first-person viewing constraints, {\bf OURS} synthesizes gaze-contingent retinal image with superior perceptual quality than alternative synthesis solutions.
+That is, with immersive, high FoV, and first-person viewing constraints, {\bf OURS} synthesizes gaze-contingent retinal images with superior perceptual quality than alternative synthesis solutions.
 The latter typically aims to reconstruct the full image over the entire field-of-view.
@@
 Complementary to the subjective measurement (\Cref{sec:study:user}), we evaluate the perceived quality in an objective fashion.
-Specifically, under all four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT} and {\bf NeRF} at individual eccentricity ranges (up to $110$ deg, the capability of the VR HMD) via partial images.
+Specifically, under all four different scenes (indoor/outdoor, artificial/realistic), we compare {\bf OURS}, {\bf F-GT}, and {\bf NeRF} at individual eccentricity ranges (up to $110$ deg, the capability of the VR HMD) via partial images.
 For each eccentricity range, we compare the deep perceptual similarity (LPIPS) \cite{zhang2018unreasonable} across all scenes.
@@
 Smaller values indicate higher perceptual similarity.
-For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $420$ data per scene.
+For each scene, we sample $20$ views with gazes at the middle of the display, resulting in $20$ data per eccentricity value, $21$ eccentricity value per scene ($5$ deg step size), thus $420$ data per scene.
 We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together.
@@
 \Cref{fig:lpips} plots LPIPS values across all scenes and eccentricity ranges ($5$ deg step size). 
-From foveal to near eccentricity (<=$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.26$).
+From foveal to near eccentricity ($\leq40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\eta^2 = 0.26$).
 That is, \textbf{OURS} shows significantly lower LPIPS than \textbf{NeRF} ($p<.001^{***}$), although higher than \textbf{F-GT} ($p<.001^{***}$).
@@
 
-From mid- to far- periphery (>$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.16$).
+From near- to far- eccentricity (>$40$ deg), we observed significant effects of the stimuli on LPIPS with a ``large'' effect size ($\omega^2 = 0.16$).
 \textbf{OURS} shows significantly higher LPIPS than \textbf{NeRF}.
@@
 \paragraph{Discussion}
-In the fovea and near-periphery, the observation revealed our method's significantly higher perceptual quality than alternative neural representation solutions.
-This is evidenced by the significantly stronger perceptual similarity to \textbf{F-GT} by comparing between \textbf{OURS} and \textbf{NeRF}.
+In the foveal and near-periphery, the observation revealed our method's significantly higher visual quality than alternative neural representation solutions.
+This is evidenced by the significantly stronger perceptual similarity to \textbf{F-GT} by comparison between \textbf{OURS} and \textbf{NeRF}.
 The subjective study (\Cref{sec:study:user}) evidenced the marginally lower similarity than \textbf{F-GT} as statistically unnoticeable among users.
@@
 The latter has been shown to display identical perceptual similarity than full resolution rendering in VR \cite{Patney:2016:TFR}.
-Thus, the findings show that \textbf{OURS} doesn't compromise the peripheral vision's quality with its significantly enhanced synthesis acuity in fovea.
+Thus, the findings show that \textbf{OURS} doesn't compromise the peripheral vision's quality with its significantly enhanced synthesis acuity in the fovea.
 
 Note that the discoveries also agree with the observations from \Cref{sec:study:user}.
-That is, in addition to the significantly faster performance  ($99.7\%$ time reduction per frame as in \Cref{sec:study:intra}\nothing{\qisun{Check if I noted it right.}}), our method showed superior perceptual quality than \textbf{NeRF} under first-person, high resolution, and immersive viewing of 3D scenes.
+That is, in addition to the significantly faster performance  ($99.8\%$ time reduction per frame as in \Cref{sec:study:intra}\nothing{\qisun{Check if I noted it right.}}), our method showed superior perceptual quality than \textbf{NeRF} under first-person, high resolution, and immersive viewing of 3D scenes.
 Meanwhile, our synthesized views are perceptually identical to the traditionally rendered images with a complete 3D mesh.
-It yet achieves significant data storage-saving with only about $0.5$mb (up to $99.5\%$ in our experiment).
+It yet achieves significant data storage-saving with only about $0.5$MB neural model (up to $99.5\%$ saving from 3D mesh in our experiment).
 
@@
 % apply objective metric to |cref{fig:optimization}, aka different sample numbers
-Virtual and augmented reality demands high frame rates along with high quality to ensure immersive and comfortable experience.
+Virtual and augmented reality demands high frame rates along with high quality to ensure an immersive and comfortable experience.
 Our neural synthesis method introduces both spatial and angular (stereo) visual perception for real-time performance (\Cref{sec:method:representation}).
@@
 
-For high resolution ($1440\times1600$), high field-of-view ($110$deg) stereo images required by VR display, our system completes all computation (including gaze-contingent neural-inference and elemental images composition) in $31.8$ms per frame without stereo foveation, \dncR{as shown in \Cref{tbl:ablation} \st{shows a time consumption breakdown of our method, together with {\bf NeRF}}}.
-Compared with {\bf NeRF} \dncR{(the bottom row of \Cref{tbl:ablation})}, our spatial foveation \dncR{\st{for adaptive resolution}} significantly improves the system performance from offline computation (in seconds) to interactive speed (\dncR{about 30 frames per second\st{$31.8$ ms per frame}}).
-
-Tailored for VR displays, our \dncR{stereo foveation\st{adaptive stereo synthesis}} further reduces the computation time \dncR{and achieves high performance of more than 50 frames per second \st{by more than $10$ms/frame}}, contributing to a temporally continuous viewing experience without loss of perceived resolution and quality.
+For high resolution ($1440\times1600$), high field-of-view ($110$ deg) stereo images required by VR display, our system completes all computation (including gaze-contingent neural-inference and elemental images composition) in $31.8$ms per frame without stereo foveation, as shown in \Cref{tbl:ablation} .
+Compared with {\bf NeRF} (the bottom row of \Cref{tbl:ablation}), our spatial foveation significantly improves the system performance from offline computation (in seconds) to interactive speed (about 30FPS).
+Our stereo foveation further reduces the computation time and achieves a high performance of more than 50 frames per second, contributing to a temporally continuous viewing experience without loss of perceived resolution and quality.
 
 %Specifically, the significant performance gain benefits from the spatial and stereo foveation that do not compromise the visual quality.
-\begin{table}[!htb]
+\begin{table}[thb]
     \centering
@@
     {%
-    The numbers show the time consumption of each component and the overall system per frame. All units are in millisecond.
+    The numbers show the average time consumption of each component and the overall system per frame. All units are in millisecond.
     \nothing{\qisun{(Jan 25) I have update the time consumption of {\bf NeRF}. See if that is correct.}}

### 2021-03-09 TOG/6-conclusion.tex
submission

@@
 \label{sec:conclusion}
-\zh{consistently using high eccentricity rather far periphery?}
 In this paper, we present a gaze-aware neural scene representation and view synthesis method tailored for future portable virtual reality viewing experiences.
-Especially, we overcome the limitations of existing neural rendering approaches for high performance, resolution, and fidelity.
-Our network individually synthesizes foveal, mid- and far-periphery retinal images which are then blended to form a wide field of view image matching retinal acuity.
+Specifically, we overcome the limitations of existing neural rendering approaches for high performance, resolution, and fidelity.
+Our network individually synthesizes foveal, mid-, and far-periphery retinal images, which are then blended to form a wide field-of-view image matching retinal acuity.
 This is achieved via adapting the neural network model to the foveated visual acuity and stereopsis.
@@
 Our multi-spherical representation introduces higher quality when the VR camera is within the first sphere.
-As shown in \Cref{fig:large_translation}, when the virtual camera translates largely within a highly occluded scene, our method still synthesizes the accurate views but with a decline in foveal image quality.
+As shown in \Cref{fig:large_translation}, when the virtual camera translates largely within a highly occluded scene, our method still synthesizes the accurate views but with declined foveal image quality.
 This is due to the accumulated error of the concentric spherical representation (\Cref{eq:imageError}).
 Multiscale coordinates and networks that consider various level-of-details of the 3D space have shown their effectiveness of interpolating geometries \cite{winkler2010multi}.
-Developing a multiscale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, to extend the applicability of our method to large scale scenes with strong occlusions.
+Developing a multiscale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, extending the applicability of our method to large-scale scenes with strong occlusions.
 \begin{figure}[!ht]
     \centering
-    \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.pdf}\label{fig:large_translation:small}}\hspace{1em} % or 2_gas_0.3.png
-    \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.pdf}\label{fig:large_translation:large}} % or 2_gas_0.9.png
+    % \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.pdf}\label{fig:large_translation:small}}\hspace{1em} % or 2_gas_0.3.png
+    \subfloat[small translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.3.png}\label{fig:large_translation:small}}\hspace{1em}
+    % \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.pdf}\label{fig:large_translation:large}} % or 2_gas_0.9.png
+    \subfloat[large translation]{\includegraphics[width=0.47\linewidth]{TOG/figs/large_tran/2_gas_0.9.png}\label{fig:large_translation:large}}
     \Caption{Comparing foveal quality with regards to translation range. }
     {%
-    Using our method, \subref{fig:large_translation:small} shows a fovea image \dnc{from training dataset} with a small-scale translation box (0.3 meters); \subref{fig:large_translation:large} shows a fovea image with a large-scale translation box (0.9 meters).
-In the 2D illustrative inset: the blue circles indicates the distribution of the spheres; the orange arrow shows the eye's translation range.
+    Using our method, \subref{fig:large_translation:small} shows a fovea image \dnc{from training dataset} with a small-scale translation box (0.3m); \subref{fig:large_translation:large} shows a fovea image with a large-scale translation box (0.9m).
     }
@@
 Currently, we sample the scene fully based on eccentricity, considering acuity and stereopsis.
-However, we envision that fine grained visual sensitivity analysis, such as luminance \cite{Tursun:2019:LCA} or depth \cite{Sun:20:OE}, would provide more insights on achieving further higher quality and/or faster performance.
+However, we envision that fine-grained visual sensitivity analysis, such as luminance \cite{Tursun:2019:LCA} or depth \cite{Sun:20:OE}, would provide more insights on achieving even higher quality and/or faster performance.
 Also, our multi-spherical representation is specialized to the scene for optimal quality.
@@
 
-For simplicity, the spatial-temporal joint optimization connects the output precision and latency to the number of the spheres ($\sphereNum$) but not their radii $\mathbf{\sphereRadius}$.
+For simplicity, the spatial-temporal joint optimization in \Cref{sec:method:optimization} connects the output precision and latency to the number of the spheres ($\sphereNum$) but not their radii $\mathbf{\sphereRadius}$.
 This is due to the significant amount of training with parameter sampling.
-Incorporating the parameters into single training process may significantly reduce the time consumption for the optimization.
+Incorporating the parameters into a single training process may significantly reduce the time consumption for the optimization.
 With the adaptive training process, a content-aware distribution (i.e., $\mathbf{\sphereRadius}$) of the spheres would further improve the synthesis quality and performance.
 %\qisun{note to myself: Right now we use did not optimize for radii but only number of spheres}
+
+\zh{shall we end with non-limitation sentences?}\qisun{yep, future work is positive than limitations.;)}

### 2021-03-09 TOG/fig_blending.tex
submission

--- 2021-01-28 TOG/4-method.tex
@@
-\section{Method}
-\label{sec:method}
-In this section, we present our perception-guided neural scene synthesis for virtual reality.
-Based on a concentric spherical representation (\Cref{sec:method:representation}), our system predominantly comprises of two main steps: visual- and stereo- acuity adaptive view synthesis (\Cref{sec:synthesis}), and image-based real-time rendering to composite stereo display frames (\Cref{sec:method:blending}).
-We then craft an analytical spatial-temporal model to guide the optimization of our intra-system variables for desired precision-performance balance (\Cref{sec:method:optimization}).
-\input{TOG/fig_notations}
-\subsection{Egocentric Neural Representation}
-\label{sec:method:representation}
-The recent single-object-oriented ``outside-in'' view synthesis methods \cite{sitzmann2019deepvoxels,mildenhall2020nerf} typically represent the training targets using uniform voxelization.
-However, immersive VR environments introduce open challenges for such parameterization due to the rapid variation of the egocentric (first-person) viewing (e.g., \Cref{fig:teaser:scene}).
-As a consequence, the neural representation on large virtual environments typically suffer from ghosting artifacts, low resolution, or slow speed (e.g., the right triangle in \Cref{fig:teaser:quality}).
-%We tailor our method specifically for egocentric (first-person) views in immersive VR environments.
-%rendering of immersive viewer-centered ``inside-out'' scenes is challenging owing to the rapid variation of scene content. 
-\paragraph{Egocentric coordinates}
-To tackle this problem we are inspired by the recent panorama dis-occlusion methods \cite{Lin:DeepPanorama,Benjamin:2020:RTV,Broxton:immersiveLF} to depict the rapidly varying first-person views with concentric spheres.
-This representation enables robust rendering at real-time rates and allows for 6DoF interaction to navigate complex immersive environments.
-As visualized in \Cref{fig:notations}, our representation is parameterized with the number of concentric spheres ($\sphereNum$) and their respective radii ($\mathbf{\sphereRadius}=\{\sphereRadius_i\},i\in [1,\sphereNum]$).
-Our network aims at predicting the function $\mlpFunc(\hat{\SpatialPt})\triangleq(r,g,b,d)$ of any position $\mathbf{\hat{\SpatialPt}}({\sphereRadius_{i,i\in[1,\sphereNum]},\theta,\phi})$ on these $\sphereNum$ spherical surfaces.
-%\dnc{$\mathbf{q}$ may be more proper}\qisun{Just to avoid confusion with \Cref{fig:notations}.
-What do you think?}\dnc{Then perhaps we can use $\hat{q}$ as shown in \Cref{fig:notations} I think the alphabet 'q' means it's in spherical system, 'p' never exists anywhere else}
-Here, $(r,g,b)$ and $d$ are the predicted color and density, respectively. 
-We then employ a ray marching through this intermediate function to produce images of specified views.
-
-\paragraph{Training and synthesis}
-%\dnc{Perhaps we can merge this paragraph with above section? And move the last paragraph of above section to here as a intro. Then this section is more focusing on GAZE-AWARE}
-For each scene, we approximate the function $\mlpFunc$ by training a multilayer perceptron (MLP) network, similar to \cite{mildenhall2020nerf}.
-The network is composed of $\mlpLayerNum$ fully-connected layers of $\mlpChannelNum$ channels, each using ReLu as the activation function.
-We will further determine $\mlpLayerNum$, $\mlpChannelNum$ and $\sphereNum$ via a spatial-temporal optimization in \Cref{sec:method:optimization}.
-Once trained, the network achieves 4D prediction on the spherical voxels, on which a ray marching rendering is performed to synthesize the final frames for display.% based on users' frame-wise views. 
-Similar to \cite{mildenhall2020nerf}, we apply a trigonometric input encoding and compute the L2-distance between the volume-rendered and mesh-rendered images as the training loss function.
-We discuss the specifics of training data creation and implementations in \Cref{sec:impl}.
-
-\subsection{Gaze-Aware Acceleration}
-\label{sec:synthesis}
-\qisun{@nianchen: does this section title sounds reasonable to you?}
-%The scene representation above is created to parameterize a given 3D space for neural synthesis. 
-%the foveated rendering manner of VR display through training two orthogonal neural networks that are tailored for the perceptual acuity variances.
-Our concentric spherical coordinate system, as described in \Cref{sec:method:representation}, addresses the view variance problem in VR.
-However, it sill suffers from significant rendering latency (seconds per frame).
-This is one of the core challenges causing neural representation to be unsuitable yet for VR.
-To combat the common performance problem of view synthesis, we predict multiple elemental images based on the viewer's real-time head and gaze motions\nothing{ instead of synthesizing a globally uniform image}.
-The elemental images are tailored for the visual acuity variances.
-We detail our network design and elemental image synthesis in the following section.
-We enable real-time responsiveness via introducing gaze-adaptive synthesis mechanisms.
-\input{fig_system}
-\paragraph{Adaptive visual acuity} 
-%the user's current view $[\rayo,\camDir]$ and gaze position $\gazeDir$ that are tracked in real-time by a VR headset.
-Existing neural scene representation networks predict uniform resolution images for offline third-person viewing.
-Due to the trade-off between quality and latency during inference, they are typically executed with low quality and/or high latency when it comes large field-of-view and high resolution imagery.
-\praneeth{this line is confusing}
-Our method, without compromising performance, generates {\it perceptually} high quality frames matching the retinal acuity, i.e. how the retina perceives dynamic and stereo frames.
-% Note that foveal/peripheral retinal images requires high/low pixel-wise precision but low/high FoV (number of pixels).
-Thus, we train two orthogonal neural networks that are tailored for the perceptual acuity variances.
-Because of foveation of imagery by the human visual system \cite{Guenter:2012:F3G}, we train two orthogonal networks (i.e., $\mlpFunc$s) to synthesize foveal ($\imageFoveal(\rayo,\camDir,\gazeDir)$, $0-20$deg), and peripheral elemental images (i.e., two $\mlpFunc$s to deploy for ray marching) for each eye.
-The periphery includes mid- ($\imageMid(\rayo,\camDir,\gazeDir)$, $0-45$deg) and high- eccentricity visual fields ($\imageFar(\rayo,\camDir$), $0-110$deg), see \Cref{fig:system}.
-Note that $\imageFar$ is independent from the gaze direction $\gazeDir$. 
-%The images are defined as functions that maps individual camera rays ($\rayo,\rayd$) to an RGB color.
-\dnc{Could move to section 4?
-I think it's implementation related}\qisun{I think the PDD numbers give readers a rough impression of the varied quality across each of the image, right?} Specifically, to incorporate the display hardware properties and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $128^2$/$256^2$/$230\times256$ resolutions respectively.
-That is, the $\imageFoveal$ has the highest spatial resolution of $6.4$ pixels per degree (PDD), higher than those of $\imageMid$ ($5.7$ PDD) and $\imageFar$ ($2.33$ PDD).
-%In each frame, the two networks independently return three elemental images, as seen from the separation range in deg, gradually larger areas along eccentricity.
-An example is shown in \Cref{fig:system}.
-
-\paragraph{Adaptive stereo-acuity}
-{
-To be rendered on head-mounted displays, stereo elemental images $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$ ($l$ and $r$ represent left and right eyes respectively) are required.
-This doubles the computational cost that is critical for latency- and frame-rate- sensitive VR experience
-\nothing{The separation of the three elemental images considers the varied visual acuity.
-However, in a head-mounted stereo VR displays, the rendered image are for two eyes, resulting in double computation for $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$.
-Here $l$ and $r$ represent the projection images for left ($\rayo^l$) and right ($\rayo^r$) eyes respectively.
-Whereas, because $\imageMid^{\{l,r\}}$ and $\imageFar^{\{l,r\}}$ demand high spatial resolution due to their large eccentricity coverage, }
-}
-(please refer to \Cref{sec:study:intra} for detailed comparisons).
-
-To tackle this problem towards real-time experience, we accommodate the inference process with the adaptive stereo-acuity in the periphery.
-Specifically, besides the visual acuity, psychophysical studies have revealed human's significantly declined stereopsis while receding from the gaze point \cite{mochizuki2012magnitude}.
-Inspired this characteristic, instead of inferring $6$ elemental images, we perform the computation with $\imageFoveal^{\{l,r\}}$, $\imageMid^m$, and $\imageFar^m$, where $m$ indicates the view at the midpoint of the left and right eyes.
-The individual images are further blended for display, as detailed in \Cref{sec:method:blending}.
-\Cref{fig:mono} visualizes the stereo changes from the adaptation.
-\input{TOG/fig_mono_comparison}
-
-\subsection{Real-Time Frame Composition}\dnc{May be moved to section 4 (merge with the paragraph 'Integration').
-I think it's not closely related to our methodology and we have no key contributions about that.
-It's more like an engineering issue}\qisun{This whole section?
-I was concerned that would break the method completeness.
-Intentionally kept it short according to contribution level here.}
-\label{sec:method:blending}
-In real-time, we perform a image-based rendering that blends the elemental images. The output frames are rendered on the stereo VR displays.
-To accommodate the uniformized $\imageMid^m$, we shift it with $-/+ \frac{\rayo^r-\rayo^l}{2}$ for left/right eye respectively. 
-\dnc{Not accurate. shifting is related to mean of fovea depthmap inferred in actual implementation to minimize the misalignment between fovea and mid}
-{
-With the obtained elemental images as input, a fragment shader is executed to generate final images for each eye.
-Two adjunct layers are blended using a smoothstep function across $40\%$ of inner layer.
-\nothing{After obtaining the individual images, our fragment shader retrieves the three frames and applied a pixel wise bilinear interpolation at each $10\%$ connecting areas between two adjunct layers.}
-}
-This enhances visual consistency on the edges between layers \cite{Guenter:2012:F3G}.
-Moreover, to further preserve peripheral elemental images' visual fidelity due to its low PDD, we enhance the contrast following the mechanism of \cite{Patney:2016:TFR}, the details are visualized in \Cref{fig:method:blending}.
-%\dnc{Note: the contrast enhancement is performed on elemental images, before blending.
-Different enhancement parameters are applied for fovea, mid and periph layers}
-
-\begin{figure*}[htb]
+\begin{figure*}[ht]
     \centering
@@
     \subfloat[blending]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended2.pdf}\label{fig:method:blending:blending}}\hspace{1em}
-    \subfloat[contrast enhance]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
+    \subfloat[contrast enhancement]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
     \Caption{Our multi-layer real-time rendering system.}
@@
 \end{figure*}
-
-\subsection{Latency-Quality Joint Optimization}
-\label{sec:method:optimization}
-As a view synthesis system based on sparse egocentric representation (the density of $\sphereNum$ spheres) and neural networks (the $\mlpLayerNum\times\mlpChannelNum$), the method inevitably introduces approximation errors.
-On the other hand, these variables also determine the online computational time that involves inferring function $\mlpFunc$ and ray marching the $\sphereNum$ spheres.
-However, VR strictly demands both quality and performance. We present a spatial-temporal model that analytically depicts the correlations and optimizes the variables for ideal user experience.
-
-\paragraph{Precision loss of a 3D scene}
-As shown in \Cref{fig:notations}, under the egocentric representation, a 3D point $\pt$ is re-projected as the nearest point on a sphere that connects it to the origin point:
-\begin{equation}\label{eq:closestPoint3D}
-{\pt^\prime}(\sphereNum,\mathbf{\sphereRadius},\SpatialPt) \triangleq \sphereRadius_k \frac{\pt}{\norm{\pt}}, \ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
-\end{equation}
-
-Similar to volume-based representation, the multi-spherical system is also defined in discrete domain.
-The sparsity thus naturally introduces approximation error that compromises the synthesis quality.
-To analytically model the precision loss, we investigate the geometric relationship among the camera, the scene, and the representation.
-As illustrated in \Cref{fig:teaser:scene,fig:notations},  for a sphere (located at origin point) with radius $\sphereRadius$, its intersection (if exists) with a directional ray $\{\rayo,\rayd\}$ is
-\begin{equation}
-\intersectionFunc(\sphereRadius,\rayo,\rayd) = \rayo + \left({\left((\rayo\cdot\rayd)^2-\norm{\rayo}^2+\sphereRadius^2\right)}^{\frac{1}{2}}-\rayo\cdot\rayd\right)\rayd,
-\label{eq:raySphereIntersection}
-\end{equation}
-where $\rayo$ and $\rayd$ are the ray's origin point and normalized direction respectively.
-% \begin{equation}
-% \left\{\intersectionFunc(\sphereRadius_i,\rayo,\rayd) \right\}\ |\ i\in[1,2,\dots,\sphereNum].
-% \end{equation}
-Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt$ represented in spherical coordinates, the ray connect them is with the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$.
-This ray may intersect with more than one spheres.
-Among them, the closest one to $\SpatialPt$ is:
-\begin{equation}\label{eq:closestPointImg}
-\begin{aligned}
-    \hat{\pt}&(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt) \triangleq \\
-    &\intersectionFunc(\sphereRadius_k,\rayo,\rayd(\SpatialPt,\rayo))\ |\ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
-\end{aligned}
-\end{equation}
-In the 3D space, the offset distance $\norm{\pt^\prime-\hat{\pt}}$ indicates the precision loss at $\pt$ from the representation. By integrating over all view and scene points, we obtain:
-\begin{equation}
-\begin{aligned}
-\sparseError(\sphereNum, \mathbf{\sphereRadius})  &= \iint \norm{\pt^\prime(\sphereNum,\mathbf{\sphereRadius},\SpatialPt)-\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)} \mathbf{d}\SpatialPt \mathbf{d}\rayo,\\
-&\forall \{\rayo, \SpatialPt\} \text{ pair without occlusion in between}.
-\label{eq:sparseError}
-\end{aligned}
-\end{equation}
-$\sparseError$ depicts how the generic representation precision of a scene, given a coordinate system defined by $\sphereNum$ and $\mathbf{\sphereRadius}$.
-
-In comparison, a neural representation aims at predicting projected image given a camera position $\rayo$ and orientation $\camDir$.
-Thus, we further extend \Cref{eq:sparseError} to image space to analyze the error given a set of camera's projection matrix  $\projectionMatrix(\rayo,\camDir)$ as
-%However, typical 3D representations are not uniformly distributed points.
-Moreover, the network only predicts pixel colors/intensities but not depth.
-By extending \Cref{eq:sparseError} to the captured image space data, we obtain
-\begin{align}
-\imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo, \camDir)  = \int \norm{\projectionMatrix(\rayo,\camDir)\cdot\left(\SpatialPt-
-\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)\right)}\mathbf{d}\SpatialPt.
-\label{eq:imageError}
-\end{align}
-From \Cref{fig:notations,eq:imageError}, we observe:
-given a fixed min/max range of $\mathbf{\sphereRadius}$, $\sphereNum$ is negatively correlated to $\imgSpaceError$;
-with a fixed $\sphereNum$, the correlation between distribution of $\sphereRadius_j$ and scene content (i.e., distribution of $\SpatialPt$s) also determines $\imgSpaceError$.
-
-However, for neural scene representation, infinitely enlarging $\sphereNum$ would significantly raise the challenges in training precision and inference performance.
-Likewise, increasing prediction and rendering densities (i.e., higher $\sphereNum$ or $\mlpLayerNum$/$\mlpChannelNum$) naturally improves the image output quality for lower \Cref{eq:imageError} while significantly increases the computation during ray marching, causing quality drop stretched along time.
-In the performance-sensitive VR scenario, the latency breaks the continuous viewing experiment and may cause simulator sickness.
-Thus, we determine the coordinate variables with a content-aware optimization towards optimal quality-speed performance.
-
-\paragraph{Spatial-temporal modeling}
-Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform a spatial-temporal joint modeling to determine the optimal coordinate system ($\sphereNum$ and $\mathbf{\sphereRadius}$) for positional precision and MLP settings ($\mlpLayerNum$/$\mlpChannelNum$) for color precision that adapt to individual computational resources and scene content.
-This is achieved via a latency-precision modeling in the spatial-temporal domain.
-%Specifically, we model the prediction image $P$ with gaze position ($G$) at time $t$ as $P_G(t)$ and the ground truth retinal image (from a simulated local rendering) as $\hat{P}_G(t)$.
-%In practice, increasing the number of spheres $\sphereNum$ or number of MLP layers may affect the computational and transmissional latency but will improve the quality of individual frames (as in \Cref{eq:imageError}).
-Thus, we define the perceptual error as:
-\begin{equation}
-\begin{aligned}
-\finalError(\sphereNum, &\mlpLayerNum, \mlpChannelNum) = 
-\sum_t\int \mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)\times\\ &\norm{\projectionMatrix(\rayo_t,\camDir_t)\cdot\SpatialPt-\projectionMatrix(\rayo_{t-\latency},\camDir_{t-\latency})\cdot\pt(\sphereRadius_k,\rayo_{t-\latency},\rayd_{t-\latency})}\mathbf{d}\SpatialPt,
-%&\int  \norm{ \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_t, \rayd_t) - \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_{t-l(\sphereNum, \mathbf{\sphereRadius})}, \rayd_{t-l(\sphereNum, \mathbf{\sphereRadius})})) } \mathbf{d} t
-\label{eq:error:image}
-\end{aligned}
-\end{equation}
-where $\latency\triangleq\latency(\sphereNum, \mlpLayerNum, \mlpChannelNum)$ is the system latency with a given coordinate and network setting.
-$\mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)$ is the L1-distance between the four channels ($r$,$g$,$b$,$a$) and0 the highest values of network layer settings.
-For simplicity, we assumed uniformly distributed $\mathbf{\sphereRadius}$ with a fixed range of the spherical coverage.
-
-Suggested by Albert et al.
-\shortcite{albert2017latency}, the latency for a foveated system shall reach below \textasciitilde$50$ms for undetectable artifacts.
-Given our test device's eye-tracking latency \textasciitilde$12ms$ and the photon submission latency \textasciitilde$14ms$ (\cite{albert2017latency}), the synthesis and rendering latency shall be less than $L_0 = 24$ms.
-Thus, we determine the optimal $\{\sphereNum, \mathbf{\sphereRadius}\}$ to balance latency and precision as
-\begin{equation}
-\begin{aligned}
-    &\argmin_{\sphereNum, \mlpLayerNum, \mlpChannelNum} \finalError(\sphereNum, \mlpLayerNum, \mlpChannelNum), \\
-    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L_0
-\end{aligned}
-\end{equation}
-\Cref{fig:optimization} visualizes an example of the optimization mechanism for in fovea.
-The optimized results ($\sphereNum, \mlpLayerNum, \mlpChannelNum$) for individual networks are detailed in the next section.
-The visual quality is validated by psychophysical study (\Cref{sec:study:user}) and objective analysis (\Cref{sec:study:quality}).
-The latency breakdown of our system is reported in \Cref{sec:study:intra}.
-\input{TOG/fig_optimization}
-\input{TOG/4.5-implementation}

### 2021-03-09 TOG/fig_lpips.tex
submission

@@
     \subfloat[bedroom]{\label{fig:lpips:bedroom}\includegraphics[width=0.48\linewidth]{TOG/figs/lpips/lpips-bedroom-84.pdf}}
-    \Caption{LPIPS visualization of all scenes as well as comparison of {\bf OURS}, {\bf NeRF}, and {\bf F-GT} over the visual field.}
+    \Caption{LPIPS visualization of all scenes and comparison of {\bf OURS}, {\bf NeRF}, and {\bf F-GT} over the visual field.}
     {%
-    X-axis indicate eccentricity range ($0$ to 110).
+    X-axis indicates the eccentricity range ($0$ to 110).
 Y-axis shows the LPIPS loss \cite{zhang2018unreasonable} with a full resolution rendering as reference.
 Lower values mean more similar perceptual quality to the reference.
-The semi-transparent layer indicate first and third quarterlies.
+The semi-transparent layers indicate first and third quarterlies.
 The inset figures' corresponding eccentricity ranges are indicated by the dashed lines.

### 2021-03-09 TOG/fig_mono_comparison.tex
submission

@@
-\begin{figure}[hb]
+\begin{figure}[tbh]
     \centering
@@
     \subref{fig:mono:wo} shows the rendered image with 6 retinal sub-images ($\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$).
-    \subref{fig:mono:w} shows the rendered image with our adaptive and accelerated inference considering foveated stereoacuity ($\imageFoveal^{\{l,r\}}$, $\imageMid^{\{m\}}$, and $\imageFar^{\{m\}}$).
+    \subref{fig:mono:w} shows the rendered image with our adaptive and accelerated inference considering foveated stereoacuity ($\imageFoveal^{\{l,r\}}$, $\imageMid^{\{c\}}$, and $\imageFar^{\{c\}}$).
 Our method preserves full stereopsis in the fovea while reducing the angular resolution in the periphery for accelerated inference.

### 2021-03-09 TOG/fig_optimization.tex
submission

@@
-\begin{figure}[htb]
+\begin{figure}[t]
     \centering
@@
     {%
-    \protect\subref{fig:optimization:heat:quality}/\protect\subref{fig:optimization:heat:time} shows \protect\Cref{eq:imageError}/latency(in millisecond under pytorch implementation) of an example foveal network.
-    The values are computed with various settings of $\mlpLayerNum\times\mlpChannelNum$ (x-axis) and $\sphereNum$ (y-axis).
+    \protect\subref{fig:optimization:heat:quality}/\protect\subref{fig:optimization:heat:time} plots $\imgSpaceError$ in \protect\Cref{eq:imageError}/latency of an example foveal network during the optimization process.
+(in milliseconds\nothing{ under pytorch implementation}) .
+    The values are computed with various settings of $\mlpChannelNum\times\mlpLayerNum$ (X-axis) and $\sphereNum$ (Y-axis).
 The second row indicates a corresponding foveal image under different settings.
-Our method balances both perceptual quality and latency.\nothing{\qisun{(Jan 24) What is the unit of \subref{fig:optimization:heat:time}?
-It shouldn't be second, right?}\dnc{It's millisecond.}}%nothing
+Our method balances both perceptual quality and latency.\qisun{note to myself: add deeper discussion and insights here.}
     }

### 2021-03-09 TOG/fig_ray_bounding.tex
submission

@@
-\begin{figure}
+\begin{figure}[hb]
     \centering
-    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
-    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
-    \Caption{Comparison of ray marching bounding.}
+    \subfloat[infinite upper bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}\label{fig:impl:bounding:off}}\hspace{1em}
+    \subfloat[limited bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}\label{fig:impl:bounding:on}}
+    \Caption{The effect of ray marching bounding.}
     {%
+    Compared to infinite ray marching \subref{fig:impl:bounding:off}, bounding the ray marching further improves visual quality in \subref{fig:impl:bounding:on}.
     }
-    \label{fig:method:bounding}
+    \label{fig:impl:bounding}
 \end{figure}

### 2021-03-09 TOG/fig_system.tex
submission

@@
-\begin{figure}[bht]
+\begin{figure}[b]
     \centering
-    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_fovea.png}\label{fig:system:fovea}}\vspace{1em}
-    \subfloat[mid-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_mid.png}\label{fig:system:mid}}\vspace{1em}
+    \subfloat[fovea]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_fovea.png}\label{fig:system:fovea}}\hspace{1em}
+    \subfloat[mid-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_mid.png}\label{fig:system:mid}}\hspace{1em}
     \subfloat[far-periphery]{\includegraphics[width=0.3\linewidth]{TOG/figs/layer_blend/lobby_view0000_periph.png}\label{fig:system:far}}
 
-    \subfloat[rendering]{\includegraphics[width=0.9\linewidth]{TOG/figs/layer_blend/lobby_view0000_blendmask.png}\label{fig:system:blending}}
-    \Caption{Synthesis output and rendering.}
+    \subfloat[display image]{\includegraphics[width=0.9\linewidth]{TOG/figs/layer_blend/lobby_view0000_blendmask.png}\label{fig:system:blending}}
+    \Caption{Visual acuity adaptive synthesis and rendering mechanism.}
     {%
-    \subref{fig:system:fovea}/\subref{fig:system:mid}/\subref{fig:system:far} shows our individual gaze-contingent synthesis for fovea (within $20$deg)/mid-periphery (within $45$deg)/far-periphery(within $110$deg, the capability of our VR display), respectively.
+    \subref{fig:system:fovea}/\subref{fig:system:mid}/\subref{fig:system:far} shows our individual gaze-contingent synthesis for fovea (within $20$ deg)/mid-periphery (within $45$ deg)/far-periphery (within $110$ deg, the capability of our VR display), respectively.
     \subref{fig:system:blending} illustrates our real-time rendering system by dynamically blending the individual images to the final displayed frame.

### 2021-03-09 TOG/teaser_fig.tex
submission

@@
 \subref{fig:teaser:scene} Visualization of our active-viewing-tailored egocentric neural scene representation.
-\subref{fig:teaser:latency} Our method allows for extremely low data storage of high quality 3D scene assets and achieves high perceptual image quality and low latency for interactive virtual reality.
-Here, we compare our method with the state-of-the-art alternative neural synthesis \protect\cite{mildenhall2020nerf} and foveated rendering solutions \cite{perry2002gaze,Patney:2016:TFR} and demonstrate that our method significantly reduces time consumption and data storage (more than $99\%$ reduction for both) for first-person immersive viewing.
-The orange circle indicates viewer's gaze position.
-We also show zoomed-in views of the foveal images generated by various methods in \subref{fig:teaser:quality} and show that our method is superior compared to existing methods in achieving high perceptual image fidelity.
+\subref{fig:teaser:latency} Our method allows for extremely low data storage of high-quality 3D scene assets and achieves high perceptual image quality and low latency for interactive virtual reality.
+Here, we compare our method with the state-of-the-art alternative neural synthesis \protect\cite{mildenhall2020nerf} and foveated rendering solutions \cite{perry2002gaze,Patney:2016:TFR} and demonstrate that our method significantly reduces more than $99\%$ time consumption (from 9s to 20ms) and data storage (from 100MB mesh to 0.5MB neural model) for first-person immersive viewing.
+The orange circle indicates the viewer's gaze position.
+We also show zoomed-in views of the foveal images generated by various methods in \subref{fig:teaser:quality} and show that our method is superior compared to existing methods in achieving high perceptual image fidelity.%\qisun{The quality of Nerf is super invisible in this design to me.
+@Zhenyi, let's flip Nerf and F-GT.}
 %\qisun{(Jan 24, 2021) Let's add some latency numbers to

### 2021-03-11 TOG/rebuttal-new.md
add rebuttal-new

We thank the reviewers' valuable comments. Allow us to address the main points, followed by detailed questions.

We would like to emphasize that the main contribution of this work is devising a human-vision-matched (resolution and stereo acuity) 6DoF neural representation that can be inferred in real-time for VR near-eye displays. The method is applicable for not only virtual environments but also physical captures/spaces, compatible with view-dependent effects for future low-storage, wearable and cloud-based VR/AR where only the pre-trained scene models will need to be dynamically transmitted. The research addresses the high-latency<!--Qi: any nerf number for its resolution vs time consumption per frame?--> problem in current neural representation, and the limited expressiveness (such as correct lighting/illumination effects) in the recent image (panorama or multi-layer image)-based representation/re-projection methods.

# Necessity of neural representation method 
Q: Comparison with traditional mesh representation (R2/R4):

First, There is a misunderstanding about the applicable case of the work. Our (and NeRF's) goal is not about encoding mesh-based virtual scenes. The neural network is a representation of radiance field. That is, besides virtual scenes, physical scenes just follow exactly same computation procedure. The only data we need to train such a model are several RGB images with corresponding intrinsic and extrinsic parameters.
We apologize for not including such examples and will include new examples with both physical scenes and dynamic lighting effects to visually demonstrate the necessity.

Further, the network is not related to the complexity of scenes. The size of a mesh scene is related to its geometric and textural fineness, which may be extremely large (e.g. hundreds or even thousands MB) with full visibility (e.g., terrain), but the size of the network is a constant (0.5MB in our experiments). In the cloud-based application scenario that all data are acquired from cloud services at run-time, the data size is an important factor that affects the user experience.

Q: Comparison with MSI methods (R2):

First, our neural synthesis method naturally support correct dis-occlusions and motion parallax as the representation treats each view uniformly (inside the first sphere as in Fig. 11).
Meanwhile, for physical scenes, MSI methods (e.g. panorama-based pre-baking) require additional image-based inference of occlusions and depths. Ln267 mentions that these methods "typically suffers from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space". Directly recording a precise multi-layer panorama requires sophisticated devices.

Second, similar to NeRF, the method does support view-dependent effect that is lacking with image-based panorama pre-baking solutions. In the current demo, we didn't include global illumination for faster data preparation. We will include visual examples demonstrating it.

To clearly demonstrate the difference, we will include a visual comparison with panorama+disocclusion method.

<!--Qi: OURS with a nerf style rendering supports correct illumination with different camera views, and we will promise to show photorealistic results during minor revision cycle. We need more points beyond this though.-->
<!--ZH: Need to explain NeRF for R2.6
   (R2: why do you need to run a ray-caster with a neural network at each sample location to achieve this style of foveated rendering? The method does not support view-dependent effects, the range of motion seems very limited, and the output resolution is quite low.)-->

# Limitations

<!-- (r1) restricted position and direction, and gaze motions?
     (r2) low (angular) resolution // head volume? frequency of transmiting network?
     (r4) occlusion (motion parallel): not clearly demonstrated-->

Q: The range of motion and parallax support (R1/R2/R4)

Our gaze range covers the entire display-supported FoV ($100^\circ$). The rotational range is only determined by the computing resources and duration for training. A larger rotation means a larger training set (longer training time).  In the "gallery" scene, we trained for up to $240^\circ$ rotational range. We are confident to include a $360^\circ$ rotational result within a minor revision cycle.
For head volume, in our experiments, we tested $30cm^3$. Ln1068-1070 and Fig.11 depict the balance: although computationally the innermost sphere (that defines the supported translation range) could any large, significantly enlarge it would cause visual quality decreasing.
Developing a multi scale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, extending the applicability of our method to large-scale scenes with strong occlusions. We support motion parallax and occlusion within the range determined by the innermost sphere. Unlike the pure rotation method, we do not need to train a network for every viewpoint.

<!--Q: Occlusion and Motion parallax(R2/R4)-->

We support correct occlusion and motion parallax at any view position in the range of motion. These visual effects are demonstrated, although not significantly visible, in the video (1m20s ~ 1m25s). We will include larger translation to demonstrated more significant occlusion and motion parallax within a minor revision cycle.

<!--Qi: (3/10) I think we can potentially drop this one since R2's point isn't really about the resolution but alternative solutions, right?-->

Q: The angular resolution (R2)

Our angular resolution (6.4 PPD) is a trade-off between performance (determined by computational resource) and quality (determined by display capability).

<!--DNC: How to measure the actual PPD of VR? I've searched a value of 5.1'/pixel (about 11.8 PPD)-->

We do use full resolution ground truth images F-GT in the user study.

<!--Qi: We are asked about "WHY" the 6PPD appears the same as 16 PPD even in fovea, not repeating the content already-->
<!--DNC: "Why" is hard to talk about because to some extent, this shows that the method we regard as ground truth is not the best foveation rendering method.-->

# Other

Q: Mobile devices (R1/R2):
<!-- (r1) experiment is done on high-end machine but paper is targeting for mobile device?
     ZH: We mentioned 'mobile' only once, in the introduction L160. Shall we try and report a number on mobile VR device?
     (r2) On what hardware is this method running in real-time
     ZH: CPU is Intel(R) Xeon(R) Gold 6230R CPU@2.10GHz (256GB RAM) and one NVIDIA GTX 3090 graphics card.
     (r2) distance towards low latency, and low-storage VR/AR platforms? -->

We target at future wearable VR/AR devices that are "akin to" (Ln160) mobile devices whose low storage and computational power demands dynamically fetching content from cloud. Our method enables fast fetching with only 0.5mb for a range of local camera 6DoF motions so that the system can keep running without any network-induced latency. Since it allows local translation, the method can fetch (for future positions) while rendering in parallel. 
That is, the research targets at the "constant" viewing experience without long pre-installation. It compares to the browser games vs. traditionally locally installed games.

<!--Further, on mobile devices, the supported resolution (PPD) is lower, demanding low inference computation.-->

<!--Qi: @Nianchen, the best if you could quickly try to run that on a mobile phone or having collaborators do that asap
We care about the experience without long installation, compariable to web-based gaming vs traditional gaming.
-->

Q: Static scene encoding (R1/R3/R4):
<!-- (r1) no user interaction, mobile devices can store
     (r3) scope should be clarified-->
We currently focus on encoding/decoding static scene in this project. Encoding dynamic and interactive scenes in neural scene representation is an orthogonal challenges that are getting attention to the neural rendering community: "D-NeRF: Neural Radiance Fields for Dynamic Scenes", arxiv 2020.
We thank R3 for the suggestion of clarifying the scope. We will tone down and state this limitation with static scenes and future attempts to extend the model with dynamic scenes while also incorporating temporal visual perception such as saccades (R4).

<!--## About encoding performance (r3.5, r4.3):
(r3) scope should be clarified
(r4) viewing is instant, but creation is not (encoding/training is slow)-->

This research assumes that the encoding of the scene (network training) is executed in a large-scale cloud service only once (which is practically available), so the computing overhead is ignored.

Q: User study (R1/R3/R4/R5):

<!-- (r1, r4) short, gaze fixed / static
     (r1) leave no time to experience ego-motion parallax effects
     (r3) narrow foveation methods selection-->

The stimuli's duration was designed according to the requirement (0.3-0.4s) of preventing refocusing and shifting (Ln811) as studied in:
"Dynamics of accommodation responses of the human eye" Campbell & Westheimer,  The J. physiology, 1960

Otherwise, the gaze changes will cause unfair comparisons, resulting in bias.

<!--We asked users to fixate at certain gaze positions so everyone's views were consistent across conditions and participants. (TBD) We set duration to 300ms so it is close to saccade experience(?)-->

<!-- 3/9
1) 300ms > per frame
2) even not enough, already proved ours > nerf
motion parallax is exactly our beneifts than panorama synthesis we will compare with panorama
it is fair for each view
-->

With eye-tracking in VR, foveated rendering is chose as GT because that has been shown to be perceptually identical to a full resolution rendering. Higher resolution on the peripheral range has minimal effects on perception. We have done an orthogonal pilot study verifying the identicality between F-GT and GT. We will include the details. 
<!-- we are compatible to any image-space foveation methods-->

Q: Source code (R3/R5):

Yes, we will make the entire source code publicly open, including models and data.

Q: Storage comparison with [Broxton2020] (R2)

First, we would like to point out a calculation mistake. The result of $(6/10)^3*10$ is not $0.06$ but $2.16$. Second, [Broxton2020] involves video compression while our is the raw size. Some practical compression methods can be applied to furthur reduce the size (to 0.3MB tested by Zip).

<!-- Q: Experiment about storage? (r2.3)
What baseline requires 100MB of storage claimed in the introduction? The original 3D asset? There must be a middle ground which improves storage somewhat, without having to evaluate a neural network for every output pixel. -->

Q: The comparison should allow NeRF an additional 96 channels (R4)

We are using the same parameters that were used in our setup. Nevertheless, the performance of such NeRF configuration is still very slow.

<!--DNC: Same network parameters means same network capacity. Use large network capacity to compare our small network capacity is unfair.-->

Q: Why our method is better than baseline (i. e. NeRF)? (R4)

We use different coordinates (egocentric coordinates) while baseline uses 3D volume. We are for different tasks. 3D volume is not suitable for our inside-out view mode.

Q: Artifacts in video (R4)

We've checked that the double image artifacts (1m30) are artifacts caused by video codec. We will choose a proper codec for our demo video within minor revision cycles. <!-- black splotches, where they come from? (1m03) TO BE CHECKD-->

Q: Isn't the representation just alpha composited front to back? (R4)
Ray cast is different from alpha blending. it is an integration of density. We refer the reviewer to ["NeRF":Milden-hall et al.2020] for the detail.

## Should we reply to questions below?

Q: (Misunderstanding) need depth reconstruction for 360 photos application. (R4)

The neural encoding the scene information, including the depth. So we do not need extra depth information in run-time. Depth information can be inferred by ray casting using output density.

Q: Explanation to Fig. 7. (R4)

The bounding refers to the range of sampling in ray casting (the near and far plane), which affects the distribution of sampling distances. <!--we set upper bound to 10000 in (a) to approximate the infinite because sampling are done in disparity space ($1/distance$)-->

Q: Is NeRF here also running on a TensorRT implementation? (R4) 
No.
<!--ZH: explain the potential effects here.-->

Q: Composition of the user study (R5)
<!--"The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions." how did this come about? 3 pairs of conditions x 2 scenes? (r5.2)-->
<!--"Each condition from an individual scene consists of 3 different gaze positions." I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped? (r5.3)-->

6 trials (2 scenes x 3 gaze positions) x 6 pairs (permutation of 2 conditions). We predefined 3 gaze positions for each scene. Participants were asked to fixate each predefined gaze position for each trial and made the choice.

Q:  What is meant by perceptual similarity? color similarity? (R5)
<!--"LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image."-->

Perceptual similarity is not only color similarity. It indicates the similarity of human visual perception.
<!--ZH(TBD): IPIPS-->

Q: What is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to? (R5)
<!--"We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together."-->

IV: three conditions (F-GT, OURS, NeRF). DV: LPIPS value at eccentricity=5, 10, 15, ... ,105.

<!--DNC Remark:
1. 把view-dependent作为我们的一个重要研究目标是否合适？因为我们在之前的submission中没有相关表述和评估，现在突然变成了一个重点在开篇被提及，会不会让人感觉很奇怪？
2. PDD是一个质量相关的问题，虽然R2主要针对的不是我们的质量问题，但如果这个“WHY”我们不予回答，会不会让R2更加认为我们的方法不够好？
3. About comparison with foveation rendering method 看起来也是一个比较共性的问题（r1.5, r2.1, r3.4）-->

<!--
1. fixed
2. Why means since your res is already that low, WHY do you bother doing a nerf than just prebake. I think we answered the nessisity of using nerf than panorama/MPI
Re(DNC): I mean this: >>>
   In Technical Correctness and Reproducibility of R2:
   However, I found the statements on resolution to be inconsistent throughout the paper. In Section 3.2.1 the maximum angular resolution of the system is stated to be 6.4 PPD, however, the experiments compare with high resolution 13 PPD images (1440x1600 with a 110 FOV). How come no one was able to pick up on this drastic change in resolution during the user study. Were the ground truth images blurred to a maximum of 6.4PPD?
   <<< Anyway, it's a problem related to 3 so we can discuss them together later.
3. Yeah, I dont have a good clue of that yet. Do you have some good answers/excuses? Maybe ask Zhenyi too.
  Re(DNC): May be we can discuss later
-->

### 2021-03-11 TOG/rebuttal-old.md
add rebuttal-new

We thank reviewers for your valuable comments. Allow us to address the main points, followed by detailed questions.

We would like to emphasize that the main contribution of this work is devising a human-visual-system-matched real-time 6DoF neural synthesis method (***suitable for not only virtual environments but also physical captures/spaces, compatible with view-dependent effects***) for future low-storage and wearable VR/AR where only the pre-trained scene models need to be stored. 
The research addresses the high-latency(~xxx ms)/low-resolution(~xxx) problem in current neural synthesis, and the limited expressiveness (such as view-dependent lighting/illumination effects) in recent image (panorama or multi-layer image)-based representation/reprojection methods .

# Shared
## OURS vs. alternatives (why not panorama synthesis or multi-layer image, multi-sphere image representation)  (R2/R4/R2.4)

We will include a visual comparison with panorama+disocclusion method.
The method does support view-dependent effect, similar to NeRF. In the current demo, we didn't include illumination for faster data preparation. We will include visual examples demonstrating it.
Comparing with existing neural synthesis, it already reaches significantly higher resolution.

Add: Ln267
<!--Qi: OURS with a nerf style rendering supports correct illumination with different camera views, and we will promise to show photorealistic results during minor revision cycle. We need more points beyond this though.-->
<!--ZH: Need to explain NeRF for R2.6
   (R2: why do you need to run a ray-caster with a neural network at each sample location to achieve this style of foveated rendering? The method does not support view-dependent effects, the range of motion seems very limited, and the output resolution is quite low.)-->

## Constraints (r1.6, r2.5, r2.6, r4.4, r4.7):
<!--   (r1) restricted position and direction, and gaze motions?
   (r2) low (angular) resolution // head volume? frequency of transmiting network?
   (r4) occlusion (motion parallel): not clearly demonstrated-->

* Gaze and head rotation

Our gaze range covers the entire display-supported FoV (100deg). The rotational range is only determined by the computing resources and duration for training. A larger rotation means a larger training set (longer training time).  In the "gallery" scene, we trained for up to 240deg rotational range. We are confident to include a 360 rotational result within a minor revision cycle.

* Translation (head volume)

In our experiments, we tested $30cm^3$. Ln1068-1070/Fig.11 depict a limitation: developing a multi scale network that synthesizes the image from global to local level-of-details would be an interesting direction for future work, extending the applicability of our method to large-scale scenes with strong occlusions. We support motion parallax and occlusion within a certain range. Unlike the pure rotation method, we do not need to train a network for every viewpoint.
In the demo scenes<!--Qi: point to the exact time range-->, there are few close objects, so the motion parallax is not significantly demonstrated. We will include larger scale translation and rotation range within a minor revision cycle. 0.5MB is not for "one view", but a local "range" as visualized in Fig.11.

* Angular resolution

Our angular resolution (6 PPD) is a trade-off between performance (determined by computational resource) and quality (determined by display capability). 
and yet as verified by user-study, users cannot distinguish us from the ground truth (10+ PDD). However, this already significantly outperforms existing neural synthesis solutions. 
<!--Qi: We are asked about "WHY" the 6PPD appears the same as 16 PPD even in fovea, not repeating the content already-->

## Mobile devices (r1.1, r2.2):
<!--   
(r1) experiment is done on high-end machine but paper is targeting for mobile device?

ZH: We mentioned 'mobile' only once, in the introduction L160. Shall we try and report a number on mobile VR device?

(r2) On what hardware is this method running in real-time
ZH: CPU is Intel(R) Xeon(R) Gold 6230R CPU@2.10GHz (256GB RAM) and one NVIDIA GTX 3090 graphics card.

(r2) distance towards low latency, and low-storage VR/AR platforms?-->

We target at future wearable VR/AR devices that are "akin to" (Ln160) mobile devices whose low storage and computational power demands dynamically fetching content from cloud. Our method enables fast fetching with only 0.5mb so that the system can keep running without any network-induced latency. Since it allows local translation, the method can fetch (for future positions) while rendering in parallel. Further, on mobile devices, the supported resolution (PPD) is lower, demanding low inference computation.

<!--Qi: @Nianchen, the best if you could quickly try to run that on a mobile phone or having collaborators do that asap-->

## Static scene encoding (r1.2, r1.3, r3.5, R4):
<!--   (r1) no user interaction, mobile devices can store
   (r3) scope should be clarified-->
We focus on encoding/decoding static scene in this project. 
Encoding dynamic scenes and interaction support are orthogonal challenges that are getting attention to the neural rendering community:
"D-NeRF: Neural Radiance Fields for Dynamic Scenes", arxiv 2020.
We thank R3 for the suggestion of clarifying the scope. We will tone down and clarify this limitation for static scene and future intention to match the model with dynamic scenes and temporal visual perception such as saccades (R4).

   <!--Accept r3’s suggestion and express scope more clearly in later revisions-->
   <!--ZH: we should probably write down "a clearer scope" directly instead of claim we 'will'. (R3: the main goal of the present work is to focus on bringing the rich static environments in VR with efficient rendering 
   or this work considers only optimizing the decoding/rendering part of the problem, while the encoding is still an open research problem and the submission explicitly leaves it out of the scope.)-->

<!--## About encoding performance (r3.5, r4.3):
   (r3) scope should be clarified
   (r4) viewing is instant, but creation is not (encoding/training is slow)-->
   This research assumes that the encoding of the scene (network training) is executed in a large-scale cloud service only once (which is practically available), so the computing overhead is ignored. <!--This is not the scope of this article. 
   Accept r3’s suggestion and express scope more clearly in later revisions (ZH: better pick one or write a version of scope)-->

## User study (r1.4, r3.4, r4.6): (R1/R3/R4)
   <!--(r1, r4) short, gaze fixed / static
   (r1) leave no time to experience ego-motion parallax effects
   (r3) narrow foveation methods selection-->
   <!--DNC: The content of the user study is arranged in this way to minimize the interference of other non-core factors on the experimental results (such as the performance defect of the eye tracker, the VR novice is not familiar with the equipment and the environment, etc.) 
   The focus of this article is not to study how to provide more accurate foveation rendering, so the most typical method is selected as the ground truth-->
The user study is simulating how user experience different conditions. 
Similar to similar studies [REF], the short duration is guarantees fairness among trials (avoid shifting LnXXXX).
We asked users to fixate at certain gaze positions so everyone's views were consistent across conditions and participants. (TBD) We set duration to 300ms so it is close to saccade experience(?)

<!-- 3/9
1) 300ms > per frame
2) even not enough, already proved ours > nerf

motion parallax is exactly our beneifts than panorama synthesis we will compare with panorama
it is fair for each view

-->

With eye-tracking in VR, foveated rendering is chose as GT because that has been shown to be perceptually identical to a full resolution rendering. Higher resolution on the peripheral range has minimal effects on perception. We have done an orthogonal pilot study verifying the identicality between F-GT and GT. We will include the details. <!--(TODO: read the claim of the 2002 paper.)-->

## Publish source code (r3.3, r5.1):
Yes, we will make the entire source code publicly open, including models and data.
<!--
About written (r3.1, r3.2, r4.14, r5.6):
Qi: just ignore these
   (r3) improve clarity / bib consistency
   (r4) typo / style
   (r5) data flow figures; 
   (r5) what is the overall loss function, how was the training done (batches? training dataset?) ZH: we can try best to increase R5's score by just claiming we will add whatever [s]he thought is helpful.
-->
## Individual
Review 1

<!--Qi: what is this parapgraph for?-->
0.5MB is the upper bound of the storage space we need (or the storage space we need is unrelated with the complexity and scale of the scene), and the storage space required for traditional 3D resources is closely related to the fineness of the model/texture (no upper bound)*

* Foveation rendering as ground truth (r1.5)

Review 2
* in angular resolution (you need fewer pixels per sphere, but also fewer spheres), this amounts to (6/10)^3 *10 = 0.06MB per frame for the maximum 6PPD resolution used in this paper.

* Angular resolution: 6.4 PPD vs ground truth 13 PPD? (r2.1)
  <!--DNC: the users are asked to choose an overall perceptually better quality, not the fovea region only. as LPIPS shown, we have better quality in mid and far periphery, 
  ZH: the above explanation is not even clear to me.
  ZH: We are using F-GT (foveated ground truth images) instead of high resolution ground truth.
  (ZH: I am confused again about the reason we choose F-GT, does that mean our final retina images has passed two foveated rendering here?)
* Experiment about storage? (r2.3)
    What baseline requires 100MB of storage claimed in the introduction? The original 3D asset? There must be a middle ground which improves storage somewhat, without having to evaluate a neural network for every output pixel.
* Other multi-sphere representation (r2.4)
    In particular, I would also like to see how a multi-sphere image representation like [Broxton2020] or [Attal2020] would perform. For example, in [Broxton2020] they report 10MB per frame for 10PPD data. Since storage is cubic in angular resolution (you need fewer pixels per sphere, but also fewer spheres), this amounts to (6/10)^3 *10 = 0.06MB per frame for the maximum 6PPD resolution used in this paper.
Review 3 (None)

Review 4
* not comparable between neural and mesh representation (r4.2)
* ZH: 0.5MB storage for our model is not only for one view. (more to go here).
* ZH: 0.5MB is the upper bound of the storage space we need. The storage space we need is unrelated to the complexity and scale of the scene, however the storage space required for traditional 3D resources is closely relevant to the fineness of the model/texture, which has no upper bound.
* A fairer comparison: allow NeRF an additional 96 channels (r4.8)
   ZH: We are using the same parameters that were used in our setup. Nevertheless, the performance of such NeRF configuration is still very slow. 
  <!--DNC: Same network parameters means same network capacity. Use large network capacity to compare our small network capacity is unfair.-->
* Why our is better? foveat network: still represents high-res whole scene, why baseline is worse? (4.9)
  <!--DNC: We use different coordinates (Egocentric coordinates) while baseline uses 3D volume. We are for different tasks. 3D volume is not suitable for our inside-out view mode.-->
* Artifacts in video:
  1m30: double image artifacts (r4.5)
  1m03: black splotches, where they come from? (r4.10)
  ZH: we have black holes in mid-periphery (why?)
* Minor questions:
  *Misunderstanding: need depth reconstruction for 360 photos application. (r4.1)
  <!--DNC: The neural encoding the scene information, including the depth. So we do not need extra depth information in run-time. Depth information can be inferred by ray casting using output density-->
  Not understand Fig 7 (r4.11)
  <!--DNC: The bounding refers to the range of sampling in ray casting (the near and far plane), which affects the distribution of sampling distances. we set upper bound to 10000 in (a) to approximate the infinite because sampling are done in disparity space (1/distance)-->
  Isn't the representation just alpha composited front to back? (r4.12)
  <!--DNC: Ray cast is different from alpha blending. it is an integration of density. Refer to NeRF for the detail equation-->
  Is NeRF here also running on a TensorRT implementation? (r4.13)
  <!--DNC: No-->
  ZH: explain the potential effects here.

Review 5
* Hard to read (Not understand)
  "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions."
  --> how did this come about? 3 pairs of conditions x 2 scenes? (r5.2)
  ZH: 6 trials (2 scenes x 3 gaze positions) x 6 pairs (permutation of 2 conditions). We predefined 3 gaze positions for each scene. Participants were asked to fixate each predefined gaze position for each trial and made the choice.  
  "Each condition from an individual scene consists of 3 different gaze positions."
  --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped? (r5.3)
  "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image."
  --> What is meant by perceptual similarity? color similarity? (r5.4)
ZH: Perceptual similarity is not only color similarity. It indicates the similarity of human visual perception.IPIPS (TBD)
  "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together."
  --> What is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to? (r5.5)
ZH: IV: three conditions (F-GT, OURS, NeRF). DV: LPIPS value at eccentricity=5, 10, 15, ... ,105.

# Raw Collections
## Review 1
1. Target mobile device but experiment on high-end machine?
2. Static scene: mobile devices can store. how about dynamic scene?
3. No interaction
4. user study: fix gaze, short
5. foveation rendering as ground truth
6. restricted position and direction?

## Review 2
1. Angular resolution: 6.4 PPD vs ground truth 13 PPD?
2. Performance: hardware? distance from low latency, and low-storage VR/AR platforms?
3. experiment about storage?
4. MSI: 10PPD - 10 MB -> 6PPD - 0.06MB?
5. head volume? frequency of transmiting network?
6. powerful computer but low resolution. MSI can achieve same storage real-time on mobile
7. why ray-caster? why not bake to MSI? (Under constraint of no view-dependent effects, small motion range, low resolution)

## Review 3
1. Improvement of clarity: the introduction of the egocentric coordinates might be a bit dense/slow to read
2. Reference: make bibliography more consistent
3. Publish source code
4. User study: the selection of foveation methods to compare is somewhat narrow
5. The scope should be clearly stated: static and receive side (decoder) only

## Review 4
1. Misunderstanding: 360 photos application: need depth reconstruction, e.g., Serrano et al. 2019 IEEE VR.
2. not comparable between nerual and mesh representation
3. Encode is slow, viewing is instant, but creation is not
4. Occlusion: small.
5. double image artifacts at 1m30s in video
6. Static user study: only a subset of experience
8. A fairer comparison: allow NeRF an additional 96 channels
9. Why our is better? foveat network: still represents high-res whole scene, why baseline is worse?
10. Artifacts: 1m03 in video: black splotches, where they come from?
11. (Minor) Not understande Fig 7
12. (Minor) Isn't the representation just alpha composited front to back?
13. (Minor) Is NeRF here also running on a TensorRT implementation?
14. Typos/style:
    L154: "remarkably" <- can remove this; the reader can determine whether it is remarkable or not.
    L324: using 'd' for density here is not a huge problem but might be confused for 'depth'. Later on (L614) this is refered ot as 'a'.
    L1064: "unprecedented performance" <- please tone it down a bit; every technique is a trade-off somewhere.
    Duplicated citation of Attal et al.

## Review 5
1. Hard to read
2. "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions." --> how did this come about? 3 pairs of conditions x 2 scenes?
3. "Each condition from an individual scene consists of 3 different gaze positions." --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped?
4. "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image." What is meant by perceptual similarity? color similarity?
5. "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together." --> Unfortunately this is too concise an explanation to make sense -- what is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to?
6. suggestion of writing/figures.
